### 2013-05-13 Rcode/BFcompare.rmd
first commit

try comparing BF for normal vs servin and stephens bf

```{r}
lABF = function(k,T ){return(((k/2)*T^2))}
lBF = function(k,T,n){return((-n/2)*log(1-(k/n)*T^2))}

n=seq(1,100,length=100)
T=2
k=.4/T^2

plot(n,lBF(k,T,n)-lABF(k,T),type="l")
abline(h=0,col=2)
```
Note that lBF>lABF always. So lABF is "biased" towards the null?
Is that right? Seems wrong, as the normal assumption is somehow
less conservative than the t?

Of course x<log(1+x)
```{r}
x=seq(0,1,length=100)
plot(x,log(1+x))
abline(a=0,b=1)
```
which seems to imply it is true.

Now let's compare
H0: \hat{\beta} \sim S T_\nu
H1: \hat{\beta} \sim \beta + S T_\nu
with \beta \sim N(0,W)
Here S is an estimate of the standard error of betahat.

the BF is the integral over beta of 
the ratio of T densities:
```{r}
BFt=function(W,betahat,S,nu){
  beta = rnorm(1000000,0,sd=sqrt(W))
  bft=mean(dt((betahat-beta)/S,nu) / dt(betahat/S,nu))
  #bfn= mean(dnorm((betahat-beta)/S)/dnorm(betahat/S))
  bfn= dnorm(betahat,0,sd=sqrt(W+S^2))/dnorm(betahat,0,sd=S)
  Zeff = effective.Zscore(betahat/S,nu)
  Seff = betahat/Zeff
  betahat.eff = Zeff*S
  cat(pnorm(Zeff),pnorm(betahat.eff/S),Zeff,betahat,Seff,pnorm(betahat/Seff),"\n")
  bfn.Seff= dnorm(betahat,0,sd=sqrt(W+Seff^2))/dnorm(betahat,0,sd=Seff)
  bfn.beff= dnorm(betahat.eff,0,sd=sqrt(W+S^2))/dnorm(betahat.eff,0,sd=S)
  return(list(bft=bft,bfn=bfn,bfn.Seff=bfn.Seff,bfn.beff=bfn.beff))
}
BFt(1,10,1,2)

```

Now transform T to Z be
```{r}
effective.Zscore = function(T,nu){
  return(qnorm(pt(T,nu)))
}
Z = effective.Zscore(T,n)
plot(n,lBF(k,T,n)-lABF(k,Z),type="l")
abline(h=0,col=2)

```