### 2013-05-13 talks/ilike-slides.Rmd

% Apaptive Shrinkage and False Discovery Rates by Laplace Approximation
% Matthew Stephens
% 2013/5/13

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(cache=TRUE,autodep=TRUE)
dep_auto()
library("qvalue")
```

# Motivation

- A common problem: you have imperfect measurements
of many ``similar" things, and wish to estimate their values.

- Particularly common in genomics. For example,
 a very common goal is to compare the mean
expression (activity) level of many genes in two conditions.

# Example: Mouse Heart Data

```{r, include=FALSE}
setwd("~/Documents/git/BayesFDR/talks/")
## load Poisson_binomial and ash functions 
source("../../stat45800/code/Rcode/PoissonBinomial.funcs.R")  
source("../Rcode/ash.R") 
x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
```

- Data on 150 mouse hearts, dissected into left and right ventricle
(courtesy Scott Schmemo, Marcelo Nobrega)

# Example: Mouse Heart Data
```{r}
head(x)
```

```{r, echo=FALSE}
cc = x[, 2:5]
g = c(-1, -1, 1, 1)
ngene = dim(x)[1]

cc.assoc = counts.associate(cc, g, 1)
zdat.ash = cc.assoc$zdat.ash
zdat = cc.assoc$zdat

# two-sided test
ttest.pval = function(t, df) {
    pval = pt(t, df = df, lower.tail = T)
    pval = 2 * ifelse(pval < 0.5, pval, 1 - pval)
    return(pval)
}

tscore = zdat[3, ]/zdat[4, ]
pval = ttest.pval(tscore, df = 2)

```

# Borrowing Strength

- For gene $j$, denote the difference in mean between condition 0 and condition 1 by $$\beta_j:= \mu^0_j - \mu^1_j.$$

- As measurements are made on only two
samples in each condition, error in estimates
of $\beta_j$ are appreciable.

- Fundamental idea: measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.

# False Discovery Rates

-Standard practice is to estimate an effect size $\beta_j$ and standard error $s_j$ for each gene.

-Then convert this to a $p$ value for each gene, e.g. by a $t$ test on $\beta_j/s_j$.

-Then use the distribution of $p$ values to estimate the false discovery rate (FDR) at a given threshold.

# False Discovery Rates

```{r, echo=FALSE}
hist(pval,prob=TRUE)
qval = qvalue(pval)
abline(h=qval$pi0,col="red",lwd=2)
```

# False Discovery Rates

```{r, echo=FALSE}
h=hist(pval,prob=TRUE)
qval = qvalue(pval)
abline(h=qval$pi0,col="red",lwd=2)
rect(0,0,0.05,qval$pi0,col="red")
rect(0,qval$pi0,0.05,h$intensities[1],col="green")
abline(v=0.05,lwd=3)
```

#FDR methods are very widely used!

- Why?

- Partly, they make sense.

- Partly, they are generic. Once you have the $p$ values, you don't really care where they came from.

#Issues

- Different precision of different measurements 
is forgotten once the $p$ value is computed.

- Borrows information for testing, not for estimation.

#An Alternative

- Instead of summarizing each gene by a single number ($p$ value)
summarize by two numbers, $\hat\beta_j$ and its standard error $\hat\s_j$.

- Then do inference for $\beta_j$ by Laplace approximation.

# Example: Wavelets

# Reproducible research

This document is produced with **knitr**, **Rstudio** and **Pandoc**.

Here is my session info:

```{r session-info}
print(sessionInfo(), locale=FALSE)
```

# Misc issues

- the plots are too wide? use the chunk option `out.width` which will be used in `<img width=... />`, e.g. `out.width=400px`