### 2013-05-13 talks/ilike.2013.tex

% CVSId: $Id: Example.tex,v 1.1.1.1 2000/11/28 11:15:12 exupery Exp $
\documentclass[ignorenonframetext]{beamer}

\usepackage{tikz}
\usepackage{amsbsy}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{algorithm}
\usepackage{multirow}
\usepackage{algorithmic}
\usepackage{color} 
\usepackage{graphicx}
\usepackage{tabularx}
%\usepackage{hyperref}
\usepackage{eepic}
\usepackage{booktabs}
\usepackage{epic}
\def \hatd { \widehat{ d } }
\def \dhatin {\hatd^{\mbox{\tiny IN}}}
\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
\def \mvn{\text{MVN}}
\def\mn{\text{MN}}
\def \iw{\text{IW}}
\def \fp {fastPHASE\ }
\def\var{\rm Var}
\def\BF{\text{BF}}
\def\BFfull{\text{BF}_\text{all}}
\def\mvr{\text{BMVR}}
\def\one{\mbf{1}}
\def\zero{\mbf{0}}
\def\sxx{S_{xx}}
\def\syy{S_{yy}}
\def\rss{\text{\rm RSS}}
\def\I{I}
\def\D{D}
\def\U{U}
\def\p{p}
\def\g{G}
\def\bfav{BF_{\text{av}}}
\def\pall{p_\text{all}}

%\ColorFoot{2}
\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
                                % mode too
\def\BF{\rm BF}
\def\hfull{\textcolor{red}{H_\text{all}}}
\def\hnull{\textcolor{blue}{H_0}}

\mode<article>
{
  \usepackage{fullpage}
  \usepackage{pgf}
  \usepackage{hyperref}
  \setjobnamebeamerversion{sanger.beamer}
}

\mode<presentation>
{
  \usetheme{Dresden}

  \setbeamercovered{transparent}
}

\usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\def \fp {fastPHASE\ }

\title[Adaptive Shrinkage by Laplace Approximation]{Adaptive Shrinkage and False Discovery Rates by Laplace Approximation}
\author{Matthew Stephens} \date{i-like, Warwick,  May 2013 }

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Motivation}  
A common problem: you have imperfect measurements
of many ``similar" things, and wish to estimate their values.

This is particularly common in genomics. For example,
 a very common goal is to compare the mean
expression (activity) level of many genes in two conditions.

Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference.

Typically expression
measurements are made on only a small number of
samples in each condition - sometimes as few as one
sample in each condition. Thus the error in estimates
of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
in estimates of $\beta_j$ still greater.

Marcelo's Heart data as an example.

Wavelets as a second application.
\end{frame}

\begin{frame}{Borrowing Strength}
A fundamental idea is
that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
\end{frame}

\begin{frame}{False Discovery Rates}

What is actually done in practice?

\end{frame}

\begin{frame}{Genetic Association Studies}

So a typical GWAS analysis involves fitting millions of simple regressions, and testing effects for significance.

\bigskip

Notation: $y$ for phenotype, $g$ for genotype, $\beta$ for genetic effects:

$$y_i = \mu + \beta g_i + \epsilon_i \qquad (i=1,\dots,n)$$
\end{frame}

\begin{frame}{Genetic Association Studies and Heterogeneity}

We have been developing statistical methods for association mapping in multiple subgroups, incorporating heterogeneity of effects.
See also Lebrec et al (2010); Han and Eskin (2011,2012). 
\bigskip

Motivating examples include:
\begin{enumerate}
 \item The ``Global Lipids Consortium" Genome-wide Association Study meta-analysis (Teslovich et al, 2010).
 \item Gene expression analysis among multiple tissues (e.g.~Dimas et al, 2009).
 \end{enumerate}
\end{frame}

 \begin{frame}{Example 1: Global Lipids meta-analysis}

GWAS data from the Global Lipids consortium (Teslovich et al, 2010) on $>100,000$ individuals
from at least 25 separate studies.

\bigskip 
Four phenotypes: total cholesterol (TC), low-density lipoprotein cholesterol
(LDL-C), high-density lipoprotein cholesterol (HDL-C) and triglycerides (TG).

\bigskip
The original association analyses performed a {\it fixed-effects meta-analysis}.
That is, assume the effects in each subgroup  are all equal ($\beta_s = \beta \, \forall s$ ), and test $H_0: \beta=0$.

\bigskip
They reported a total of 95
SNPs as being associated with one or more phenotypes.

\end{frame}

\begin{frame}{Example 1: Global Lipids meta-analysis}

{\bf Question:} Given that this study involved 25 separate subgroups, many with quite different recruitment criteria
(e.g. some were recruited as cases for a particular disease; others were recruited as controls; etc),
would an analysis that allows for heterogeneity across studies identify more associations?
 \end{frame}

 \begin{frame}{Example 2: eQTL studies across multiple cell-types}

 Dimas et al (2009), measured expression data in 75 individuals, in 3 cell types: Fibroblasts, LCLs and T-cells.

 \bigskip

A key goal was to identify genetic variants associated with expression that were shared among
 cell types, or were specific to some subset of cell-types.  (Identifying eQTLs specific to individual cell types may shed insight into
 cell-type-specific regulation mechanisms.)

 \bigskip

 Original analysis performed association analysis separately in each cell type, and then looked at the overlap
 of the resulting associations.  The overlap was small (14\%), and they concluded that many eQTLs occur in only one cell type.

 \end{frame}

 \begin{frame}{Example 2: eQTL studies across multiple cell-types}

 {\bf Question:} Incomplete power may cause this analysis to underestimate sharing; does a joint analysis of all cell types
 come to the same conclusion?
\end{frame}

\begin{frame}{Methods}

Focus first on meta-analysis, where effect $\beta$ may vary across subgroups $s$:
$$y_{si} = \mu_s + \beta_s g_i + \epsilon_{si} \; \text{ with }\epsilon_{si} \sim \mathcal{N}(0,\sigma_s^2).$$

\bigskip

Primary goal of the meta-analysis is to identify SNPs for which there is strong evidence
against $H_0: \beta_s \equiv 0 \, \forall s$. 
\end{frame}

\begin{frame}{Alternatives to $H_0$}

To assess evidence against $H_0$ we introduce a set of alternative models, indexed
by parameters $\phi, \omega$, to be compared with $H_0$.

\bigskip

$H_1(\phi, \omega): \beta_s \text{ normally distributed about common mean } \bar \beta$.
$$\beta_s |\bar \beta \sim N(\bar\beta,\phi^2); \qquad \bar\beta \sim N(0, \omega^2).$$

\bigskip

{\bf Note 1:} $\phi=0$ corresponds to the usual ``fixed effects" alternative, $\beta_s = \bar \beta \, \forall s$.

\bigskip
{\bf Note 2:} Can alternatively work with the ``standardized" effect sizes, $b_s = \beta_s/\sigma_s$, which 
generally yields similar (but not identical) results.
\end{frame}

\begin{frame}{Bayes Factors}

The Bayes Factor
$$\BF(\phi,\omega) = p(y | g, H_1(\phi, \omega)) / p(y | H_0)$$
measures the support for $H_1(\phi,\omega)$ vs $H_0$, with large
values indicating strong evidence against $H_0$.

\bigskip
Although $\BF(\phi,\omega)$ depends on priors for nuisance parameters $(\mu,\sigma^2_s)$, it is not very sensitive,
and sensible default choices exist.

\bigskip 
Hyperparameters $\phi$ and $\omega$ must be chosen to reflect expected effect sizes and levels of heterogeneity
(but can average over several values to reflect uncertainty in choice of appropriate values).
\end{frame}

\begin{frame}{Computation}

$\BF(\phi,\omega)$ can be quickly and accurately approximated by Laplace approximation.

\bigskip

In the simplest cases these approximations depend only on the summary statistics
in each study, $\hat\beta_s$ and $\rm{se}(\hat\beta_s)$.
(Details: Wen and Stephens, 2011).
\end{frame}

\begin{frame}{Bayes Factors and standard test statistics}

This framework includes some commonly-used frequentist test statistics as special cases.
\bigskip

For example, if we allow $\omega$ to vary across SNPs according to the inverse
of the standard error of $\bar \beta$ then $BF(\phi=0, \omega)$ is monotonic with the weighted $Z$ score
\begin{equation} \label{weighted.Zscores.eqn}
\mathcal{Z} = \frac{\sum_s w_s Z_s}{\sqrt{\sum_{s'} w_{s'}^2}}
\end{equation}
where $Z_s=\hat\beta_s/ {\rm se}(\hat{\beta}_s)$ and $w_s =  {\rm se}(\hat{\beta}_s)^{-1}$. (Details: Wen and Stephens, 2011.)

\bigskip
In other words, we can see what implicit models for $\beta$ are assumed by standard methods.
\end{frame}

\begin{frame}{Example 1: Global Lipids meta-analysis}

{\bf Question:} Given that this study involved 25 separate subgroups, many with quite different recruitment criteria
(e.g. some were recruited as cases for a particular disease; others were recruited as controls; etc),
would an analysis that allows for heterogeneity across studies identify more associations?

 \bigskip

 \pause{{\bf Answer:} Not much!}
 \end{frame}

 \begin{frame}{Results: Global Lipids GWAS}

\begin{itemize}
\item Searched genome-wide for SNPs with strong signal when allowing
for heterogeneity ($\BF_{\rm het}>10^6$) but not when assuming no heterogeneity ($\BF_{\rm no-het}<10^6$).
\item 42 SNPs satisfied these criteria.
\item But 36 of these were driven by apparently strong associations in a single study (Framingham heart Study), and seemed likely to be due to data processing errors.
\item Two more showed similarly suspicious patterns (association in just one study, a subset of the WTCCC). 
\end{itemize}
\end{frame}

\begin{frame}{Results: Global Lipids GWAS}

\begin{table}[h!t]
\begin{center}
\begin{tabular}{ l c c c c }
\toprule
Phenotype & SNP & Gene & $\log_{10} ({\rm BF}_{\rm no-het})$ &  $\log_{10}({\rm BF}_{\rm het})$ \\
\midrule
LDL & rs1800978 & {\footnotesize  ABCA1} & 5.2  & 6.0  \\
TG  & rs1562398 & {\footnotesize nr KLF14} & 5.3 &  6.5 \\
%TG &  rs765547  & {\footnotesize Flanking LPL}   & 1.7 & 7.9 & 5.8  \\ 
HDL & rs11229165 & {\footnotesize nr OR4A16} & 4.6 & 6.4 \\
HDL & rs7108164 & {\footnotesize nr OR4A42P} & 4.2 & 6.3 \\ 
%HDL & rs11984900 & {\footnotesize N.A. } &-1.1 & 16.6 & 6.2 \\
%HDL & rs6995137 & {\footnotesize Flanking SFRP1} & -0.4 & 6.9 & 4.8 \\
\bottomrule
\end{tabular}
%\caption{\label{lipids.miss.tbl} Association signals that show strong association under the models allowing for heterogenetiy (${\rm ABF^{EE}_{cefn}} \text{ or } {\rm ABF_{maxH}^{EE}} \ge 10^6$) but less strong under a model with no heterogeneity (${\rm ABF^{EE}_{fix}}< 10^6$). It seems likely that the last two of these represent false positive associations, but we include them in the table for completeness (see text for discussion).}
\end{center}
\end{table}
\end{frame}

\begin{frame}{Example 2: eQTL sharing across cell-types}

 In meta-analysis application the primary goal was to reject the global null, $H_0: \beta_s =0 \, \forall s$.

 \bigskip

 Mapping eQTLs among multiple cell-types (subgroups) differs 
 in that we care more about {\it which} $\beta_s$ are non-zero, and patterns
 of sharing among subgroups.

 \bigskip

E.g. Dimas et al (2009), identified eQTLs separately in 3 cell types, and asked which eQTLs are shared among cell types (subgroups).

 \end{frame}

 \begin{frame}{Example 2: eQTL sharing across cell-types}

 To address this we expand our alternative models  $H_1(\phi,\omega)$ to
 allow that effects may be zero in some subgroups.

\bigskip

Introduce a \alert{configuration} $\gamma$ indicating which subgroups have non-zero effect.
\begin{itemize}
\item E.g. $\gamma=[110]$ corresponds to non-zero effect in the first two subgroups.
\end{itemize}

\bigskip

See also Han \& Eskin (PloS Genetics, 2012).

%\pause
%\bigskip
%An eQTL can have different effects in different tissues ($\beta_1 \neq \beta_2$):
%\begin{itemize}
% \item $b_s \sim \mathcal{N}(\bar{b}, \alert{\phi^2})$ and $\bar{b} \sim \mathcal{N}(0,\alert{\omega^2})$
%\item $\omega$: magnitude of the average effect
%\item $\phi$: magnitude of heterogeneity
% \item both are specified by a grid of values
%\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Bayesian Model Averaging and hierarchical modeling}

The support in the data for configuration $\gamma$ can be measured by the Bayes Factor
$$\BF_\gamma(\phi,\omega) = \frac{p(y | H_1(\gamma,\phi,\omega))}{p(y | H_0)}$$.

\bigskip

Overall evidence against $H_0$ can be measured by averaging over $\gamma, \phi, \omega$:
$\text{BMA} = \sum_{\gamma,\phi,\omega} \eta_{\gamma,\phi,\omega} \text{BF}_\gamma(\phi,\omega)$

\bigskip

Estimate proportions $\eta_{\gamma,\phi,\omega}$ using a hierarchical model to combine information across genes.

\end{frame}

%\begin{frame}
%\frametitle{Bayes Factor (BF) and hierarchical model (HM)}
%\begin{itemize}
%\item Quantify evidence for an eQTL in configuration $\gamma$:
%\bigskip
%$\text{BF}_\gamma = \frac{\mathsf{P}(\text{data } | \text{ eQTL in configuration }\gamma)}{\mathsf{P}(\text{data } | \text{ no eQTL in any tissue})}$
%\pause
%\bigskip
%\bigskip
%\item Estimate configuration proportions $\eta_\gamma$ by borrowing information across genes (pooling).
%\end{itemize}
%\end{frame}

%\begin{frame}
%\frametitle{Tools to fulfill the two goals}
%\begin{itemize}
%\item Detect eQTLs by Bayesian Model Averaging (BMA):
%\bigskip
%$\text{BMA} = \sum_\gamma \eta_\gamma \text{BF}_\gamma$
%\pause
%\bigskip
%\bigskip
%\item Identify in which tissue(s) they are active via their posterior:
%\bigskip
%\small{$\mathsf{P}(\text{SNP is in configuration }\gamma \; | \text{ data, SNP is eQTL}) = \frac{\eta_\gamma \text{BF}_\gamma}{\sum_\gamma \eta_\gamma \text{BF}_\gamma}$}
%\end{itemize}
%\end{frame}

%-----------------------------------------------------------------------------

%\section{Simulations}

%-----------------------------------------------------------------------------

%\section{Analysis of three cell types from Dimas \textit{et al.}}

\begin{frame}
\frametitle{Example 2: eQTL studies across multiple cell-types}

Dimas et al (2009), measured expression data in 75 individuals, in 3 cell types: Fibroblasts, LCLs and T-cells.

\bigskip

They identified eQTLs separately in each cell-type, and found small overlap of results (14\%).

\bigskip

{\bf Question:} Incomplete power may cause this analysis to underestimate sharing; does a joint analysis of all cell types
 come to the same conclusion?

\end{frame}

\begin{frame}
\frametitle{Joint Analysis Increases Power}
\begin{center}
  \begin{tikzpicture}
    \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width=\textwidth,height=\textheight,keepaspectratio=true]{figures/box-forest_weak-consistent_ENSG00000090924-rs755690}};
  \end{tikzpicture}
\end{center}
%\begin{center}
%  \small{Example of gene ENSG00000090924 and SNP rs755690.}
%\end{center}
\end{frame}

\begin{frame}
\frametitle{Gain in power from the joint analysis}
%\only<1>{
%\begin{center}
%%\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true,clip,trim=0cm 0cm 5.6cm 0cm]{dimas}%
%  \begin{tikzpicture}
%    \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true,clip,trim=0cm 0cm 5.6cm 0cm]{figures/dimas}};
%    \begin{scope}[x={(image.south east)},y={(image.north west)}]
%      % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
%      % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
%      % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y}; }
%      \draw[red,ultra thick,rounded corners] (0.19,0.64) rectangle (0.44,0.81);
%      \draw[red,ultra thick,rounded corners] (0.68,0.64) rectangle (0.94,0.81);
%    \end{scope}
%  \end{tikzpicture}
%\end{center}
%}
%\only<2>{
\begin{center}
\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true,clip,trim=11.6cm 0cm 0cm 0cm]{figures/dimas}%
\end{center}
%}
\end{frame}

\begin{frame}
\frametitle{Joint analysis suggests much more sharing of eQTLs}
%\begin{table}[!ht]
%\begin{tabular}{|c|c|c|}
%\hline
%Configuration & Hierarchical model & Tissue-by-tissue \\
%\hline
%F-L-T & \alert<2-3>{0.865} [0.807, 0.983] & \alert<2-3>{0.142} \\
%L-T & \alert<4>{0.046} [0.003, 0.104] & 0.056 \\
%F-L & 0.005 [0.000, 0.028] & 0.099 \\
%F-T & 0.001 [0.000, 0.016] & 0.056 \\
%F & \alert<3>{0.025} [0.000, 0.070] & \alert<3>{0.253} \\
%L & \alert<3>{0.051} [0.000, 0.115] & \alert<3>{0.265} \\
%T & \alert<3>{0.007} [0.000, 0.033] & \alert<3>{0.130} \\
%\hline
%\end{tabular}
%\begin{flushleft}\small{The results for the hierarchical model were obtained with the correlated Bayes Factors and the EM algorithm. The results for the tissue-by-tissue analysis were obtained by fixing an FDR of 0.05 in each tissue.}\end{flushleft}
%\end{table}
%\only<5>{
%\alert{Tissue-by-tissue analysis} doesn't account for different power between tissues $\Rightarrow$ likely to \alert{over-estimate tissue specificity}.
%}
\begin{center}
%\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true]{venn-diag_config-props_sep-vs-join}%
\begin{tikzpicture}
  \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true]{figures/venn-diag_config-props_sep-vs-join}};
  \begin{scope}[x={(image.south east)},y={(image.north west)}]
    % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
    % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
    % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y}; }
    \only<1>{
      \draw[red,ultra thick,rounded corners] (0.43,0.49) rectangle (0.57,0.6);
    }
%    \only<2>{
%      \draw[red,ultra thick,rounded corners] (0.14,0.67) rectangle (0.27,0.78);
%      \draw[red,ultra thick,rounded corners] (0.7,0.67) rectangle (0.84,0.78);
%      \draw[red,ultra thick,rounded corners] (0.43,0.12) rectangle (0.57,0.23);
%    }
    % \only<3>{
    %   \draw[red,ultra thick,rounded corners] (0.6,0.40) rectangle (0.7,0.51);
    % }
  \end{scope}
\end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Wrong tissue-specific call by the tissue-by-tissue analysis}
\only<1>{
  \begin{center}
    \begin{tikzpicture}
      \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width=\textwidth,height=\textheight,keepaspectratio=true]{figures/box-forest_wrong-specific_ENSG00000106153-rs4948093}};
      \begin{scope}[x={(image.south east)},y={(image.north west)}]
        % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
        % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
        % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y}; }

        %\only<2>{
          \draw[red,ultra thick,rounded corners] (0.77,0.65) rectangle (0.85,0.72);
                  \draw[red,ultra thick,rounded corners] (0.77,0.47) rectangle (0.85,0.54);
            \draw[red,ultra thick,rounded corners] (0.77,0.28) rectangle (0.85,0.35);
%              \draw[red,ultra thick,rounded corners] (0.77,0.65) rectangle (0.85,0.52);
          %\draw[red,ultra thick,rounded corners] (0.71,0.91) rectangle (0.92,0.99);
        %}
      \end{scope}
    \end{tikzpicture}
  \end{center}

  \begin{center}
    \small{Example of gene ENSG00000106153 and SNP rs4948093 (MAF=0.23).}

    \small{See also Ding \textit{et al.} (2010, AJHG).}
  \end{center}
}
\end{frame}

%-----------------------------------------------------------------------------

%\section{Conclusions and perspectives}

\begin{frame}[fragile]
\frametitle{The next challenge - more subgroups!}
\begin{itemize}
\item This ``configuration-based" framework can deal satisfactorily with, perhaps, 6-10 subgroups.
\item The NIH GTEX project is currently collecting data on upwards of 20 tissues.
\item More generally, in genomics, one might have hundreds of  ``observations" on each unit...
\item ... and, potentially, relevant covariates.
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{The next challenge - a general framework for data integration?}

\begin{itemize}
\item Summarize the ``data" on each SNP in each subgroup (or experimental condition) as $(\hat\beta, se(\hat\beta))$.
\item Arrange these in a big $p$ by $S$ matrix.
\item Goal: identify the elements that correspond to non-zero (or ``large") $\beta$, exploiting combined structure
across rows of the matrix.
\end{itemize}

Challenge: exploit the many available tools -- clustering, PCA, factor analysis,etc -- to do this in a flexible and powerful way.

\end{frame}

\begin{frame}
\frametitle{Acknowledgments}

\begin{itemize}
\item  Xiaoquan Wen,  Timoth\'ee Flutre, Jonathan Pritchard. 
\item Global Lipids Consortium for making data available
\item Manoulis Dermitzakis for expression data.
\item Funding: NHGRI and NIH GTEX consortium.
\end{itemize}

Selected References:
\begin{itemize}
\item Wen \& Stephens (2011, arXiv), Wen (2012, arXiv), Flutre et al (2013, PloS Genetics).
\item Han \& Eskin (2011, AJHG; 2012, PloS Genetics).
\item Ding \textit{et al.} (2010, AJHG).
\item Lebrec \textit{et al.} (2010, SAGMB).
\end{itemize}
\url{http://stephenslab.uchicago.edu/publications.html}
\end{frame}

\begin{frame}
\frametitle{Gain in power from the joint analysis}
\only<1>{
\begin{center}
%\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true,clip,trim=0cm 0cm 5.6cm 0cm]{dimas}%
  \begin{tikzpicture}
    \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true,clip,trim=0cm 0cm 5.6cm 0cm]{figures/dimas}};
    \begin{scope}[x={(image.south east)},y={(image.north west)}]
      % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
      % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
      % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y}; }
      \draw[red,ultra thick,rounded corners] (0.19,0.64) rectangle (0.44,0.81);
      \draw[red,ultra thick,rounded corners] (0.68,0.64) rectangle (0.94,0.81);
    \end{scope}
  \end{tikzpicture}
\end{center}
}
\only<2>{
\begin{center}
\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true,clip,trim=11.6cm 0cm 0cm 0cm]{figures/dimas}%
\end{center}
}
\end{frame}

%  \begin{frame}{Why Jointly Analyze?}
%\begin{enumerate}
%\item Power gains if eQTLs are shared among tissues.
%\item Without a joint analysis, very hard to assess tissue specificity due to lack of power.
% \end{enumerate}
% \end{frame}
% 
%

% Similar, but different, issues, arise when analyzing eQTLs in multiple tissues

%\begin{frame}{LDL Subfractions}  
%\begin{center}
%\includegraphics[width=2in,height=2in]{subfrac_top3_av.pdf}
%\end{center}
%\end{frame}

\begin{frame}{Where next?}
\begin{itemize}
\item Larger-scale problems (e.g.~GTEx collecting data on 30 tissues)
%\item More complex outcome structures (e.g. small time series).
\item Multi-SNP multi-phenotype? (e.g.~Verzilli et al (2005); Banerjee et al (2008)).
%\item Clustering problems generally (Genotype unobserved).
\item Dealing with non-normality; Outliers; Binary outcome with intermediate quantitative phenotypes.
\item ``Response" phenotypes. (Maranville et al, PloS Genetics, 2011).
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Some eQTLs may be shared; others tissue-specific}
\only<1>{
\begin{center}
\includegraphics[width=0.95\textwidth,height=0.8\textheight,keepaspectratio=true]{figures/ex_config11_var0-5_boxplot}%
\end{center}
}
\only<2>{
\begin{center}
\includegraphics[width=0.95\textwidth,height=0.8\textheight,keepaspectratio=true]{figures/ex_config10_var0-5_boxplot}%
\end{center}
}
\only<3>{
\begin{center}
\includegraphics[width=0.95\textwidth,height=0.8\textheight,keepaspectratio=true]{figures/ex_config00_var0-5_boxplot}%
\end{center}
}
\end{frame}

\end{document}