### 2013-05-14 talks/ilike-slides.Rmd
final draft of ilike slides; added papers

% Adaptive Shrinkage and False Discovery Rates by Laplace Approximation
% Matthew Stephens
% 2013/5/13

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(cache=TRUE,autodep=TRUE)
dep_auto()
library("qvalue")
```

# Outline

* Prelude
* Allegro (ma non troppo)

# Prelude

- Consider testing the null hypothesis $H_0: \beta=0$, vs the alternative $H_1: \beta \neq 0$ in the
logistic regression model:
$$\log \frac{p(Y_i=1|X_i=x)}{p(Y_i=0|X_i=x)}  = \mu + x \beta$$

- Specifically, consider computing the Bayes Factor
$$BF:=\frac{p(Y | X, H_1)}{p(Y| X ,H_0)}.$$

- In genome-wide association studies, we may wish to do this 
for millions of different genetic variants ($X$).

# Prelude

$$BF= \frac{\int p(Y | \mu, \beta, X) p_1(\mu,\beta | X) \, d\mu d\beta}{\int p(Y | \mu, \beta=0, X) p_0(\mu | X) \, d\mu},$$
where $p_0$ and $p_1$ denote priors under $H_0$ and $H_1$.

- These integrals generally don't have closed forms, but being 
low-dimensional they are simple to approximate.

- For $p_1: \beta \sim N(0,\phi^2)$, 
Wakefield, 2009 (see also Johnson, 2008) suggested a particularly simple \emph{Approximate Bayes Factor} (ABF)
based on the maximum likelihood estimate, $\hat\beta$, and 
its (estimated) standard error $s$.

# Prelude

$$ABF = \sqrt{1-k} \exp(0.5 k T^2)$$
where $k:= \phi^2/(s^2 + \phi^2)$ and $T:= \hat\beta/s$.

- ABF arises if we assume $\hat{\beta} | s, \beta \sim N(\beta, s^2)$ and treat $\hat{\beta}$ as the observed ``data".

- Equivalently ABF can be derived as a ``Laplace approximation",
approximating the likelihood
$L(\beta)$ as Normal, centered on $\hat\beta$,
with variance $s^2$: $$L(\beta) \propto \exp[- 0.5(\beta-\hat\beta)^2/s^2].$$

# Comments on ABF

* This is not the moxt accurate Laplace approximation one might consider.

* However, it has some nice features.  
    * The approximation is independent of prior.
    * Applicable to any regression where $\hat\beta$ and $s$ are available.
    * Easily computed using results of standard software or published analyses (e.g. CI).

* A simple transformation of $T$ can improve accuracy for small samples (analagous to $t$ test vs $Z$ test); Wen and Stephens, Arxiv.

# Extensions of ABF

- Similar ideas can be used to compute 
ABFs in slightly more complex settings.

- Eg In Wen and Stephens, we consider
$S$ subgroups, and approximate the BF for
$H_0: \beta_s =0$ for all $s$, vs a general alternative
$H_0: \beta_s \neq 0$.

# Allegro (ma non troppo)

- The problem: you have imperfect measurements
of many ``similar" things, and wish to estimate their values.

- Particularly common in genomics. For example,
 a very common goal is to compare the mean
expression (activity) level of many genes in two conditions.

# Example: Mouse Heart Data

```{r, echo=FALSE}
setwd("~/Documents/git/ash/talks/")
## load Poisson_binomial and ash functions 
source("../../stat45800/code/Rcode/PoissonBinomial.funcs.R")  
source("../Rcode/ash.R")

x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
xx = rowSums(x[,2:5])
x = x[xx>0,]
xx = xx[xx>0]
```

- Data on 150 mouse hearts, dissected into left and right ventricle
(courtesy Scott Schmemo, Marcelo Nobrega)

```{r echo=FALSE}
head(x)
```

```{r, echo=FALSE}

cc = x[, 2:5]

g = c(-1, -1, 1, 1)
ngene = dim(x)[1]

cc.assoc = counts.associate(cc, g, 1)
zdat.ash = cc.assoc$zdat.ash
zdat = cc.assoc$zdat

# two-sided test
ttest.pval = function(t, df) {
    pval = pt(t, df = df, lower.tail = T)
    pval = 2 * ifelse(pval < 0.5, pval, 1 - pval)
    return(pval)
}

tscore = zdat[3, ]/zdat[4, ]
pval = ttest.pval(tscore, df = 2)
qval = qvalue(pval)

highxp = xx>1000 # select high expressed genes
pval.high = pval[highxp]
qval.high = qvalue(pval.high)
cc.assoc.high = counts.associate(cc[highxp,],g,1)
zdat.ash.high = cc.assoc.high$zdat.ash
zdat.high = cc.assoc.high$zdat

```

# False Discovery Rate analyses

- Standard practice: analyses use False Discovery Rates

    - e.g. Benjamini and Hochberg, 1995; Storey and Tibshirani, 2003, which have roughly 18k and 4k citations respectively!

- Typical analysis proceeds roughly as follows:

    - Estimate an effect size $\beta_j$ and standard error $s_j$ for each gene.

    - Convert this to a $p$ value for each gene, e.g. by a $t$ test on $\beta_j/s_j$.

    - Use the distribution of $p$ values to estimate the false discovery rate (FDR) at a given threshold.

# False Discovery Rates

```{r, echo=FALSE,fig.height=4,fig.cap=""}
hist(pval,prob=TRUE,main="p value distribution, all genes",ylim=c(0,4),xlab="p value")
abline(h=qval$pi0,col="red",lwd=2)

```

# False Discovery Rates

```{r, echo=FALSE,fig.height=4,fig.cap="FDR=0.21"}
h=hist(pval,prob=TRUE,main="p value distribution, all genes",ylim=c(0,4),xlab="p value")
qval = qvalue(pval)
abline(h=qval$pi0,col="red",lwd=2)
rect(0,0,0.05,qval$pi0,col="red")
rect(0,qval$pi0,0.05,h$intensities[1],col="green")
abline(v=0.05,lwd=3)
```

# FDR problem: different genes have different precision/power

```{r, echo=FALSE,fig.height=4,fig.cap=""}
hist(log10(xx),main="Counts vary considerably across genes", xlab="log10(counts)")
```

# FDR problem: lower count genes, less power, add noise

```{r, echo=FALSE,fig.height=4,fig.cap=""}
h=hist(pval[xx<1000],prob=TRUE,ylim=c(0,4),main="p values, low count genes",xlab="p value",breaks=seq(0,1,length=21))
qval.low=qvalue(pval[xx<1000])
abline(h=qval.low$pi0,col="red",lwd=2)
rect(0,0,0.05,qval.low$pi0,col="red")
abline(v=0.05,lwd=3)
```

# FDR problem: higher count genes, more power
```{r, echo=FALSE, fig.height=4,fig.cap="FDR=0.10"}

h=hist(pval.high,prob=TRUE,ylim=c(0,4),main="p values, high count genes",xlab="p value")

abline(h=qval.high$pi0,col="red",lwd=2)
rect(0,0,0.05,qval.high$pi0,col="red")
rect(0,qval.high$pi0,0.05,h$intensities[1],col="green")
abline(v=0.05,lwd=3)
#qval.high$q[pval.high<0.05] gives FDR for p<0.05
```

# Adaptive Shrinkage

- Fundamental idea: use hierarchical modelling so measurements of $\beta_j$ for each gene improve inference for $\beta$ at other genes.

- Despite a long-standing literature on these types of methods - e.g. Greenland and Robins 1991, Efron and Tibshirani 2002, Gelman et al 2012 -  they are much less widely used (in genomics at least).

- Possibly this is due, in part, to the lack of a simple, flexible, and generic implementation?

# Generic adaptive shrinkage via Laplace approximation

- Summarize data on each gene by two numbers, $\hat\beta_j$ and its standard error $s_j$. (a la Wakefield; Greenland and Robins 1991)

- Approximate likelihood for $\beta_j$ by 
$$L(\beta_j) \propto \exp(-0.5 (\beta_j - \hat\beta_j)^2/s_j^2).$$
(``Laplace Approximation")

- Borrow information by assuming $\beta_j$ are iid $\sim g(\cdot; \pi)$, where $\pi$ are hyperparameters to be estimated.

- Letting $g(\cdot; \pi)$ be a mixture of normal distributions provides
both flexibility, and analytic calculations.

    - very small variances can capture effects that are ``effectively" zero.

# An important special case

- Focus on the special case where $g(\cdot; \pi)$ can be
assumed unimodal and symmetric about zero.

- Then the posterior mean, $E(\beta_j | \hat\beta, s, \hat\pi)$ is a ``shrinkage" estimate of $\beta_j$.

- And $p(\beta_j > 0 | \hat\beta, s, \hat\pi)$ can be used
to identify $j$ for which the sign of $\beta_j$ can be confidently determined (analogous to test of $\beta_j =0$; Gelman et al, 2012).

- Because $\pi$ is estimated from the data, the amount
of shrinkage is adaptive to the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each gene.

# Example: ASH applied to mouse data

```{r fig.height=4,echo=FALSE,fig.cap=""}
hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
```

# Example: ASH applied to mouse data

```{r fig.height=4,echo=FALSE,fig.cap=""}
hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
t=seq(-2,2,length=201)
lines(t,dnorm(t,0,sd=0.032), col=2, lwd=2)
```

# Example: ASH applied to mouse data

```{r fig.height=4,echo=FALSE,fig.cap=""}
hist(zdat.ash$PosteriorMean,main="Shrunken estimates",xlab="shrunk betahat",xlim=c(-0.1,0.1),prob=T)
lines(t,dnorm(t,0,sd=0.032), col=2, lwd=2)
```

# Shrinkage is adaptive to information

```{r, echo=FALSE,fig.height=4}
#hist(zdat[3,])
#plot(zdat[3,],zdat.ash$PosteriorMean,xlim=c(-0.2,0.2))
#points(zdat[3,16677],zdat.ash$PosteriorMean[16677],col=2,pch=16)
#points(zdat[3,16079],zdat.ash$PosteriorMean[16079],col=2,pch=16)
#x[16677,]
#x[16079,]
```

```{r, echo=FALSE}
#temp = counts.associate(cc,c(-1,1,-1,1),1)
#temp.tscore = temp$zdat[3,]/temp$zdat[4,]
#temp.pval = ttest.pval(temp.tscore,df=2)
#plot(temp$zdat.ash$localfdr,temp.pval)
#identify(temp$zdat.ash$localfdr,temp.pval)
```

```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
```

# Shrinkage is adaptive to information

```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
points(zdat.ash$localfdr[15325],pval[15325],col=2,pch=16)
points(zdat.ash$localfdr[16123],pval[16123],col=2,pch=16)
```

# Shrinkage is adaptive to information

```{r, echo=FALSE}
cbind(x[,1:5],pval,zdat.ash$localfdr)[c(15325,16123),]
```

# Summary: FDR vs ASH

- Both provide a rational approach to identifying ``significant" findings.

- Both are generic and modular: once you have the summary data, you can forget where they came from.

- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values) precision of different measurements 
can be better accounted for.

- ASH borrows information for estimation, as well as testing.

# Next steps?

- Extend to allow $g(\cdot;\pi)$ to depend on covariates $X$.

- Extend to allow for correlations in the measured $\hat\beta_j$.

# Thanks

- to the several postdoctoral researchers and students
who have worked with me on related topics.

- Especially: William Wen, Timothee Flutre, Scott Powers, Heejung Shim, Zhengrong Xing, and Ester Pantaleo.

- And to the NIH for funding, and i-like for inviting me.

# Reproducible research

- This document is produced with **knitr**, **Rstudio** and **Pandoc**.

- For more details see my \tt{stephens999/ash} repository at http://www.github.com/stephens999/ash

- Website: http://stephenslab.uchicago.edu

# Pandoc Command used

pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  ilike-slides.md -o ilike-slides.pdf

Here is my session info:

```{r session-info}
print(sessionInfo(), locale=FALSE)
```

# FDRs for higher count genes affected by lower count genes

```{r,echo=FALSE,fig.height=4,fig.cap=""}
plot(qval.high$q, qval$q[xx>1000],main="",xlab="FDR, high count genes", ylab="FDR, all genes")
```

# Some odd things in the data

```{r, echo=FALSE,fig.height=3,fig.width=3}
  plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
  x[tail(order(zdat[3,])),]
```