### 2013-05-20 Rcode/intro.rmd
extended intro.rmd

### Project Goals

The initial goal of the ASH (Adaptive SHrinkage) project is to provide simple, generic, and flexible methods to derive ``shrinkage-based" estimates and credible intervals for unknown quantities $\beta=(\beta_1,\dots,\beta_J)$, given only estimates of those quantities ($\hat\beta=(\hat\beta_1,\dots, \hat\beta_J)$) and their corresponding estimated standard errors ($s=(s_1,\dots,s_J)$).

Although shrinkage-based estimation can be motivated in various ways, our key goal here is to combine information across the multiple measurements $j=1,\dots,J$ to improve inference for each individual $\beta_j$. By improved inference, we mean both
improved average accuracy of point estimates, which is
the traditional focus of shrinkage-based methods, \emph{and} improved assessments of uncertainty.

By ``adaptive" shrinkage we 
have two key properties in mind. First, the appropriate amount of shrinkage is determined from the data, rather than being pre-specified. Second, the amount of shrinkage undergone by each $\hat\beta_j$ will depend on the standard error $s_j$: measurements with high standard error will undergo more shrinkage than measurements with low standard error.

As an important special case, we intend these methods to address the "multiple comparisons" problem, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero. Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\hat\beta_j/s_j$,
and then applying a generic procedure, such as that of Benjamini 
and Hochberg (1995?) or Storey (2001?), designed to control or
estimate the false discovery rate (FDR) or the positive FDR (Storey, 2001?). In essence we aim to provide analagous
generic methods that work directly with two numbers for each 
measurement $(\hat\beta_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic). Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the 
uncertainty in each measurement $\hat\beta_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis. The potential for shrinkage-based estimation to
help address multiple comparisons problems has been highlighted
previously, including Greenland and Robins (1991),
Efron (2008) and Gelman et al (2012). [Note, check also Louis, JASA, 1984]

### Methods Outline

The methods are based on treating the vectors $\hat\beta$
and $s$ as ``observed data", and then performing inference for $\beta$ from these observed data, using a standard hierarchical modelling framework
to combine information across $j=1,\dots,J$.

Specifically, 
we assume that the true 
$\beta_j$ values are independent
and identically distributed from some distribution $g(\cdot ; \pi)$, where $\pi$ is a hyper-parameter to be estimated. 
Then, given $\beta$, we assume that $(\hat\beta_j,s_j)$ are independent across $j$, and depend only on $\beta_j$. Putting these together, the joint model for the unobserved $\beta$ and the observed $\hat\beta, s$ is:
\begin{align}
p(\hat\beta, s, \beta | \pi) & = \prod_j g(\beta_j ; \pi) p(\hat\beta_j, s_j | \beta_j) \\
& = \prod_j g(\beta_j ; \pi) L(\beta_j; \hat\beta_j,s_j).
\end{align}
The specific choices of $g$ and $L$ are described below.

We fit this hierarchical model using the following "Empirical Bayes" approach. First we estimate the hyper-parameters $\pi$ by maximizing the likelihood
$$L(\pi; \hat\beta, s) := p(\hat\beta, s | \pi) = \int p(\hat\beta, s, \beta | \pi) d\beta.$$
Then, given this estimate $\hat\pi$, we compute the conditional distributions $$p(\beta_j | \hat\pi, \hat\beta, s) \propto g(\beta_j; \pi) L(\beta_j; \hat\beta_j, s_j).$$ 
In principle we would
prefer to take a full Bayes approach that accounts for uncertainty in $\pi$, but, at least for now, we compromise this principle for the simplicity of the EB approach.

[put picture of hierarchical model here]

The conditional distributions $p(\beta_j | \hat\pi, \hat\beta, s)$ 
encapsulate uncertainty in the values for $\beta_j$, combining information across
$j=1,\dots,J$. The combining of the information occurs through estimation of
$\pi$, which involves all of the data, and it is

These conditional distributions can be conveniently summarized
in various ways, including point estimates (e.g. the posterior means or medians),
and credible intervals/regions.

The key components of this hierarchical model
are the distribution $g$ and the likelihood $L(\beta_j; \hat\beta_j, s_j)$. We make the following choices for these.

1. The likelihood for $\beta_j$ is normal, centered on $\hat\beta_j$, with standard deviation $s_j$.
That is, 
$$L(\beta_j; \hat\beta_j, s_j) \propto \exp[-0.5(\beta_j-\hat\beta_j)^2/s_j^2]. \quad (**)$$

2. The distribution $g(\cdot; \pi)$ is a mixture of zero-centered normal distributions, 
$$g(\cdot; \pi) = \sum_{k=1}^K pi_k N(\cdot; 0, \sigma^2_k).$$
In practice, we currently fix the number of components $K$ to be large, and take the variances $\sigma_1<\sigma_2<\dots<\sigma_K$ to be fixed, and vary from very small (possibly 0), to very large --  sufficiently large that typically $\hat\pi_K=0$.

The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways. For example, we can write $p(\hat\beta_j, s_j | \beta_j) = p(\hat\beta_j | s_j, \beta_j)p(s_j | \beta_j)$. Now, if we are willing to assume that
$s_j$ alone contains no information about $\beta_j$, or equivalently that $p(s_j | \beta_j)$ does not depend on $\beta_j$, then
$L(\beta_j) \propto p(\hat\beta_j | s_j, \beta_j)$,
and if $\hat\beta_j | s_j, \beta_j \sim N(\beta_j, s_j^2)$,
as is often asymptotically the case, then we obtain the likelihood (**) above.

An alternative motivation is to think of this as a normal
approximation to the likelihood from the 
raw data $D_j$ that were used to compute
$\hat\beta_j$ and $s_j$. Then if we observed these data
the likelihood for $\beta$ would be 
$p(D_j | \beta_j)$, and a Taylor series expansion of the log likelihood around the maximum likelihood estimate $\hat\beta_j$  yields $$l(\beta_j) \approx l(\hat\beta_j) + 0.5* (\beta_j - \hat\beta_j)^2 l''(\hat\beta_j).$$ [Fill in details?]

The idea of using a mixture of normal distributions for $g$ 
also seems very natural: mixtures of normals provide a flexible family of distributions able to provide a good approximation to any true underlying $g$; and 
when combined with the normal likelihood they give
an analytic form for the conditional distribution $p(\beta_j | \hat\pi, \hat\beta_j, s_j)$ (also a mixture of normals).
The constraint that these normals be centered at zero may seem initially less natural. 
Certainly this constraint could be relaxed in principle.
However, we view it as a convenient way to impose an assumption 
that $g$ is unimodal with its mode at 0,
which we view as a plausible assumption in many 
settings, and one which may be helpful to avoid
"overfitting" of $g$. (Using normal distributions centered at 0
also imposes an assumption
that $g$ is symmetric about zero, which we view as less
plausible, but represents a compromise between simplicity
and flexibility.
In cases where this assumption seems wildly inappropriate one could perhaps
improve results by applying the
model separately to positive and negative values of $\betahat$.)

Finally, using a large number of normal components with a wide range of variances,
rather than, say, a smaller number of components with the variances
to be estimated, is simply for computational convenience. With fixed
variances there exists a very simple EM algorithm to maximise the likelihood
in $\pi$. Obtaining maximum likelihood estimates for the variances could certainly
be implemented with a little
more work, but it is unclear whether this would result in practically-important gains in many situations of interest.

### Computation Outline

As outlined above, we fit the model using the following Empirical Bayes procedure:
1. Estimate $\pi$ by maximizing the likelihood $L(\pi)$.
2. Compute the conditional distributions $p(\beta_j | \hat\beta_j, s_j, \hat\pi)$.

Using a normal likelihood $L(\beta_j)$, and assuming
$g$ to be a mixture of normal distributions with fixed variances, 
yields a simple EM algorithm
for estimating $\pi$ in Step 1, and simple analytic forms for the conditional
distributions in Step 2.

### A simple example

Load in some functions.
```{r}
setwd("~/Documents/git/ash/Rcode/")
set.seed(32327)
## load Poisson_binomial and ash functions 
source("../Rcode/ash.R") 
library("qvalue")
```

Here we simulate data, effectively as in Efron, 2008, Section 7.
```{r}
truenull = c(rep(0,1000),rep(1,9000))
beta = c(rnorm(1000,-3,1),rep(0,9000))
s= rep(1,10000)
betahat = beta+rnorm(10000,beta,s)

beta.ash=ash(betahat,s)
#compute the usual zscore and corresponding p value
zscore = betahat/s
pval = pchisq(zscore^2,df=1,lower.tail=F)
qval = qvalue(pval)

plot(betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,4),xlim=c(-7,4))
abline(h=0)
abline(a=0,b=1,col=2)

plot(qval$q,2*beta.ash$qval,main="comparison of ash and q value qvalues")
abline(a=0,b=1)

o = order(beta.ash$qval)

plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=2,type="l")
lines(cumsum(truenull[o])/(1:10000),2*beta.ash$qval[o])
abline(a=0,b=1)
```

It seems that in this case the ash q values underestimate the
FDR slightly. Possibly this is the assymetry.
Try shrinking positive and negatives separately:
```{r}
pos= betahat>0
betapos.ash=ash(betahat[pos],s[pos])
betaneg.ash = ash(betahat[!pos],s[!pos])
lfdr = rep(0,length(betahat))
lfdr[pos] = betapos.ash$localfdr
lfdr[!pos] = betaneg.ash$localfdr
qv = qval.from.localfdr(lfdr)
o = order(qv)
plot(cumsum(truenull[o])/(1:10000),2*qv[o],type="l")
abline(a=0,b=1)
```
No, that's not it.

So maybe it is the non-monotonic behaviour of the true beta values in this case.
Let's try putting in a component at -1.5 to make the data fit this assumption better:

```{r}
truenull = c(rep(0,2000),rep(1,8000))
beta = c(rnorm(1000,-3,1),rnorm(1000,-1.5,1),rep(0,8000))
s= rep(1,10000)
betahat = beta+rnorm(10000,beta,s)

beta.ash=ash(betahat,s)
#compute the usual zscore and corresponding p value
zscore = betahat/s
pval = pchisq(zscore^2,df=1,lower.tail=F)
qval = qvalue(pval)
plot(qval$q,2*beta.ash$qval,main="comparison of ash and q value qvalues")
abline(a=0,b=1)

o = order(beta.ash$qval)

plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=2,type="l")
lines(cumsum(truenull[o])/(1:10000),2*beta.ash$qval[o])
abline(a=0,b=1)

```

### Motivation

It is common in statistics that you measure many ``similar" things imperfectly, and wish to estimate their values. The situation arises commonly in the kinds of genomics applications I am often involved in, but also in other areas of statistics.
In genomics, for example, a very common goal
is to compare the mean
expression (activity) level of many genes in two conditions.
Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
measurements are made on only a small number of
samples in each condition - sometimes as few as one
sample in each condition. Thus the error in estimates
of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
in estimates of $\beta_j$ still greater.

A fundamental idea is
that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.

Note on multiple comparisons: it isn't really a "problem" but an "opportunity". This viewpoint also espoused by Greenland and Robins. It isn't the number of tests that is relevant (the false
discovery rate at a given threshold does not depend on the number of tests). It is the *results* of the tests that are relevant.