### 2013-08-28 Rcode/jash_test.rmd
Add things for joint ash

Model
========================================================
The observed values is in a $N\times n$ matrix $Y$ ($N$ tests and $n$ replicate observations for each tests), where \[y_{jk}|\beta_j,\tau_j\sim N(\beta_j,1/\tau_j),\ k=1,2,...,n,\ j=1,2,...,N\]
We assume that $\beta_j$ (mean), $\tau_j$ (precision) come from the Normal-Gamma mixture prior with $M\times L$ components:
\[\beta_j|\tau_j\sim N(\mu_j,1/(c_l\tau_j)),\quad \tau_j\sim Gamma(a_m,b_m)\quad w.p. \pi_{ml}\]
where $\sum_{m,l}\pi_{ml}=1,\ m=1,...,M,\ l=1,...,L$. Here $a_m>0,b_m>0,c_l>0,\mu_j$ are known constants (default setting $\mu_j=0$). Note that $c_l$ can be chosen as $\infty$, and thus result in a point mass at zero in $\beta$'s prior.

Posterior distribution
=======================
Hence, the conjugate postrior dis
tribution of $(\beta_j,\tau_j)$ given $Y$ ias also a Normal-Gamma mixture distribution, i.e. $\tau_j|Y\sim Gamma(\tilde{a},\tilde{b})$, $\beta_j|\tau_j,Y\sim N(\tilde{mu},1/\tilde{\tau})$. The full posterior distribution for $\beta_j$ is $P(\beta_j|Y)=\int P(\beta_j|\tau_j,Y)P(\tau_j|Y)d\tau_j$, which is hard to compute. For convenience, we can try using the distribution of $\beta_j|\hat{\tau_j},Y$ for further inference, where $\hat{\tau_j}=E(\tau_j|Y)$. Later we will compare the sampled $P(\beta_j|Y)$ to $P(\beta_j|\hat{\tau_j},Y)$.

Note: $P(\beta_j|Y)$ is feasible to compute, though very messy. I am working on it...

Estimate hyper-parameters
=======================
We estimate the hyper-parameters $\pi$ by maximizing the likelihood $L(\pi;Y)=\int P(Y,\beta,\tau|Y)d\beta d\tau$.

Note: I find that sometimes if $a_m,b_m,c_l$ are not well chosen (e.g. the resulting $Var(Y|\beta,\tau)$ for each components overlap), then it is hard for the EM algorithm to converge.

FDR and q-values
================
Same as ash, we use false sign rate for inference.

Some examples
=============

```{r}
setwd("/Volumes/PERSONAL/MS")
set.seed(22222)
source("../Rcode/ash.R") 
source("../Rcode/jash.R")

#simulate N observations with n replicates for each, Nnull under the null
#with altmean and altsd being the mean and sd of beta under the alternative
simdata = function(N, Nnull, altmean, altsd, betasd, n) {
  null = c(rep(1, Nnull), rep(0, N - Nnull))
  beta = c(rep(0, Nnull), rnorm(N - Nnull, altmean, altsd))
  Y = matrix(rnorm(N*n, rep(beta,n), rep(betasd,n)), ncol=n)
  return(list(null = null, beta = beta, Y = Y))
}
```

Simulate 10000 tests, with 2000 alternatives, β∼N(0,sd=3), and 8000 null values (β=0).
```{r}
ss = simdata(10000, 8000, 0, 3, 1, 3)
beta.jash = jash(ss$Y)
beta.jash.auto = jash(ss$Y,auto=TRUE)
```

We use ash for comparison. Here we use the test-wise sample-mean and sqrt(sample-variance/n) as the inputs of ash. (So the sebetahat here is not the true standard error of betahat!)
```{r}
betahat = apply(ss$Y,1,mean)
sebetahat = apply(ss$Y,1,sd)/sqrt(dim(ss$Y)[2])
beta.ash.auto = ash(betahat,sebetahat,auto=TRUE)
```

The histogram of true $\beta$:
```{r}
hist(ss$beta, prob = TRUE, breaks = seq(-8, 8, length = 20))
```

Both jash and ash give the Posterior means $E(\beta|Y)$, compare the shrinked means using two methods:
```{r}
plot(betahat,beta.jash$PosteriorMean,xlab='Observed betahat',ylab='Posterior mean',cex=0.5,pch=20)
points(betahat,beta.ash.auto$PosteriorMean, col=2, cex=0.5,pch=20)
title('Black: jash; Red: ash.auto')
```

Compare the sampled full posterior distribution $P(\beta|Y)$ to the conditional posterior distribution $P(\beta|\hat{tau},Y)$, where $\hat{tau}=E(\tau|Y)$. Take the first two tests for illustration. It seems that in this case, the conditional distribution fits the full distribution well. (However, I am worried that it will make difference when $\tau$ has large scale or wide range. NEED TESTING!)

```{r}
# Compare sampled distribution P(beta|Y) to the normal mixture distribution P(beta|Y,hat(tau))
# Red lines are the conditional posterior distribution P(beta|Y,hat(tau)) 
par(mfrow=c(1,2))
sample.jash = posterior_sample_jash(beta.jash$post,1000,1:2)
hist(sample.jash$beta[1,],prob = TRUE, breaks = seq(-2, 2, length = 20),ylim=c(0,2.5),main='First test')
x = seq(-2, 2, length = 1000)
normdens = dnorm(outer(x,rep(1,length(beta.jash$a.vec))),mean=beta.jash$condpost$normmean[rep(1,1000),],sd=beta.jash$condpost$normsd[rep(1,1000),])
mixnormdens = rowSums(beta.jash$condpost$condpi[rep(1,1000),]*normdens)
lines(x,mixnormdens,type='l',col=2)

hist(sample.jash$beta[2,],prob = TRUE, breaks = seq(-2.5, 2.5, length = 20),ylim=c(0,2.5),main='Second test')
x = seq(-2.5, 2.5, length = 1000)
normdens = dnorm(outer(x,rep(1,length(beta.jash$a.vec))),mean=beta.jash$condpost$normmean[rep(2,1000),],sd=beta.jash$condpost$normsd[rep(2,1000),])
mixnormdens = rowSums(beta.jash$condpost$condpi[rep(2,1000),]*normdens)
lines(x,mixnormdens,type='l',col=2) 
```

Compare the actual FDR and estimated FDR for ash, jash, jash.auto (a way of automatically selecting $a_m$,$b_m$,$c_l$). It seems that jash performs a bit better than ash (which makes sense, since the estimated sebetahat for ash is not accurate).
```{r}
# Compare ash.auto, jash, jash.auto
o.ash = order(beta.ash.auto$qval)
o.jash = order(beta.jash$qval)
o.jash.auto = order(beta.jash.auto$qval)
par(mfrow=c(1,1))
plot(cumsum(ss$null[o.jash])/(1:10000),type='l',ylim=c(0,1),ylab='FDR' )
lines(cumsum(ss$null[o.ash])/(1:10000), col=2)
lines(cumsum(ss$null[o.jash.auto])/(1:10000), col='blue')
lines(sort(beta.jash$qvalue),lty=2)
lines(sort(beta.ash.auto$qvalue),lty=2,col=2)
lines(sort(beta.jash.auto$qvalue),lty=2,col='blue')
title('Dotted line: estimated FDR; Solid line: actual FDR')
legend('topleft',col=c('black','blue','red'),lty=c(1,1,1),legend=c('jash','jash.auto','ash.auto'))
```

The fitted $\pi$ for jash: (sdvec gives the theoretical $SD(y_{jk}|a_m,b_m,c_l)$ for each component)
```{r}
beta.jash$fitted
bata.jash.auto$fitted
```

We can also have a point mass at zero for $\beta$'s prior, by adding components with $c_l=\infty$. Compare the jash results with or without point mass:
```{r}
# Use point mass
beta.jash.pm = jash(ss$Y,usePointMass=TRUE)
#beta.ash.pm = ash(betahat,sebetahat,auto=TRUE,usePointMass=TRUE)
beta.jash$fitted
beta.jash.pm$fitted
o.jash.pm = order(beta.ash.pm$qvalue)
plot(cumsum(ss$null[o.jash])/(1:10000),type='l',ylim=c(0,1),ylab='FDR' )
lines(cumsum(ss$null[o.jash.pm])/(1:10000), col=2)
lines(sort(beta.jash$qvalue),lty=2)
lines(sort(beta.jash.pm$qvalue),lty=2,col=2)
title('Dotted line: estimated FDR; Solid line: actual FDR')
legend('topright',col=c('black','red'),lty=c(1,1),legend=c('jash','jash.pm'))
```

For ash, maybe the ``df'' option can resolve the understimating FDR issue? It works to some extent, but still not very satisfying.
```{r}
# Use df modification for ash
beta.ash.t = ash(betahat,sebetahat,auto=TRUE,df=2)
o.ash.t = order(beta.ash.t$qval)
plot(cumsum(ss$null[o.jash])/(1:10000),type='l',ylim=c(0,1),ylab='FDR' )
lines(cumsum(ss$null[o.ash.t])/(1:10000), col=2)
lines(sort(beta.jash$qvalue),lty=2)
lines(sort(beta.ash.t$qvalue),lty=2,col=2)
title('Dotted line: estimated FDR; Solid line: actual FDR')
legend('topright',col=c('black','red'),lty=c(1,1),legend=c('jash','ash.t'))
```

If the there are many replicates for each test, ash should yield pretty good results? Yes, it is. (But I am trying to figure out why jash performs this bad, maybe it has something to do with the settings of $a_m,b_m,c_l$)
```{r}
ss.largen = simdata(10000, 8000, 0, 3, 1, 20)
beta.largen.jash = jash(ss.largen$Y)
beta.largen.jash.auto = jash(ss.largen$Y,auto=TRUE)

betahat.largen = apply(ss.largen$Y,1,mean)
sebetahat.largen = apply(ss.largen$Y,1,sd)/sqrt(dim(ss.largen$Y)[2])
beta.largen.ash = ash(betahat.largen,sebetahat.largen,auto=TRUE)

o.largen.ash = order(beta.largen.ash$qval)
o.largen.jash = order(beta.largen.jash$qvalue)
o.largen.jash.auto = order(beta.largen.jash.auto$qvalue)
plot(cumsum(ss$null[o.largen.jash])/(1:10000),type='l',ylim=c(0,1),ylab='FDR' )
lines(cumsum(ss$null[o.largen.ash])/(1:10000), col=2)
lines(cumsum(ss$null[o.largen.jash.auto])/(1:10000), col='blue')
lines(sort(beta.largen.jash$qvalue),lty=2)
lines(sort(beta.largen.ash$qvalue),lty=2,col=2)
lines(sort(beta.largen.jash.auto$qvalue),lty=2,col='blue')
title('Large num of replicates case')
legend('topleft',col=c('black','blue','red'),lty=c(1,1,1),legend=c('jash','jash.auto','ash.auto'))
```

Try some different data: the issue of underestimating FDR is still noticable.
```{r}
truenull = c(rep(0, 1000), rep(1, 9000))
beta = c(rnorm(1000, -3, 1), rep(0, 9000))
s = rep(1, 10000)
Y = matrix(rnorm(30000, rep(beta,3), rep(s,3)),ncol=3)
Y.betahat = apply(Y,1,mean)
Y.sebetahat = apply(Y,1,sd)/sqrt(dim(Y)[2])
Y.ash.auto = ash(Y.betahat,Y.sebetahat,auto=TRUE)
Y.jash = jash(Y)
Y.jash.auto = jash(Y,auto=TRUE)

o = order(Y.jash.auto$qvalue)
plot(cumsum(truenull[o])/(1:10000), Y.jash.auto$qvalue[o], col = 2, type = "l",ylim=c(0,0.9))
abline(a = 0, b = 1)
```

Works undone
=============
* Speed up the function!
* Derive full posterior distribution $P(\beta|Y)$
* Better ways to (automatically) select $a_m,b_m,c_l$?
* Issue of underestimating false sign rate
* Use other approaches to estimate $\pi$ (e.g. VB?)
* Try more data (with different scales, underlying distributions, etc)