### 2013-10-01 Rcode/test.mix.Rmd
added test.mix.Rmd; renamed ash.test.R as test.ash.Rmd; updated VBEM back to output post as well as pihat

################################# TESTING #####################################
```{r}
source("mix.R")
temp = normalmix(c(0.5,0.5),c(-3,2),c(1,1))
x = seq(-5,5,length=100)
plot(x,dens(temp,x),type="l")
all.equal(compdens(temp,x),matdnorm(x, temp$mean,temp$sd))

plot(x,mixcdf(temp,x),type="l")

plot(x,compdens_conv(temp,x,0.01)[1,],type="l")
plot(x,compdens_conv(temp,x,0.1)[2,],type="l")
plot(x,dens_conv(temp,x,0.2),type="l")
plot(x,dens_conv(temp,x,0.001),type="l")
plot(x,dens_conv(temp,x,10),type="l")

```

```{r}

temp2 = unimix(c(0.5,0.5),c(-3,3),c(-1,4))
plot_dens(temp2)

plot(x,compdens_conv(temp2,x,0.01)[1,],type="l")
plot(x,compdens_conv(temp2,x,0.1)[2,],type="l")
plot(x,dens_conv(temp2,x,0.2),type="l")

compcdf_post(temp2,0,c(1,2,3),c(10,10,10))
compcdf_post(temp2,-2,c(-2,3),c(0.1,1))

plot_post_cdf(temp2,betahat=c(-2),sebetahat=10)
plot_post_cdf(temp2,betahat=c(-2),sebetahat=1,col=2)
```