### 2013-10-01 Rcode/test.ash.Rmd
added test.mix.Rmd; renamed ash.test.R as test.ash.Rmd; updated VBEM back to output post as well as pihat

```{r}

source("ash.R")
set.seed(100)

#Test VBEM
abf = rbind(c(1,0,0,0),c(0,1,0,0),c(0,0,1,0),c(0,0,1,0),c(0,0,1,0))
eps = 1e-10
abf[abf==0] = eps #replace 0 with small number
print(all.equal(VBEM(abf,c(1,1,1,1))$post,c(2,2,4,1)))
print(all.equal(VBEM(abf,c(1,2,1,1))$post,c(2,3,4,1)))

#simulate n beta-hat values, nnull under the null
#with altmean and altsd being the mean and sd of beta under the alternative
simdata=function(n, nnull, altmean, altsd, betasd){
  null = c(rep(1,nnull),rep(0,n-nnull))
  beta = c(rep(0,nnull),rnorm(n-nnull,altmean,altsd))
  betahat = rnorm(n,beta,betasd)
  return(list(null=null,beta=beta,betahat=betahat,betasd=betasd))
}

ss = simdata(10000,8000,0,2,1)

system.time((beta.ash=ash(ss$betahat,ss$betasd)))
system.time((beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)))
system.time((beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")))
system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)))

hist(ss$beta,prob=TRUE,breaks=seq(-7,7,length=20))
x= seq(-4,4,length=10000)
lines(x,density(beta.ash,x),col=2)
lines(x,density(beta.ash.auto,x),col=3)
lines(x,density(beta.ash.vb.uniform,x),col=4)
lines(x,density(beta.ash.vb.null,x),col=5)
beta.ash$fitted.f
beta.ash.auto$fitted.f
beta.ash.vb.uniform$fitted.f

ss = simdata(10000,10000,0,2,1)
system.time((beta.ash=ash(ss$betahat,ss$betasd)))
system.time((beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)))
system.time((beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")))
system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)))

hist(ss$beta,prob=TRUE,breaks=seq(-7,7,length=20))
x= seq(-4,4,length=10000)
lines(x,density(beta.ash,x),col=2)
lines(x,density(beta.ash.auto,x),col=3)
lines(x,density(beta.ash.vb.uniform,x),col=4)
lines(x,density(beta.ash.vb.null,x),col=5)
beta.ash$fitted.f
beta.ash.auto$fitted.f
beta.ash.vb.uniform$fitted.f
beta.ash.vb.null$fitted.f

beta.ash.vb.uniform$fit$loglik
beta.ash.vb.uniform$fit$converged

```