### 2013-10-04 Rcode/test.mix.Rmd
improved package installation

```{r}
############## HELPER FUNCTIONS #################################################
#OUTPUT: k by n matrix of normal densities
matdnorm = function (x, mu, sigma, log=FALSE) 
{
  k=length(mu)
  n=length(x)
  d = matrix(rep(x,rep(k,n)),nrow=k)
  return(matrix(dnorm(d, mu, sigma, log),nrow=k))
}
```

################################# TESTING #####################################
```{r}
setwd("~/Documents/git/ash/")
source("Rcode/mix.R")
source("Rcode/ash.R")
source("Rcode/ash.oldfuncs.R")
library("testthat")
temp = normalmix(c(0.5,0.5),c(0,0),c(1,2))

expect_that(comp_mean(temp),equals(c(0,0)))
expect_that(mixmean(temp),equals(0))
expect_that(mixsd(temp),equals(sqrt(2.5)))

comp_postmean(temp,0,1)
comp_postsd(temp,0,1)
postsd(temp,0,1)

set.seed(100)
beta = rnorm(100)
betahatse = abs(rnorm(100))
betahat = rnorm(100, beta,betahatse)

pd= posterior_dist(temp,betahat,betahatse)
mv=normmix.mv(pd)
expect_that(mv$mean,equals(postmean(temp,betahat,betahatse)))
expect_that(sqrt(mv$var),equals(postsd(temp,betahat,betahatse)))

x = seq(-5,5,length=100)
plot(x,dens(temp,x),type="l")
all.equal(compdens(temp,x),matdnorm(x, temp$mean,temp$sd))

plot(x,mixcdf(temp,x),type="l")

plot(x,compdens_conv(temp,x,0.01)[1,],type="l")
plot(x,compdens_conv(temp,x,0.1)[2,],type="l")
plot(x,dens_conv(temp,x,0.2),type="l")
plot(x,dens_conv(temp,x,0.001),type="l")
plot(x,dens_conv(temp,x,10),type="l")

```

```{r}
temp2 = unimix(c(0.5,0.5),c(-3,3),c(-1,4))
plot_dens(temp2)

plot(x,compdens_conv(temp2,x,0.01)[1,],type="l")
plot(x,compdens_conv(temp2,x,0.1)[2,],type="l")
plot(x,dens_conv(temp2,x,0.2),type="l")

compcdf_post(temp2,0,c(1,2,3),c(10,10,10))
compcdf_post(temp2,-2,c(-2,3),c(0.1,1))

plot_post_cdf(temp2,betahat=c(-2),sebetahat=10)
plot_post_cdf(temp2,betahat=c(-2),sebetahat=1,col=2)
```

```{r}
etruncnorm(10,20,0,1)
my_etruncnorm(10,20,0,1)

etruncnorm(-20,-10,0,1)
my_etruncnorm(-20,-10,0,1)

etruncnorm(-20,-10,5,1)
my_etruncnorm(-20,-10,5,1)

etruncnorm(-30,-20,5,2)
my_etruncnorm(-30,-20,5,2)

etruncnorm(-Inf,-38,0,1)
my_etruncnorm(-Inf,-38,0,1)

```

```{r}
  comp_postmean(temp2, c(1,2,3), c(10,10,10))
  comp_postmean(temp2, c(1,2,3), c(0.1,0.1,0.1))
```