### 2013-10-24 paper/Rcode/NoisyIllustration.rmd
start draft of paper

The following R code simulates some data where the effects are normally distributed, and the first 500 observations have good precision, and the next 500 have poor precision.

```{r}
#install q value package
#source("http://bioconductor.org/biocLite.R")
#biocLite("qvalue")
setwd("~/Documents/git/ash/Rcode/")
library("qvalue")
library("lattice") #library for some of the plots

#set up some data with mixture of values of s
set.seed(100)
s.good = 0.5
s.poor = 10
J.good= 500
J.poor = 500
J=J.good+J.poor
beta = c(rnorm(J,0,1))
s = c(rep(s.good,J.good),rep(s.poor,J.poor))
betahat = beta + rnorm(J,0,s)
#compute the usual zscore and corresponding p value
zscore = betahat/s
pval = pchisq(zscore^2,df=1,lower.tail=F)
```
As expected the $p$ values from the poor observations are approximately uniform, whereas those from the good observations  are enriched for small $p$ values:

```{r}
p= data.frame(pval=pval,type=c(rep("GOOD",500),rep("POOR",500)))
histogram(~pval|type,p,layout=c(1,2),xlim=c(0,1),breaks=seq(from=0,to=1,length=20),type="count")
```

Now what happens if we apply FDR methods to the all the data is
that the uniform $p$ values from the poor observations
add noise relative to looking at the good observations only. This impacts both the estimate of $pi_0$ from qvalue, and the number of findings that are significant at a given FDR.
```{r}
qq.all = qvalue(p$pval)
qq.good = qvalue(p$pval[p$type=="GOOD"])
print(c(qq.good$pi0, qq.all$pi0))
print(c(sum(qq.good$qvalues<0.05),sum(qq.all$qvalues<0.05)))
```

In contrast, if you use adaptive shrinkage, it makes no difference whether you use all the observations or the just the good ones -
the noise is ignored (as it should be!)

```{r}
library("ashr")
beta.ash.all = ash(betahat,s)
beta.ash.good = ash(betahat[p$type=="GOOD"],s[p$type=="GOOD"])
plot(beta.ash.good$PosteriorMean,beta.ash.all$PosteriorMean[p$type=="GOOD"])
plot(beta.ash.good$PositiveProb,beta.ash.all$PositiveProb[p$type=="GOOD"])
```