### 2013-11-01 Rcode/EmpiricalNull.rmd
added work on talk for NS lab meeting

```{r}
  require(ashr)
  hh = read.table("../data/nejm_brca_release_edit.csv",sep=",",skip=3)
  subset = apply(hh, 1, max)<20
  hh = hh[subset,]

  labs = read.table("../data/nejm_brca_release_edit.csv",sep=",",skip=1,nrows=1)
  labs = 1*(labs=="BRCA1") + 2 * (labs=="BRCA2")

  hh.betahat = apply(hh[,labs==1],1,mean) - apply(hh[,labs==2],1,mean)
  n1 = sum(labs==1)
  n2 = sum(labs==2)
  hh.sebetahat = sqrt(apply(hh[,labs==1],1,var)/n1 + apply(hh[,labs==2],1,var)/n2)
  hh.zscore = hh.betahat/hh.sebetahat
  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
  hist(hh.pval,probability=TRUE,xlab="p value",main="Distribution of p values for Hedenfalk et al data",nclass=40,col=5)
  hh.q = qvalue(hh.pval)
  abline(h=hh.q$pi0,col=2,lwd=2)
  sum(hh.q$q<0.05)

```

Here lambda1>1,lambda2>0 allows for an "inflation factor": betahat is N(beta,lambda1*sebetahat+lambda2) or in other words, betahat is the sum of beta
and a "confounding" beta, which is distributed as N(0,(lambda1-1) sebetahat +lambda2)

```{r}
  hh.ash1.0 = ash(hh.betahat,hh.sebetahat,lambda1=1)
  hh.ash1.1 = ash(hh.betahat,hh.sebetahat,lambda1=1.1)
  hh.ash1.2 = ash(hh.betahat,hh.sebetahat,lambda1=1.2)
  hh.ash1.5 = ash(hh.betahat,hh.sebetahat,lambda1=1.5)
  hh.ash2.0 = ash(hh.betahat,hh.sebetahat,lambda1=2.0)
  cat(get_loglik(hh.ash2.0),get_loglik(hh.ash1.5),get_loglik(hh.ash1.2),get_loglik(hh.ash1.1),get_loglik(hh.ash1.0))

  plot(hh.ash1.0,-5,5)
  plot(hh.ash1.1,-5,5)
  plot(hh.ash1.5,-5,5)
  plot(hh.betahat/hh.sebetahat,hh.ash2$PositiveProb)
```

Now try applying empirical null to z scores, expecting that lambda1 won't be
identifiable. However, there will be a maximum value it can take.
```{r}
  hh.ashz1.0 = ash(hh.betahat/hh.sebetahat,1,lambda1=1)
  hh.ashz1.1 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.1)
  hh.ashz1.2 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.2)
  hh.ashz1.5 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.5)
  hh.ashz1.6 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.6)
  hh.ashz1.7 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.7)
  hh.ashz1.8 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.8)
  hh.ashz1.9 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.9)

  hh.ashz2.0 = ash(hh.betahat/hh.sebetahat,1,lambda1=2.0)
  cat(get_loglik(hh.ashz2.0),get_loglik(hh.ashz1.9),get_loglik(hh.ashz1.8),get_loglik(hh.ashz1.7),get_loglik(hh.ashz1.6),get_loglik(hh.ashz1.5),get_loglik(hh.ashz1.2),get_loglik(hh.ashz1.1),get_loglik(hh.ashz1.0))

  plot(hh.betahat/hh.sebetahat,hh.ashz1.6$PositiveProb)
  points(hh.betahat/hh.sebetahat,hh.ashz1.5$PositiveProb,col=2)
```