### 2013-11-01 talks/NSmeet2013.Rmd

% Shrinkage, False Discovery Rates, and an Alternative to the Zero Assumption
% Matthew Stephens
% 2013/11/1

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(cache=TRUE,autodep=TRUE)
dep_auto()
library("qvalue")
library("ashr")
```

# The Canonical Genomics Experiment

- Measure lots of things, with error

- Get estimates of effects $\beta_j$ ($\hat\beta_j$) and their standard errors $s_j$

- Turn these into Z-scores, $z_j = \hat\beta_j/s_j$

- Turn these into $p$ values, $p_j$

- Apply `qvalue` 
to identify findings ``significant" at a given False Discovery Rate.

- ...?

# Example: BRCA1 vs BRCA2 expression

```{r, echo=FALSE}
  #Set up Analysis of Hedenfalk BRCA data
  setwd("~/Documents/git/ash/paper/Rcode")

  altcol="cyan" #colors to use
  nullcol="blue" 
  nc=40 #number of bins in histograms
  ncz = 100 # number of bins in z score histograms

  hh = read.table("../../data/nejm_brca_release_edit.csv",sep=",",skip=3)
  subset = apply(hh, 1, max)<20
  hh = hh[subset,]

  labs = read.table("../../data/nejm_brca_release_edit.csv",sep=",",skip=1,nrows=1)
  labs = 1*(labs=="BRCA1") + 2 * (labs=="BRCA2")

  hh.betahat = apply(hh[,labs==1],1,mean) - apply(hh[,labs==2],1,mean)
  n1 = sum(labs==1)
  n2 = sum(labs==2)
  hh.sebetahat = sqrt(apply(hh[,labs==1],1,var)/n1 + apply(hh[,labs==2],1,var)/n2)
  hh.zscore = hh.betahat/hh.sebetahat
  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)

```

```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

```

# Example: BRCA1 vs BRCA2 expression

```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  hh.q = qvalue(hh.pval)
  abline(h=hh.q$pi0,col=nullcol,lwd=2)

  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
  #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)

```

# Example: BRCA1 vs BRCA2 expression

```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  hh.q = qvalue(hh.pval)
  abline(h=hh.q$pi0,col=nullcol,lwd=2)

  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
  #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)

  abline(v=0.1,lwd=2,col=2)

  text(0.05,1.2,labels="A",col=2,cex=1.2)  
  text(0.05,0.4,labels="B",col=2,cex=1.2)  
  text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(hh.q$pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
```

# Example 2: Mouse Heart Data

```{r, echo=FALSE}
setwd("~/Documents/git/ash/talks/")
## load Poisson_binomial and ash functions 
source("../../stat45800/code/Rcode/PoissonBinomial.funcs.R")  
#source("../Rcode/ash.R")

x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
xx = rowSums(x[,2:5])
x = x[xx>0,]
xx = xx[xx>0]
```

```{r echo=FALSE}
head(x)
```

- Data on 150 mouse hearts, dissected into left and right ventricle
(courtesy Scott Schmemo, Marcelo Nobrega)

```{r, echo=FALSE}

cc = x[, 2:5]

g = c(-1, -1, 1, 1)
ngene = dim(x)[1]

cc.assoc = counts.associate(cc, g, 1)
zdat.ash = cc.assoc$zdat.ash
zdat = cc.assoc$zdat

# two-sided test
ttest.pval = function(t, df) {
    pval = pt(t, df = df, lower.tail = T)
    pval = 2 * ifelse(pval < 0.5, pval, 1 - pval)
    return(pval)
}

tscore = zdat[3, ]/zdat[4, ]
pval = ttest.pval(tscore, df = 2)
qval = qvalue(pval)

highxp = xx>1000 # select high expressed genes
pval.high = pval[highxp]
qval.high = qvalue(pval.high)
cc.assoc.high = counts.associate(cc[highxp,],g,1)
zdat.ash.high = cc.assoc.high$zdat.ash
zdat.high = cc.assoc.high$zdat

```

# Example 2: Mouse Heart Data

```{r, echo=FALSE,fig.height=4,fig.cap=""}
h=hist(pval,prob=TRUE,ylim=c(0,4),main="p values, all genes",xlab="p value",breaks=seq(0,1,length=21),col=altcol)
abline(h=qval$pi0,col=nullcol,lwd=2)
h$density=rep(qval$pi0,length(h$density))  
plot(h,col=nullcol,add=TRUE,freq=FALSE)
abline(v=0.05,lwd=3)
text(0.6,3,labels=paste0("FDR = ",round(max(qval$qval[pval<0.05]),2)),cex=1.2)
```

# FDR problem 1: different genes have different precision/power

```{r, echo=FALSE,fig.height=4,fig.cap=""}
hist(log10(xx),main="Counts vary considerably across genes", xlab="log10(counts)")
```

# FDR problem 1: lower count genes, less power, add noise

```{r, echo=FALSE,fig.height=4,fig.cap=""}
h=hist(pval[xx<1000],prob=TRUE,ylim=c(0,4),main="p values, low count genes",xlab="p value",breaks=seq(0,1,length=21),col=altcol)
qval.low=qvalue(pval[xx<1000])
pval.low=pval[xx<1000]
abline(h=qval.low$pi0,col=nullcol,lwd=2)
h$density=rep(qval.low$pi0,length(h$density))  
plot(h,col=nullcol,add=TRUE,freq=FALSE)
abline(v=0.05,lwd=3)
#text(0.6,3,labels=paste0("FDR = ",round(max(qval.low$qval[pval.low<0.05]),2)),cex=1.2)
```

# FDR problem 1: higher count genes, more power
```{r, echo=FALSE, fig.height=4,fig.cap=""}

h=hist(pval.high,prob=TRUE,ylim=c(0,4),main="p values, high count genes",xlab="p value",col=altcol)

abline(h=qval.high$pi0,col=nullcol,lwd=2)
h$density=rep(qval.high$pi0,length(h$density))  
plot(h,col=nullcol,add=TRUE,freq=FALSE)
abline(v=0.05,lwd=3)
text(0.6,3,labels=paste0("FDR = ",round(max(qval.high$qval[pval.high<0.05]),2)),cex=1.2)
#qval.high$q[pval.high<0.05] gives FDR for p<0.05
```

# FDR problem 1: q values increased by low count genes
```{r,echo=FALSE, fig.height=4,fig.cap=""}
plot(qval$qval[xx>1000],qval.high$qval,xlab="q values from all gene analysis", ylab="q values from high-count gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
abline(a=0,b=1,col=2)
```

# Problem 2: The Zero Assumption (ZA)

- The standard `qvalue` 
approach assumes that all the $p$ values near 1 are null.

- Analogously, one can assume that all Z scores near 0 are null. Efron refers to this as the ``Zero Assumption".

- This allows us to estimate $\pi_0$ ``conservatively" using the density of $p$ values near 1.

# Problem 2: The ZA

- The ZA seems initially natural.

- However, it turns out to imply unrealistic assumptions about the distribution of non-zero effects.

```{r, echo=FALSE, include=FALSE}
require(fdrtool)
  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
require(locfdr)
  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
require(mixfdr)
  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
```

# Implied distribution of $p$ values under $H_1$

```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  hh.q = qvalue(hh.pval)
  abline(h=hh.q$pi0,col=nullcol,lwd=2)

  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
  #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)

  abline(v=0.1,lwd=2,col=2)

  text(0.05,1.2,labels="A",col=2,cex=1.2)  
  text(0.05,0.4,labels="B",col=2,cex=1.2)  
  text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(hh.q$pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
```

# Implied distribution of Z scores under alternative (fdrtool)

```{r, echo=FALSE,fig.height=4,fig.cap=""}
#plot a histogram of z scores, highlighting the alternative distribution
#of z scores that is implied by localfdr values lfdr.
  nullalthist = function(z,lfdr,...){
    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
    h$density = (1-avlfdr) * h$density
    plot(h,add=TRUE,col=altcol,freq=FALSE)
  }

  nullalthist(hh.zscore,hh.fdrtool$lfdr,main="fdrtool")  
```

# Implied distribution of Z scores under alternative (locfdr)

```{r, echo=FALSE,fig.height=4,fig.cap=""}

  nullalthist(hh.zscore,hh.locfdr$fdr,main="locfdr")
```

# Implied distribution of Z scores under alternative (mixfdr)

```{r, echo=FALSE,fig.height=4,fig.cap=""}

  nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
  dev.off()
```

# Problems: Summary

- By summarizing each observation by a $Z$ score or $p$ value, 
standard fdr tools ignore precision of different measurements

- Standard tools make the ZA, which implies actual effects have a (probably unrealistic) bimodal distribution. [and tends to overestimate $\pi_0$, losing power]

- Also standard tools focus only on zero vs non-zero effects. (eg what if we would
like to identify genes that have at least a 2-fold change?)

# Proposed Alternative

- Instead of working with $z$ scores or $p$ values, work with two numbers 
$(\hat\beta_j,s_j)$ for each test.

- *Incorporate the precision* of the observations $\hat\beta$ into the likelihood.
Specifically, we approximate likelihood for $\beta_j$ by a normal (``Laplace") approximation: 
$$L(\beta_j) \propto \exp(-0.5 (\beta_j - \hat\beta_j)^2/s_j^2).$$
[From $\hat\beta_j \sim N(\beta_j, s_j)$]

- *Directly model the underlying distribution of $\beta$*, using a unimodal family of distributions $g$; estimate $g$ from the data.

# Proposed Alternative: More details

- A convenient way to model $g$ is by a mixture of 0-centered
normal distributions: 
$$g(\beta; \pi) = \sum_{k=1}^K \pi_k N(\beta; 0, \sigma^2_k)$$

- The estimating $g$ comes down to estimating $\pi$ - easily done
by maximum likelihood (EM algorithm) or variational Bayes.

- By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
we get a fairly flexible unimodal symmetric distribution.

- Alternatively, a mixture of uniforms, with 0 as one end-point of the range,
provides still more flexibility, and in particular allows for asymmetry. (Grenander 1953 shows this provides the non-parametric mle in the ``no error" case; Campy + Thomas do the equal-errors case)

# Illustration: $g$ a mixture of 0-centered normals

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, dnorm(x,0,1),type="l",ylim=c(0,2),ylab="density")
lines(x, dnorm(x,0,0.1))
lines(x, dnorm(x,0,0.5))
```

# Illustration: $g$ a mixture of 0-centered normals

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, 0.5*dnorm(x,0,1)+0.5*dnorm(x,0,0.1),type="l",ylim=c(0,2),ylab="density")
```

# Illustration: $g$ a mixture of 0-anchored uniforms

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, dunif(x,0,1),type="s",ylim=c(0,5),ylab="density")
lines(x, dunif(x,0,0.2),type="s")
lines(x, dunif(x,0,0.5),type="s")
lines(x, dunif(x,-0.3,0),type="s",col=2)
lines(x, dunif(x,-0.4,0),type="s",col=2)
lines(x, dunif(x,-2,0),type="s",col=2)
```

# Illustration: $g$ a mixture of 0-anchored uniforms

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, 0.1*dunif(x,0,3)+ 0.3*dunif(x,0,0.2) + 0.2*dunif(x,0,0.5) + 0.1* dunif(x,-0.3,0) + 0.1*dunif(x,-0.3,0)+0.2*dunif(x,-0.4,0),type="s",ylim=c(0,2),ylab="density")
```

# Illustration: BRCA data

```{r}
hh.ash = ash(hh.betahat,hh.sebetahat)
hh.ash.hu = ash(hh.betahat, hh.sebetahat,mixcompdist="halfuniform")
```

```{r, echo=FALSE,fig.height=4,fig.cap=""}
h=hist(hh.betahat,freq=FALSE,main="BRCA, effect sizes",ylim=c(0,2),nclass=nc)
x = seq(-3,3,length=100)
lines(x,density(hh.ash,x),col=2,lwd=2)
lines(x,density(hh.ash.hu,x),col=3,lwd=2)
```

# Illustration: BRCA data
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.ashz$ZeroProb,main="ash")
```

# Illustration: BRCA data
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
```

# Issue: identifiability of $\pi_0$

- For estimating False Discoveries, we are asking whether $\beta_j = 0$.

- However, the data cannot distinguish between $\beta_j = 0$ and $\beta_j$ "very
small"

- Similarly, the data can't tell us what proportion of g should be exactly on zero, vs near zero

- Of course this is true regardless of the method used!

# Issue: identifiability of $\pi_0$

- The Zero assumption (ZA) solves the identifiability problem by assuming that
there *are* no $\beta_j$ near zero!

- The ZA makes $\pi_0$ identifiable.

- Another view is that the estimate of $\pi_0$ under ZA will systematically tend to overestimate $\pi_0$, and so is ``conservative".

- That is it provides an ``upper bound" on $\pi_0$

# Identifiability of $\pi_0$: Solution 1

- We replaced the ZA with the unimodal assumption on g.

- This does not make $\pi_0$ identifiable, but it does effectively provide an upper bound on $\pi_0$.

- Indeed, we saw that when we estimated $\pi_0$ under the ZA the data
then contradicted the unimodal assumption on g. Thus the upper bound is
more conservative than under ZA.

- In practice, implement upper bound by putting prior on $\pi_0$ that encourages it to be big, then estimate $\pi$ by posterior mean (VB).

# Identifiability of $\pi_0$: Solution 2

- Identifiability of $\pi_0$ is primarily an issue if we insist on asking question is $\beta_j=0$?

- How about we change focus: assume *none* of the $\beta_j$ are zero ("one group approach"), and ask for which $\beta_j$ are we confident about the sign (Gelman et al, 2012).

- Positive and negative effects are often treated differently in practice anyway.

- That is we replace fdr with False Sign Rate (fsr), the probability that if we say an effect is positive (negative), it is not.

- Example: suppose we estimate that $\Pr(\beta_j<0)=0.975$ and $\Pr(\beta_j>0)=0.025$. Then we report $\beta_j$ as a ``(negative) discovery", and estimate its fsr as 0.025.

# A technicality

- Suppose you estimate $\Pr(\beta_j<0)=0.98$,  $\Pr(\beta_j>0)=0.01$, $\Pr(\beta_j=0) = 0.01$.

- Should you declare an fdr of 0.01 or 0.02?

- Maybe fsr makes more sense anyway?

# Unification of one-group and two-group solutions

- A possible concern with one-group model: if we assume all $\beta$ are non-zero, but really some are actually zero, might we not underestimate the fsr?

- Consider our example, with $\Pr(\beta_j>0)=0.025$. If we actually allowed
$\beta_j=0$ then possibly all of this probability might actually land at $\beta_j=0$.

- I argue that, assuming symmetry of $g$ near 0, this also provides an upper
bound of how much of the $\Pr(\beta_j<0)=0.975$ might also move to 0.

- Therefore a more conservative estimate of the fsr might be 0.05 (or, more generally, double what you get allowing for point mass)

# Estimation

- This approach also provides a full posterior distribution for each $\beta_j$.

- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg genes with at least 2-fold difference between conditions

# Example: BRCA data

Compare fit of $g$ both allowing $\beta_j=0$ and not allowing it.
```{r, echo=FALSE, includ=FALSE}
  hh.ash.fdr = ash(hh.betahat,hh.sebetahat,usePointMass = TRUE,prior="nullbiased",VB=TRUE)
  hh.ash.shrink = ash(hh.betahat,hh.sebetahat)
```

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot(x,cdf.ash(hh.ash.fdr,x),col=2,type="l",lwd=3,ylab="cdf")
  lines(x,cdf.ash(hh.ash.shrink,x),col=3,lwd=3,lty=2)
```

# Example: BRCA data

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot(hh.ash.fdr$localfsr,2*hh.ash.shrink$localfsr,xlab="fsr with point mass", ylab="2 * fsr without point mass", xlim=c(0,0.1),ylim=c(0,0.1))
  abline(a=0,b=1,col=2,lwd=2)
```

- Because $\pi$ is estimated from the data, the amount
of shrinkage is adaptive to the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each gene.

# Example: ASH applied to mouse data

```{r, fig.height=4,echo=FALSE,fig.cap=""}
hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
```

# Example: ASH applied to mouse data

```{r, fig.height=4,echo=FALSE,fig.cap=""}
hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
t=seq(-2,2,length=201)
lines(t,dnorm(t,0,sd=0.032), col=2, lwd=2)
```

# Example: ASH applied to mouse data

```{r, fig.height=4,echo=FALSE,fig.cap=""}
hist(zdat.ash$PosteriorMean,main="Shrunken estimates",xlab="shrunk betahat",xlim=c(-0.1,0.1),prob=T)
lines(t,dnorm(t,0,sd=0.032), col=2, lwd=2)
```

# Shrinkage is adaptive to information

```{r, echo=FALSE,fig.height=4}
#hist(zdat[3,])
#plot(zdat[3,],zdat.ash$PosteriorMean,xlim=c(-0.2,0.2))
#points(zdat[3,16677],zdat.ash$PosteriorMean[16677],col=2,pch=16)
#points(zdat[3,16079],zdat.ash$PosteriorMean[16079],col=2,pch=16)
#x[16677,]
#x[16079,]
```

```{r, echo=FALSE}
#temp = counts.associate(cc,c(-1,1,-1,1),1)
#temp.tscore = temp$zdat[3,]/temp$zdat[4,]
#temp.pval = ttest.pval(temp.tscore,df=2)
#plot(temp$zdat.ash$localfdr,temp.pval)
#identify(temp$zdat.ash$localfdr,temp.pval)
```

```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
```

# Shrinkage is adaptive to information

```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
points(zdat.ash$localfdr[15325],pval[15325],col=2,pch=16)
points(zdat.ash$localfdr[16123],pval[16123],col=2,pch=16)
```

# Shrinkage is adaptive to information

```{r, echo=FALSE}
cbind(x[,1:5],pval,zdat.ash$localfdr)[c(15325,16123),]
```

# Compare q values for high count genes, with and without low count genes

```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
plot(zdat.ash$qvalue[xx>1000],zdat.ash.high$qvalue,xlab="q values from all data",ylab="q values from high only")
abline(a=0,b=1,col=2,lwd=2)
```

# BRCA1: Compare $\pi_0$ estimates
```{r}
  c(hh.fdrtool$param[3],
  hh.locfdr$fp0[1,3],
  hh.mixfdr$pi[1],
  hh.ashz$fitted.g$pi[1])
```

# BRCA1: Compare number significant at fdr<0.05

```{r}
  c(sum(hh.fdrtool$lfdr<0.05),
  sum(hh.locfdr$fdr<0.05),
  sum(hh.mixfdr$fdr<0.05),
  sum(hh.ashz$ZeroProb<0.05))
```

# Summary: FDR via Adapative Shrinkage

- Both provide a rational approach to identifying ``significant" findings.

- Both are generic and modular: once you have the summary data, you can forget where they came from.

- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values) precision of different measurementscan be better accounted for.

- ASH borrows information for estimation, as well as testing.

# Other Applications

- Widely applicable: perhaps anywhere (?) 
where shrinkage is appropriate, requiring only an estimated
effect size and standard error for each object.

- Could also use effect size estimate and $p$ value for each variable, by converting to effect size estimate and (pseudo-) standard error.

- Currently applying it to wavelet shrinkage applications.

# Guarantees?

- ``I think you have some nice ideas. How will you convince
people to use them?" (C Morris)

# Next steps?

- Extend to allow $g(\cdot;\pi)$ to depend on covariates $X$.

- Extend to allow for correlations in the measured $\hat\beta_j$.

# Thanks

- to the several postdoctoral researchers and students
who have worked with me on related topics.

- Especially Mengyin Lu who coded the VB algorithm.

# Reproducible research

- This document is produced with **knitr**, **Rstudio** and **Pandoc**.

- For more details see my `stephens999/ash` repository at `http://www.github.com/stephens999/ash`

- Website: `http://stephenslab.uchicago.edu`

# Pandoc Command used

`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  ilike-slides.md -o ilike-slides.pdf`

(alternative to produce html slides; but figures would need reworking)
`pandoc -s -S -i -t dzslides --mathjax NSmeet2013.md -o NSmeet2013.html`

Here is my session info:

```{r session-info}
print(sessionInfo(), locale=FALSE)
```

# Some odd things in the data

```{r, echo=FALSE,fig.height=3,fig.width=3}
  plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
  x[tail(order(zdat[3,])),]
```