### 2013-12-15 paper/Rcode/ashsim.rmd
updated paper and code

```{r}
  require("ashr")
  ```{r}
opts_knit$set(progress = TRUE, verbose = TRUE,root.dir="~/Documents/git/ash/paper/Rcode")
require(ashr)
require(qvalue)
require(fdrtool)
require(mixfdr)
require(locfdr)
require(ggplot2)
#simulate from mixture of normals
#compare normal, uniform and half-uniform mixtures
#bsd gives standard deviation of beta
ashsim=function(mixmean,mixsd,mixpi,bsd=1,seedval = 100,nsamp=1000,niter=50){  
  set.seed(seedval)  
  beta =list()
  betahatsd=list()
  betahat = list()
  fit.ash.n = list()
  fit.ash.u = list()
  fit.ash.hu = list()
  fit.ash.hu.vb=list()
  fit.ash.true = list()
  fit.mixfdr= list()
  fit.ash.fdr.n = list()

  fit.mixfdr.enull= list()
  fit.mixfdr.J10= list()
  fit.mixfdr.J10P0= list()
  fit.mixfdr.J100= list()
  k = length(mixmean)
  for(i in 1:niter){
    comp = sample(1:k,nsamp,prob=mixpi,replace=TRUE)
    sd = mixsd[comp]
    mean = mixmean[comp]
    beta[[i]] = rnorm(nsamp,mean,sd)
    betahatsd[[i]] = bsd
    betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
    fit.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="shrink")
    fit.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="shrink")
    fit.ash.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="shrink")
    fit.ash.hu.vb[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform",method="shrink",VB=TRUE)
    fit.ash.true[[i]] = ash(betahat[[i]],betahatsd[[i]],g=normalmix(mixpi,mixmean,mixsd))
    fit.mixfdr[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE)

    fit.ash.fdr.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="fdr")
    #fit.ash.fdr.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="fdr")
    #fit.ash.fdr.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="fdr")
    fit.mixfdr.enull[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=FALSE,plot=FALSE)
    fit.mixfdr.J10[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE,J=10)
     #fit.mixfdr.J100[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE,J=100)
    fit.mixfdr.J10P0[[i]] = try(mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE,J=10,P=0))
  }
  return(list(beta =beta,
  betahatsd=betahatsd,
  betahat = betahat,
  fit.ash.n = fit.ash.n,
  fit.ash.u = fit.ash.u,
  fit.ash.hu = fit.ash.hu,
  fit.ash.hu.vb=fit.ash.hu.vb,
  fit.mixfdr = fit.mixfdr,
  fit.ash.fdr.n = fit.ash.fdr.n,
  fit.ash.true=fit.ash.true,
 # fit.ash.fdr.u = fit.ash.fdr.u,
 # fit.ash.fdr.hu = fit.ash.fdr.hu,
  fit.mixfdr.enull = fit.mixfdr.enull,
  fit.mixfdr.J10 = fit.mixfdr.J10,
  #fit.mixfdr.J100= fit.mixfdr.J100,
  fit.mixfdr.J10P0=fit.mixfdr.J10P0))
}    
```

```{r, cache=TRUE}
sim1= ashsim(c(0,0,0),c(1,1,2),c(1/3,1/3,1/3),niter=20,nsamp=1000)
sim2= ashsim(c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7),rep(1/7,7),niter=20,nsamp=1000)
sim3= ashsim(c(-2,-1,0,1),c(2,1.5,1,1),c(1/4,1/4,1/3,1/6),niter=20,nsamp=1000)

```

Experiment with mixFdr to see how penalization effects fit. 
```{r}
  betahat = sim1$betahat[[1]]
  betahatsd = sim1$betahatsd[[1]]
  set.seed(111) #to ensure reproducibility
  fit.mixfdr.P0 = mixFdr(betahat/betahatsd,noiseSD=1,theonull=TRUE,P=0,plot=FALSE)
  cdf.mixfdr = function(a,x){mixcdf(normalmix(a$pi,a$mu,a$sigma-1),x)}
  x = seq(-4,4,length=100)

  res = data.frame(beta=x,
                  mixfdr.P0 = t(cdf.mixfdr(fit.mixfdr.P0,x)),
                  mixfdr = t(cdf.mixfdr(sim1$fit.mixfdr[[1]],x)),
                  ash.n = t(cdf.ash(sim1$fit.ash.n[[1]],x)$y),
                  ash.fdr.n=t(cdf.ash(sim1$fit.ash.fdr.n[[1]],x)$y))

  truth = t(cdf.ash(sim1$fit.ash.true[[1]],x)$y)

  res.melt=melt(res, id.vars=c("beta"),variable.name="Method")   
  res.melt$Penalization= ifelse((res.melt$Method=="mixfdr" | res.melt$Method=="ash.fdr.n"),"Default","Minimal")
  res.melt$Methodtype=as.character(res.melt$Method)
  res.melt$Methodtype[res.melt$Method=="mixfdr.P0"]="mixfdr"
  res.melt$Methodtype[res.melt$Method=="ash.fdr.n"]="ash"
  res.melt$Methodtype[res.melt$Method=="ash.n"]="ash"

cbbPalette <- c("#56B4E9", "#009E73", "#000000","#F0E442", "#0072B2", "#E69F00",  "#D55E00", "#CC79A7")
pdf("figures/penaltycompare.pdf")
ggplot(res.melt, aes(x=beta)) +
  geom_line(aes(x=beta,y=value,color=Methodtype,linetype=Penalization,group=Method),size=1.5,alpha=0.8) +
  geom_line(data=res, aes(y=truth,color="truth"),alpha=1,size=0.5) +
  scale_colour_manual(name = 'Method', values = cbbPalette)
dev.off()
```

```{r}
plot_examples = function(sims){
  sim1= sims[[1]]
  sim2= sims[[2]]
  sim3= sims[[3]]
len=length(sim1$beta[[1]])
res = data.frame(beta = c(sim1$beta[[1]],sim2$beta[[1]],sim3$beta[[1]]), Scenario = c(rep(1,len),rep(2,len),rep(3,len)))
x = seq(-8,6,length=100)
y = dnorm(x)
dens1 = data.frame(x=x,dtrue=t(density(sim1$fit.ash.true[[1]],x)$y), Scenario=1)
dens2 = data.frame(x=x,dtrue=t(density(sim2$fit.ash.true[[1]],x)$y), Scenario=2)
dens3 = data.frame(x=x,dtrue=t(density(sim3$fit.ash.true[[1]],x)$y), Scenario=3)
dens=rbind(dens1,dens2,dens3)
ggplot(res) + facet_grid(.~Scenario)  + 
    geom_line(data = dens, aes(x,dtrue),size=1.2,alpha=0.9,linetype=1) + scale_x_continuous(limits=c(-6,6))
}

pdf("figures/simABC_egdens.pdf",width=6.5,height=2)
plot_examples(list(sim1,sim2,sim3))
dev.off()
```

```{r}
  rmse = function(x,y){sqrt(mean((x-y)^2))}
  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
  plot_rmse = function(sims,inczero=FALSE,incbetahat=FALSE){
    res=list()

    for(i in 1:length(sims)){
      err.bayes = mapply(get_rmse.ash,sims[[i]]$fit.ash.true,sims[[i]]$beta)
      err.ash.n= mapply(get_rmse.ash,sims[[i]]$fit.ash.n,sims[[i]]$beta)
      err.ash.u = mapply(get_rmse.ash,sims[[i]]$fit.ash.u,sims[[i]]$beta)
      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$fit.ash.hu,sims[[i]]$beta)
      err.mixfdr=mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr,sims[[i]]$beta)
      err.mixfdr.enull = mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr,sims[[i]]$beta)
       err.mixfdr.J10 = mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr.J10,sims[[i]]$beta)
       #err.mixfdr.J100 = mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr.J100,sims[[i]]$beta)
      err.ash.fdr.n = mapply(get_rmse.ash,sims[[i]]$fit.ash.fdr.n,sims[[i]]$beta)

      err.betahat = mapply(rmse,sims[[i]]$betahat,sims[[i]]$beta)
      err.zero = unlist(lapply(sims[[i]]$beta,rmse,y=0))        
      res[[i]] = data.frame(Scenario=i,bayes=err.bayes,ash.normal=err.ash.n,ash.uniform=err.ash.u,ash.halfuniform=err.ash.hu,mixfdr=err.mixfdr,
                            mixfdr.enull = err.mixfdr.enull,
                            mixfdr.J10 = err.mixfdr.J10,
                            #mixfdr.J100 = err.mixfdr.J100,
                            ash.normal.fdr = err.ash.fdr.n)
      if(inczero){
        res[[i]]=data.frame(res[[i]],zero=err.zero)
      }
      if(incbetahat){
        res[[i]]=data.frame(res[[i]],betahat=err.betahat)
      }
    }
    require(reshape2)
    res.melt = melt(res, id.vars=c("Scenario","bayes"),variable.name="Method")    
    p=ggplot(data=res.melt,aes(bayes,value,colour=Method)) +geom_point(shape=16) +
         facet_grid(. ~ Scenario,scale="free_x") +
        geom_abline(colour = "black") +
        xlab("RMSE (Optimal Bayes Rule)") +
        ylab("RMSE (other method)")
    print(p +
          coord_equal(ratio=1))
  }
  pdf("figures/rmse_biplot.pdf",width=6.5,height=2)
    plot_rmse(list(sim1,sim2,sim3))
  dev.off()

#This figure not used in paper?
pdf("figures/rmse_biplot_withzerobetahat.pdf")
    plot_rmse(list(sim1,sim2,sim3),inczero=TRUE,incbetahat=TRUE)
  dev.off()

```

```{r}
  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
  plot_rmse_boxplot = function(sims,inczero=FALSE,incbetahat=FALSE,incmixfdr=FALSE){
    res=list()

    for(i in 1:length(sims)){
      err.bayes = mapply(get_rmse.ash,sims[[i]]$fit.ash.true,sims[[i]]$beta)
      err.ash.n= mapply(get_rmse.ash,sims[[i]]$fit.ash.n,sims[[i]]$beta)
      err.ash.u = mapply(get_rmse.ash,sims[[i]]$fit.ash.u,sims[[i]]$beta)
      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$fit.ash.hu,sims[[i]]$beta)
      err.mixfdr=mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr,sims[[i]]$beta)
      err.mixfdr.enull = mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr,sims[[i]]$beta)
       err.mixfdr.J10 = mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr.J10,sims[[i]]$beta)
#       err.mixfdr.J100 = mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr.J100,sims[[i]]$beta)
      err.ash.fdr.n = mapply(get_rmse.ash,sims[[i]]$fit.ash.fdr.n,sims[[i]]$beta)         
      err.betahat = mapply(rmse,sims[[i]]$betahat,sims[[i]]$beta)
      err.zero = unlist(lapply(sims[[i]]$beta,rmse,y=0))        
      res[[i]] = data.frame(Scenario=i,
                            ash.normal=err.ash.n/err.bayes,
                            ash.uniform=err.ash.u/err.bayes,
                            ash.halfuniform=err.ash.hu/err.bayes,
                            ash.normal.fdr = err.ash.fdr.n/err.bayes
                            )
      if(incmixfdr){
        res[[i]]=data.frame(res[[i]],mixfdr=err.mixfdr/err.bayes,
                            mixfdr.enull = err.mixfdr.enull/err.bayes,
                            mixfdr.J10 = err.mixfdr.J10/err.bayes)
                            #mixfdr.J100 = err.mixfdr.J100/err.bayes)
      }
      if(inczero){
        res[[i]]=data.frame(res[[i]],zero=err.zero/err.bayes)
      }
      if(incbetahat){
        res[[i]]=data.frame(res[[i]],betahat=err.betahat/err.bayes)
      }
    }
    require(reshape2)
    res.melt = melt(res, id.vars=c("Scenario"),variable.name="Method")    
    ggplot(res.melt,aes(Method,value,color=Method)) + geom_boxplot() + facet_grid(.~Scenario)

  }
  pdf("figures/rmse_boxplot.pdf",width=6.5,height=2)
    plot_rmse_boxplot(list(sim1,sim2,sim3))
  dev.off()

    pdf("figures/rmse_boxplot_extended.pdf",width=6.5,height=2)
    plot_rmse_boxplot(list(sim1,sim2,sim3),TRUE,TRUE,TRUE)
  dev.off()

```

```{r}
  plot_loglik_boxplot = function(sims){
    res=list()
    for(i in 1:length(sims)){
      loglik.bayes = mapply(get_loglik,sims[[i]]$fit.ash.true)
      loglik.ash.n= mapply(get_loglik,sims[[i]]$fit.ash.n)
      loglik.ash.u = mapply(get_loglik,sims[[i]]$fit.ash.u)
      loglik.ash.hu = mapply(get_loglik,sims[[i]]$fit.ash.hu)

      loglik.ash.fdr.n = mapply(get_loglik,sims[[i]]$fit.ash.fdr.n)         
      res[[i]] = data.frame(Scenario=i,
                            ash.normal=loglik.ash.n-loglik.bayes,
                            ash.uniform=loglik.ash.u-loglik.bayes,
                            ash.halfuniform=loglik.ash.hu-loglik.bayes,
                            ash.normal.fdr = loglik.ash.fdr.n-loglik.bayes)

    }
    require(reshape2)
    res.melt = melt(res, id.vars=c("Scenario"),variable.name="Method")    
    ggplot(res.melt,aes(Method,value,color=Method)) + geom_boxplot() + facet_grid(.~Scenario)

  }
  pdf("figures/loglik_boxplot.pdf",width=6.5,height=2)
    plot_loglik_boxplot(list(sim1,sim2,sim3))
  dev.off()

```

```{r}
  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
  get_loglik.mixfdr = function(a,betahat,betahatsd){loglik_conv(normalmix(a$pi,a$mu,a$sigma-1),betahat,betahatsd)}
  cdf.mixfdr = function(a,x){mixcdf(normalmix(a$pi,a$mu,a$sigma-1),x)}
  plot_rmse_loglik_boxplot = function(sims){
    res=list()

    for(i in 1:length(sims)){
      err.bayes = mapply(get_rmse.ash,sims[[i]]$fit.ash.true,sims[[i]]$beta)
      err.ash.n= mapply(get_rmse.ash,sims[[i]]$fit.ash.n,sims[[i]]$beta)
      err.ash.u = mapply(get_rmse.ash,sims[[i]]$fit.ash.u,sims[[i]]$beta)
      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$fit.ash.hu,sims[[i]]$beta)
      err.mixfdr=mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr,sims[[i]]$beta)
      err.ash.fdr.n = mapply(get_rmse.ash,sims[[i]]$fit.ash.fdr.n,sims[[i]]$beta)

      res[[i]] = data.frame(Scenario=i,
                            ash.normal=err.ash.n/err.bayes,
                            ash.uniform=err.ash.u/err.bayes,
                            ash.halfuniform=err.ash.hu/err.bayes,
                            ash.normal.fdr = err.ash.fdr.n/err.bayes)

    }

     res2=list()
    for(i in 1:length(sims)){
      loglik.bayes = mapply(get_loglik,sims[[i]]$fit.ash.true)
      loglik.ash.n= mapply(get_loglik,sims[[i]]$fit.ash.n)
      loglik.ash.u = mapply(get_loglik,sims[[i]]$fit.ash.u)
      loglik.ash.hu = mapply(get_loglik,sims[[i]]$fit.ash.hu)
      loglik.mixfdr=mapply(get_loglik.mixfdr,sims[[i]]$fit.mixfdr,sims[[i]]$betahat,sims[[i]]$betahatsd)
      loglik.ash.fdr.n = mapply(get_loglik,sims[[i]]$fit.ash.fdr.n)         
      res2[[i]] = data.frame(Scenario=i,
                            ash.normal=loglik.ash.n-loglik.bayes,
                            ash.uniform=loglik.ash.u-loglik.bayes,
                            ash.halfuniform=loglik.ash.hu-loglik.bayes,
                            ash.normal.fdr = loglik.ash.fdr.n-loglik.bayes)

      }

    require(reshape2)
    res.melt = melt(res, id.vars=c("Scenario"),variable.name="Method")    
    res2.melt = melt(res2, id.vars=c("Scenario"),variable.name="Method")    
    res.melt$type='RMSE (vs Bayes Rule)'
    res2.melt$type='log(likelihood) (vs Bayes Rule)'
  ggplot(rbind(res.melt,res2.melt),aes(Method,value)) + geom_boxplot() + facet_grid(type~Scenario,scale="free_y")

  }
  pdf("figures/rmse_loglik_boxplot.pdf",width=6.5,height=5)
    plot_rmse_loglik_boxplot(list(sim1,sim2,sim3))
  dev.off()

```

looking to see why mixfdr gives worse rmse, even though loglik looks ok.
Perhaps it overshrinks?
```{r}
 plot(sim3$fit.ash.hu[[1]]$PosteriorMean, sim3$fit.mixfdr[[1]]$effectSize)
 plot(sim3$fit.ash.fdr.n[[1]]$PosteriorMean, sim3$fit.mixfdr[[1]]$effectSize)
```

problem seems to be that it overestimates pi0
```{r}
 sim3$fit.mixfdr[[1]]$pi
```
[1] 0.71660654 0.08298064 0.20041282

```{r}
temp=mixFdr(sim3$betahat[[1]]/sim3$betahatsd[[1]],noiseSD=1,theonull=TRUE,calibrate=TRUE)
temp2=mixFdr(sim3$betahat[[1]]/sim3$betahatsd[[1]],noiseSD=1,theonull=TRUE,J=10,plot=FALSE)
plot(ecdf(sim3$beta[[1]]),xlim=c(-6,6))
lines(x,cdf.mixfdr(temp,x),col=3)
lines(cdf.ash(sim3$fit.ash.n[[1]],x),col=2)
lines(cdf.ash(sim3$fit.ash.hu[[1]],x),col=2,lty=2)
```

```{r}
  plot_LR=function(sims){
    hist(unlist(lapply(sims$fit.ash.u,get_loglik))-unlist(lapply(sims$fit.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
  }

  pdf("figures/logLR.pdf")
    plot_LR(sim1)
    plot_LR(sim2)
  dev.off()
```

## Unused figures?

```{r}
plot_examples_withfit = function(sims){
  sim1= sims[[1]]
  sim2= sims[[2]]
  sim3= sims[[3]]
len=length(sim1$beta[[1]])
res = data.frame(beta = c(sim1$beta[[1]],sim2$beta[[1]],sim3$beta[[1]]), Scenario = c(rep(1,len),rep(2,len),rep(3,len)))
x = seq(-8,6,length=100)
y = dnorm(x)
dens1 = data.frame(x=x,dn= t(density(sim1$fit.ash.n[[1]],x)$y),du=t(density(sim1$fit.ash.u[[1]],x)$y),dhu=t(density(sim1$fit.ash.hu[[1]],x)$y), Scenario=1)
dens2 = data.frame(x=x,dn = t(density(sim2$fit.ash.n[[1]],x)$y),du=t(density(sim2$fit.ash.u[[1]],x)$y), dhu=t(density(sim2$fit.ash.hu[[1]],x)$y), Scenario=2)
dens3 = data.frame(x=x,dn = t(density(sim3$fit.ash.n[[1]],x)$y),du=t(density(sim3$fit.ash.u[[1]],x)$y), dhu=t(density(sim3$fit.ash.hu[[1]],x)$y), Scenario=3)
dens=rbind(dens1,dens2,dens3)
cbbPalette <- c("#56B4E9", "#009E73", "#F0E442", "#0072B2", "#E69F00", "#000000", "#D55E00", "#CC79A7")
ggplot(res, aes(x=beta)) + facet_grid(.~Scenario)  +
    geom_histogram(aes(y=..density..),binwidth=0.5,
                   colour="black", fill="white") +  
  geom_line(data = dens, aes(x,du,color='Uniform'),size=1.2,alpha=0.9,linetype=1) +
  geom_line(data = dens, aes(x,dhu,color='Half Uniform'),size=1.2,alpha=0.9,linetype=1) +
    geom_line(data = dens, aes(x,dn,color='Normal'),size=1.2,alpha=0.9,linetype=1) +
  scale_colour_manual(name = 'Method', values = cbbPalette) + scale_x_continuous(limits=c(-4,4))
}

pdf("figures/simABC_eg_withfit.pdf",width=6.5,height=2)
plot_examples_withfit(list(sim1,sim2,sim3))
dev.off()
```

```{r}
cdf.mixfdr = function(a,x){mixcdf(normalmix(a$pi,a$mu,a$sigma-1),x)}
plot_examples_cdf_withfit = function(sims){
  sim1= sims[[1]]
  sim2= sims[[2]]
  sim3= sims[[3]]
len=length(sim1$beta[[1]])
res = data.frame(beta = c(sim1$beta[[1]],sim2$beta[[1]],sim3$beta[[1]]), Scenario = c(rep(1,len),rep(2,len),rep(3,len)))
x = seq(-6,6,length=100)
y = dnorm(x)
cdf1 = data.frame(x=x,dn=t(cdf.ash(sim1$fit.ash.n[[1]],x)$y),du=t(cdf.ash(sim1$fit.ash.u[[1]],x)$y),dhu=t(cdf.ash(sim1$fit.ash.hu[[1]],x)$y), dtrue=t(cdf.ash(sim1$fit.ash.true[[1]],x)$y), dn.fdr=t(cdf.ash(sim1$fit.ash.fdr.n[[1]],x)$y),dm = t(cdf.mixfdr(sim1$fit.mixfdr[[1]],x)), Scenario=1)
cdf2 = data.frame(x=x,dn=t(cdf.ash(sim2$fit.ash.n[[1]],x)$y),du=t(cdf.ash(sim2$fit.ash.u[[1]],x)$y),dhu=t(cdf.ash(sim2$fit.ash.hu[[1]],x)$y), dtrue=t(cdf.ash(sim2$fit.ash.true[[1]],x)$y),dn.fdr=t(cdf.ash(sim2$fit.ash.fdr.n[[1]],x)$y),dm = t(cdf.mixfdr(sim2$fit.mixfdr[[1]],x)),Scenario=2)
  cdf3 = data.frame(x=x,dn=t(cdf.ash(sim3$fit.ash.n[[1]],x)$y),du=t(cdf.ash(sim3$fit.ash.u[[1]],x)$y),dhu=t(cdf.ash(sim3$fit.ash.hu[[1]],x)$y), dtrue=t(cdf.ash(sim3$fit.ash.true[[1]],x)$y),dn.fdr=t(cdf.ash(sim3$fit.ash.fdr.n[[1]],x)$y),dm = t(cdf.mixfdr(sim3$fit.mixfdr[[1]],x)),Scenario=3)

cdf=rbind(cdf1,cdf2,cdf3)
cbbPalette <- c("#56B4E9", "#009E73","#F0E442",  "#000000","#0072B2", "#E69F00", "#000000", "#D55E00", "#CC79A7")
ggplot(res, aes(x=beta)) + facet_grid(.~Scenario)  +
  geom_line(data = cdf, aes(x,dtrue,color='True'),size=1.5,alpha=0.9,linetype=1) +
  geom_line(data = cdf, aes(x,du,color='Uniform'),size=1,alpha=0.9,linetype=2) +
  geom_line(data = cdf, aes(x,dhu,color='Half Uniform'),size=1,alpha=0.9,linetype=2) +
    geom_line(data = cdf, aes(x,dn,color='Normal'),size=1,alpha=0.9,linetype=2) +
    geom_line(data = cdf, aes(x,dn.fdr,color='Normal, null-biased'),size=1,alpha=0.9,linetype=2) +
    geom_line(data = cdf, aes(x,dm,color='mixfdr'),size=1,alpha=0.9,linetype=2) +
  scale_colour_manual(name = 'Method', values = cbbPalette) + scale_x_continuous(limits=c(-6,6))
}

pdf("figures/simABC_eg_cdf_withfit.pdf",width=6.5,height=2)
plot_examples_cdf_withfit(list(sim1,sim2,sim3))
dev.off()
```