### 2014-02-24 paper/Rcode/UCtest.Rmd
added uc talk

% False Discovery Rates, A New Deal
% Matthew Stephens
% 2014/2/24

```{r, include=FALSE}
require("qvalue")
require("ashr")
require("ggplot2")
```

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE)
dep_auto()
```

# Before we get started: Getting Organized

- Over ~10 years of working with graduate students + postdocs,
I've noticed something.
- Organized researchers get more done (and better!).
- Many of them are more organized than I am!
- Thought: I should get organized; I should help others get organized.