### 2014-02-26 paper/Rcode/ashsim.rmd
rearranged simulations for paper some

```{r}
opts_knit$set(progress = TRUE, verbose = TRUE,root.dir="~/Documents/git/ash/paper/Rcode")
require(ashr)
require(qvalue)
require(fdrtool)
require(mixfdr)
require(locfdr)
require(ggplot2)   
```

```{r}
  load("ashsim_out.RData")
  source("plot_examples.R")
  source("plot_rmse.R")
```

```{r}
pdf("figures/simABC_egdens.pdf",width=6.5,height=2)
plot_examples(list(simA,simB,simC))
dev.off()
```

```{r}
  pdf("figures/rmse_biplot.pdf",width=6.5,height=2)
    plot_rmse(list(simA,simB,simC))
  dev.off()

#This figure not used in paper?
pdf("figures/rmse_biplot_withzerobetahat.pdf")
    plot_rmse(list(simA,simB,simC),inczero=TRUE,incbetahat=TRUE)
dev.off()

```

```{r}

  pdf("figures/rmse_boxplot.pdf",width=6.5,height=2)
    plot_rmse_boxplot(list(simA,simB,simC))
  dev.off()

    pdf("figures/rmse_boxplot_extended.pdf",width=6.5,height=2)
    plot_rmse_boxplot(list(simA,simB,simC),TRUE,TRUE,TRUE)
  dev.off()

```

```{r}

  pdf("figures/loglik_boxplot.pdf",width=6.5,height=2)
    plot_loglik_boxplot(list(simA,simB,simC))
  dev.off()

```

```{r}

  pdf("figures/rmse_loglik_boxplot.pdf",width=6.5,height=5)
    plot_rmse_loglik_boxplot(list(simA,simB,simC))
  dev.off()

```

looking to see why mixfdr gives worse rmse, even though loglik looks ok.
Perhaps it overshrinks?
```{r}
 plot(simC$fit.ash.hu[[1]]$PosteriorMean, simC$fit.mixfdr[[1]]$effectSize)
 plot(simC$fit.ash.fdr.n[[1]]$PosteriorMean, simC$fit.mixfdr[[1]]$effectSize)
```

problem seems to be that it overestimates pi0
```{r}
 simC$fit.mixfdr[[1]]$pi
```
[1] 0.71660654 0.08298064 0.20041282

```{r}
temp=mixFdr(simC$betahat[[1]]/simC$betahatsd[[1]],noiseSD=1,theonull=TRUE,calibrate=TRUE)
temp2=mixFdr(simC$betahat[[1]]/simC$betahatsd[[1]],noiseSD=1,theonull=TRUE,J=10,plot=FALSE)
plot(ecdf(simC$beta[[1]]),xlim=c(-6,6))
lines(x,cdf.mixfdr(temp,x),col=3)
lines(cdf.ash(simC$fit.ash.n[[1]],x),col=2)
lines(cdf.ash(simC$fit.ash.hu[[1]],x),col=2,lty=2)
```

```{r}
  plot_LR=function(sims){
    hist(unlist(lapply(sims$fit.ash.u,get_loglik))-unlist(lapply(sims$fit.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
  }

  pdf("figures/logLR.pdf")
    plot_LR(simA)
    plot_LR(simB)
  dev.off()
```

## Unused figures?

```{r}

pdf("figures/simABC_eg_withfit.pdf",width=6.5,height=2)
plot_examples_withfit(list(simA,simB,simC))
dev.off()
```

```{r}
pdf("figures/simABC_eg_cdf_withfit.pdf",width=6.5,height=2)
plot_examples_cdf_withfit(list(simA,simB,simC))
dev.off()
```

Experiment with mixFdr to see how penalization effects fit. 
```{r}
  betahat = simA$betahat[[1]]
  betahatsd = simA$betahatsd[[1]]
  set.seed(111) #to ensure reproducibility
  fit.mixfdr.P0 = mixFdr(betahat/betahatsd,noiseSD=1,theonull=TRUE,P=0,plot=FALSE)
  cdf.mixfdr = function(a,x){mixcdf(normalmix(a$pi,a$mu,a$sigma-1),x)}
  x = seq(-4,4,length=100)

  res = data.frame(beta=x,
                  mixfdr.P0 = t(cdf.mixfdr(fit.mixfdr.P0,x)),
                  mixfdr = t(cdf.mixfdr(simA$fit.mixfdr[[1]],x)),
                  ash.n = t(cdf.ash(simA$fit.ash.n[[1]],x)$y),
                  ash.fdr.n=t(cdf.ash(simA$fit.ash.fdr.n[[1]],x)$y))

  truth = t(cdf.ash(simA$fit.ash.true[[1]],x)$y)

  res.melt=melt(res, id.vars=c("beta"),variable.name="Method")   
  res.melt$Penalization= ifelse((res.melt$Method=="mixfdr" | res.melt$Method=="ash.fdr.n"),"Default","Minimal")
  res.melt$Methodtype=as.character(res.melt$Method)
  res.melt$Methodtype[res.melt$Method=="mixfdr.P0"]="mixfdr"
  res.melt$Methodtype[res.melt$Method=="ash.fdr.n"]="ash"
  res.melt$Methodtype[res.melt$Method=="ash.n"]="ash"

cbbPalette <- c("#56B4E9", "#009E73", "#000000","#F0E442", "#0072B2", "#E69F00",  "#D55E00", "#CC79A7")

pdf("figures/penaltycompare.pdf")
ggplot(res.melt, aes(x=beta)) +
  geom_line(aes(x=beta,y=value,color=Methodtype,linetype=Penalization,group=Method),size=1.5,alpha=0.8) +
  geom_line(data=res, aes(y=truth,color="truth"),alpha=1,size=0.5) +
  scale_colour_manual(name = 'Method', values = cbbPalette)
dev.off()
```