### 2014-02-26 paper/Rcode/fdr.examples.rmd
rearranged simulations; updated makefigs.R

```{r}
opts_knit$set(progress = TRUE, verbose = TRUE,root.dir="~/Documents/git/ash/paper/Rcode")
require(ashr)
require(qvalue)
require(fdrtool)
require(mixfdr)
require(locfdr)
require(ggplot2)
```

```{r}
load("sim1.RData")
```

Show illustrative example
```{r}
altcol="cyan" #colors to use
nullcol="blue" 
nc=40 #number of bins in histograms
ncz = 100 # number of bins in z score histograms

plot_FDR_hist=function(sim,iter=1){
  hh.pval = sim$pval[[iter]]
  hh.zscore = sim$zscore[[iter]]
  hh.hist =hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  hh.q = qvalue(hh.pval)
  abline(h=hh.q$pi0,col=nullcol,lwd=2)

  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)

  abline(v=0.1,lwd=2,col=2)

  text(0.05,1.2,labels="A",col=2,cex=1.2)  
  text(0.05,0.4,labels="B",col=2,cex=1.2)  
  text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(hh.q$pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
}
plot_FDR_hist(simres1,1)

```

```{r}
plot_lfdr_hist=function(sim,iter=1){
  require(fdrtool)
  hh.pval=sim$pval[[iter]]
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  hh.gren = grenander(ecdf(hh.pval))
  abline(h=min(hh.gren$f.knots),col=nullcol,lwd=2)  
  lines(hh.gren$x.knots,hh.gren$f.knots,lwd=2)
  abline(v=0.1,lwd=2,col=2)
  text(0.1,0.9,labels="a",col=2,cex=1)  
  text(0.1,0.34,labels="b",col=2,cex=1.2)  
  text(0.6,3,labels=paste0("lfdr = b/(a+b) =  ",round(min(hh.gren$f.knots)/approx(hh.gren$x.knots,hh.gren$f.knots,0.1)$y,2)),cex=1.2)
}
plot_lfdr_hist(simres1,1)
```

```{r}
#plot a histogram of z scores, highlighting the alternative distribution
#of z scores that is implied by localfdr values lfdr.
  nullalthist = function(z,lfdr,...){
    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
    h$density = (1-avlfdr) * h$density
    plot(h,add=TRUE,col=altcol,freq=FALSE)
  }

#this one puts the null on the bottom
  altnullhist = function(z,lfdr,...){
    h=hist(z, freq=FALSE,col=altcol,nclass=ncz,...)
    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
    h$density = avlfdr * h$density
    plot(h,add=TRUE,col=nullcol,freq=FALSE)
  }

  plotall_hist=function(sim,iter=1,histfun=nullalthist){
    hh.zscore=sim$zscore[[iter]]    
    par(mfcol=c(2,2))
    histfun(hh.zscore,sim$betahat.fdrtool[[iter]]$lfdr,main="fdrtool")  
    histfun(hh.zscore,sim$betahat.locfdr[[iter]]$fdr,main="locfdr")
    histfun(hh.zscore,sim$betahat.mixfdr[[iter]]$fdr,main="mixfdr")
    histfun(hh.zscore,sim$betahat.ash.n[[iter]]$lfdr,main="ash")
    par(mfcol=c(1,1))
  }

 # pdf("figures/nullalthist.pdf")
  plotall_hist(simres1,1,nullalthist)
#  dev.off()

 # pdf("figures/altnullhist.pdf")
  plotall_hist(simres1,1,altnullhist)
#  dev.off()

```