### 2014-03-04 paper/Rcode/conservativepm.rmd

```{r}
opts_knit$set(progress = TRUE, verbose = TRUE,root.dir="~/Documents/git/ash/paper/Rcode")
require(ashr)
require(qvalue)
require(fdrtool)
require(mixfdr)
require(locfdr)
require(ggplot2)
```

```{r}
load("sim1.RData")
```

#look into methods for estimating pi0 conservatively

```{r}
#  par(mfcol=c(3,3))
  plot_ecdf=function(sims){
    for(i in 1:length(sims$beta)){
       plot(ecdf(sims$beta[[i]]),xlim=c(-6,6),main=paste0("iteration ",i))
       x = seq(-6,6,length=1000)
       lines(cdf.ash(sims$betahat.ash.n[[i]],x),col=2,lwd=2)
       lines(cdf.ash(sims$betahat.ash.u[[i]],x),col=3,lwd=2)
       lines(cdf.ash(sims$betahat.ash.true[[i]],x),col=4,lwd=2)
    }
  }
  plot_ecdf(simres1)
  plot_ecdf(simres2)
```

Figure to show that estimated betahats are not so different
```{r}
plot(betahat.ash.u[[1]]$PosteriorMean,betahat.ash.n[[1]]$PosteriorMean)
abline(a=0,b=1,lwd=2,col=2)
```

QUestion: is the null-biased prior maybe a little too conservative?
Answer: log likelihoods don't suggest they are

```{r}
#hh.ashtrue = hh.ashz
#hh.ashtrue$fitted.g$pi = c(2/3,1/15,1/15,1/15,1/15,1/15)
#hh.ashtrue$fitted.g$mean = c(0,0,0,0,0,0)
#hh.ashtrue$fitted.g$sd = sqrt(c(0,1,0.2,0.4,0.8,3))

#loglik(hh.ashtrue,betahat,sebetahat)
#loglik(hh.ashz,betahat,sebetahat)
```