### 2014-03-28 examples/fsr.rmd
added examples

# FSR Example

This is an example of how the false sign rate and the false discovery rate can
be very different when a large proportion of the nulls are false.

In this example, we assume that the true $\beta_j$ ($j=1,\dots,n$) are normally distributed, $\beta_j \sim N(0,1)$. 
So no $\beta_j$ is exactly 0, and all the null hypotheses $H_j: \beta_j=0$ are false.
Since all the nulls are false, no discovery is a false discovery,
and the ``true" FDR is always 0, whatever threshold is used.

```{r}
set.seed(100)
library(ashr)
library(qvalue)
n = 10000
beta = rnorm(n)
sebetahat = 1
betahat = beta + rnorm(n,0,sebetahat)

z = betahat/sebetahat
p = pchisq(z^2, df=1,lower.tail=FALSE)
z.ash = ash(z,1,method="fdr")
z.qvalue = qvalue(p)
```

In this example all the nulls are false, so the actual proportion of true nulls, $\pi_0$, is equal to 0. However, the estimated value of $\pi_0$ can be very different from zero. From qvalue we get
an estimated $\pi_0$ of `r z.qvalue$pi0`. This makes sense when you look at the
distribution of p values, and recall that qvalue makes the assumption that all $p$ values near 1 are null (see histogram below). In contrast, ash produces a much smaller estimate of $\pi_0$ 
($\pi0=$ `r get_pi0(z.ash)`), because it allows that many $p$ values near 1 may actually arise
even when $\beta_j \neq 0$. Of course, in this case both qvalue and ash provide "conservative"
estimates of $\pi_0$, but the qvalue estimate is much more conservative.

```{r}
  hist(p,prob=TRUE)
  abline(h=z.qvalue$pi0,lwd=2)
  text(0.95, z.qvalue$pi0+0.06, "pi0 (q value)")
  abline(h=get_pi0(z.ash),col=2,lwd=2)
  text(0.95, get_pi0(z.ash)+0.06,"pi0 (ash)",col=2)
```

Because of its very conservative estimate of $\pi_0$, qvalue also
provides much more conservative estimates of the FDR. For example, at a $p$ value threshold of 0.05, qvalue estimates the FDR to be `r max(z.qvalue$q[p<0.05])` whereas ash estimates the FDR
as `r max(z.ash$qvalue[p<0.05])`. Of course, the true value of the FDR is zero, so both are conservative.

Even though the true FDR is zero at any threshold, it seems, intuitively, that "discoveries" with
a $p$ value near 1 are not very meaningful. One way to formalize this is to consider the ``false sign rate" instead of the false discovery rate. The local false sign rate (lfsr) for each discovery is the probability that the true $\beta_j$ has the opposite sign to the corresponding $z$ score, $z_j$. The tail false sign rate (FSR) at a given threshold is the average of the lfsr values for discoveries that exceed that threshold. In this example, discoveries with a $p$ value near 1 will have a high lfsr, even though they have a local false discovery rate (lfdr) of 0. When, as here, $\pi_0$ is large, the lfsr is much more meaningful than
the lfdr. When $\pi_0$ is small both measures are meaningful, 
and indeed are generally quite similar. Consequently we recommend lfsr for general usage.

In this example, at a $p$ value threshold of 0.05, $\ash$ estimates the local false sign rate (lfsr) to be `r max(z.ash$lfsr[p<0.05])` and the tail false sign rate (FSR) to be `r mean(z.ash$lfsr[p<0.05])`. The actual FSR at that threshold is `r mean(sign(z[p<0.05])!=sign(beta[p<0.05]))`.