### 2014-05-09 talks/BOG2014.Rmd
added BOG slides

% False Discovery Rates, A New Deal
% Matthew Stephens
% 2014/5/8

```{r, include=FALSE}
require("qvalue")
require("ashr")
require("ggplot2")
```

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE,dev='pdf')
dep_auto()
```

# What you can expect from this talk

- No experimental data!

- No fancy heatmaps!

- 100\% Money-back guarantee to half the False Discovery Rate (FDR) in your data!

# The Canonical Genomics Experiment

- Measure lots of things, with error

- Get estimates of effects $\beta_j$ ($\hat\beta_j$) and their standard errors $s_j$

- Turn these into Z-scores, $z_j = \hat\beta_j/s_j$

- Turn these into $p$ values, $p_j$

- Apply `qvalue` 
to identify findings ``significant" at a given FDR.

- ...?

```{r, echo=FALSE}
  #simple simulated example
  ncz = 100 # number of bins in z score histograms
  set.seed(111)
  hh.betahat = rnorm(1000,0,2)
  hh.sebetahat = 1
  hh.zscore = hh.betahat/hh.sebetahat
  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
  hh.ash = ash(hh.betahat,hh.sebetahat, method="fdr")
  hh.q = qvalue(hh.pval)
```

```{r, echo=FALSE}
  plot_FDReg_hist = function(hh.pval,pi0,nc=40,nullcol="blue",altcol="cyan",type=4,title="Distribution of p values",...){
    hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main=title,nclass=nc,col=altcol,...)
    if(type>1){
      abline(h=pi0,col=nullcol,lwd=2)

      hh.hist$density=rep(pi0,length(hh.hist$density))  
      #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
      plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
    }
    if(type>2){
    abline(v=0.1,lwd=2,col=2)
    }
    if(type>3){
      text(0.05,1.2,labels="A",col=2,cex=1.2)  
      text(0.05,0.4,labels="B",col=2,cex=1.2)  
      text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
    }
  }
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=1)                 
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=2)                 
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=3)                 
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
```

# Problem 1: The Zero Assumption (ZA)

- The standard `qvalue` 
approach assumes that all the $p$ values near 1 are null.

- Analogously, one can assume that all Z scores near 0 are null. Efron refers to this as the ``Zero Assumption".

- Seems initially natural.

```{r, echo=FALSE, include=FALSE}
require(fdrtool)
  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
require(locfdr)
  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
require(mixfdr)
  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
```

# Implied distribution of $p$ values under $H_1$

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
```

# Implied distribution of Z scores under alternative

```{r, echo=FALSE,fig.height=4,fig.cap=""}
#plot a histogram of z scores, highlighting the alternative distribution
#of z scores that is implied by localfdr values lfdr.
  nullalthist = function(z,lfdr,nullcol="blue",altcol="cyan",...){
    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
    h$density = (1-avlfdr) * h$density
    plot(h,add=TRUE,col=altcol,freq=FALSE)
  }

  nullalthist(hh.zscore,hh.fdrtool$lfdr)  
```

# FDR problem 2: different measurement precision

- In some cases the measurement precisions differ among units

- Eg effect sizes of rare SNPs have larger standard error than those of common SNPs

- Eg Expression levels of low-expressed genes have less precision than high-expressed genes

- If some effects are measured less precisely than others, those tests ``lack power"
and dilute signal, increasing FDR

```{r, echo=FALSE}
#install q value package
#source("http://bioconductor.org/biocLite.R")
#biocLite("qvalue")
library("qvalue")
library("lattice") #library for some of the plots

#set up some data with mixture of values of s
set.seed(100)
s.good = 0.5
s.poor = 10
J.good= 500
J.poor = 500
J=J.good+J.poor
beta = c(rnorm(J,0,1))
s = c(rep(s.good,J.good),rep(s.poor,J.poor))
betahat = beta + rnorm(J,0,s)
#compute the usual zscore and corresponding p value
zscore = betahat/s
pval = pchisq(zscore^2,df=1,lower.tail=F)
```

# FDR problem 2: different measurement precision

- Simulation: effects $\beta_j \sim N(0,1)$
- 500 "good" observations with low standard error ($s_j=0.5$)
- 500 "poor" observations with very high standard error ($s_j=10$)

# FDR problem 2: different measurement precision

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_FDReg_hist(pval[1:500],0,title="distribution of GOOD p values",type=1,ylim=c(0,15))
```

# FDR problem 2: different measurement precision

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_FDReg_hist(pval[501:1000],0,title="distribution of POOR p values",type=1,ylim=c(0,15))
```

# FDR problem 2: different measurement precision

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_FDReg_hist(pval[1:1000],0,title="distribution of ALL p values",type=1,ylim=c(0,15))
```
```{r, include=FALSE}
qq.all = qvalue(pval)
qq.good = qvalue(pval[1:500])
```

# FDR problem 2: different measurement precision

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_FDReg_hist(pval[1:500],qq.good$pi0,title="GOOD p values",type=4,ylim=c(0,15))
```

# FDR problem 2: different measurement precision

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_FDReg_hist(pval[1:1000],qq.all$pi0,title="ALL p values",type=4,ylim=c(0,15))
```

# Problems: Summary

Standard tools are highly conservative.

- The ZA, which implies actual effects have a (probably unrealistic) bimodal distribution; causes overestimate of $\pi_0$, losing power.

- By focussing on $p$ values, low-precision measurements can dilute high-precision measurements.

# FDR via Empirical Bayes

- Following previous work (e.g. Newton, Efron, Muralidharan) we take an empirical Bayes approach to FDR.

- Eg Efron assumes that the $Z$ scores come from a mixture of null, and alternative:
$$Z_j \sim f_Z(.) = \pi_0 N(.;0,1) + (1-\pi_0) f_1(.)$$
where $f_1, \pi_0$ are to be estimated from the data.

- Once $f_1$ and $\pi_0$ estimated, FDR calculations are straightforward.

# FDR: A New Deal

- Instead of modelling $Z$ scores, model the effects $\beta$,
$$\beta_j \sim \pi_0 \delta_0(.) + (1-\pi_0) g(.)$$

- Constrain $g$ to be unimodal about 0; estimate $g$ from data.

- *Incorporate precision* of each observation $\hat\beta$ into the likelihood.
Specifically, approximate likelihood for $\beta_j$ by assuming 
$$\hat\beta_j \sim N(\beta_j, s_j)$$

# FDR - A New Deal

- A convenient way to model $g$: mixture of 0-centered
normal distributions: 
$$g(\beta; \pi) = \sum_{k=1}^K \pi_k N(\beta; 0, \sigma^2_k)$$

- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm).

- By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
we get a flexible unimodal symmetric distribution.

- Can approximate, arbitrarily closely, any scale mixture of normals.
Includes almost all priors used for sparse regression problems (spike-and-slab, double exponential/Laplace/Bayesian Lasso, horseshoe).

# FDR - A New Deal

- Alternatively, a mixture of uniforms, with 0 as one end-point of the range,
provides still more flexibility, and in particular allows for asymmetry.

- If allow a very large number of uniforms this provides the non-parametric mle for $g$; cf Grenander 1953; Cordy + Thomas 1997.

# Illustration: $g$ a mixture of 0-centered normals

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, dnorm(x,0,1),type="l",ylim=c(0,2),ylab="density")
lines(x, dnorm(x,0,0.1))
lines(x, dnorm(x,0,0.5))
```

# Illustration: $g$ a mixture of 0-centered normals

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, 0.5*dnorm(x,0,1)+0.5*dnorm(x,0,0.1),type="l",ylim=c(0,2),ylab="density")
```

# Illustration: $g$ a mixture of 0-anchored uniforms

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, dunif(x,0,1),type="s",ylim=c(0,5),ylab="density")
lines(x, dunif(x,0,0.2),type="s")
lines(x, dunif(x,0,0.5),type="s")
lines(x, dunif(x,-0.3,0),type="s",col=2)
lines(x, dunif(x,-0.4,0),type="s",col=2)
lines(x, dunif(x,-2,0),type="s",col=2)
```

# Illustration: $g$ a mixture of 0-anchored uniforms

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, 0.1*dunif(x,0,3)+ 0.3*dunif(x,0,0.2) + 0.2*dunif(x,0,0.5) + 0.1* dunif(x,-0.3,0) + 0.1*dunif(x,-0.3,0)+0.2*dunif(x,-0.4,0),type="s",ylim=c(0,2),ylab="density")
```

# Recall Problem 1: distribution of alternative Z values multimodal
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.fdrtool$lfdr,main="qvalue")  
```

# Problem Fixed: distribution of alternative Z values unimodal
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.ash$lfdr,main="ash")
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4,title="qvalue")                 
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,get_pi0(hh.ash),type=4,title="ash")                 
```

```{r, include=FALSE}
  beta.ash.good = ash(betahat[1:500],s[1:500],method="fdr")
  beta.ash.all = ash(betahat,s,method="fdr")
  print(c(sum(beta.ash.good$qvalue<0.05),sum(beta.ash.all$qvalue<0.05)))
```

# Recall Problem 2: poor measurements dilute good measurments

Number of findings at (estimated) FDR=0.05:
```{r, echo=FALSE}
  res.tab = data.frame(qvalue = c(sum(qq.all$qvalues<0.05),sum(qq.good$qvalues<0.05)),
                       ash =c(sum(beta.ash.all$qvalue<0.05),sum(beta.ash.good$qvalue<0.05)))
  row.names(res.tab)=c("ALL","GOOD")
  kable(res.tab,row.names=TRUE,format="pandoc")
```

# A new problem: an embarrassment of riches

- If the null is mostly false, the new approach can  provide unsettling results

- The FDR can be small for all observations, even those with $p \approx 1$!

- In the illustrative example, the maximum $q$ value is `r round(max(hh.ash$qvalue),2)`

# Perhaps we didn't really understand the question?

- Problem arises only if we insist on asking question ``is $\beta_j=0$?"

- Given enough signal, we become convinced that very few of the $\beta_j=0$

- But for some $\beta_j$ we still may have little information about actual value

- Suggests a change of focus: assume *none* of the $\beta_j$ are zero ("one group approach"), and ask for which $\beta_j$ are we confident about the sign (Gelman et al, 2012).

# The False Sign Rate

- Suggestion: replace FDR with local false sign rate (lfsr), the probability that if we say an effect is positive (negative), it is not.

- Example: suppose we estimate that $\Pr(\beta_j<0)=0.975$ and $\Pr(\beta_j>0)=0.025$. Then we report $\beta_j$ as a ``(negative) discovery", and estimate its fsr as 0.025.

# Even with many signals, large $p$ values have high lfsr

```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
  res=data.frame(p=hh.pval, lfsr=hh.ash$lfsr, qvalue = hh.ash$qvalue)
  res.melt = melt(res, id.vars=c("p"),variable.name="Measure")
cbbPalette <- c("#000000", "#D55E00", "#CC79A7")
 labels = c('qvalue','lfsr')
  breaks = c("qvalue","lfsr")

  pp= ggplot(data=res.melt,aes(p,value,color=Measure)) +geom_point(shape=1) +
      geom_abline(colour = "black") +
        xlab("p-value") +
        ylab("lfsr/qvalue")

  print(pp +scale_y_continuous(limits=c(0,0.1)) +
          scale_x_continuous(limits=c(0,0.1))  +
           scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
          coord_equal(ratio=1))

```

# Summary

- ASH provides a generic approach to shrinkage estimation, as well as
false discovery (sign) rates.

- Unimodal assumption for effects reduces conservatism

- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values or $z$ scores) precision of different measurement is better accounted for.

- In high-signal contexts, False Sign Rate is preferable to False Discovery Rate.

# Reproducible Research?

- Principle: when publishing results of computational procedures, we should
publish the code that produced the results.
- "publishing figures or results without the complete software environment could
be compared to a mathematician publishing an announcement of a mathematical theorem without giving the proof" (Buckheit and Donohoe)
- “an article about a computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result.” [Claerbout]

# This talk is reproducible!

- `http://www.github.com/stephens999/ash`

# Why is reproducibility important?

- Not only because people are forgetful, error-prone, or dishonest!
- Reproducing work is also the first step to extending it.
- Helps communications among researchers (eg student + advisor).

# Thanks

- to the developers of **R**, **knitr**, **Rstudio** and **Pandoc**.

- to the several postdoctoral researchers and students
who have worked with me on related topics.

- Including Ester Pantaleo, Scott Powers, Mengyin Lu, Sen Tian, Wei Wang, Zhengrong Xing.

- NHGRI for funding.

- `ashr` package: `http://www.github.com/stephens999/ash`

# Pandoc Command used

`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  slides.md -o slides.pdf`

(alternative to produce html slides; but figures would need reworking)
`pandoc -s -S -i -t dzslides --mathjax slides.md -o slides.html`

Here is my session info:

```{r session-info}
print(sessionInfo(), locale=FALSE)
```

# Issue: you can't actually estimate $\pi_0$!

- Data cannot distinguish between $\beta_j = 0$ and $\beta_j$ "very
small"

- As a result $\pi_0$ cannot be estimated: the data can never rule out $\pi_0=0$.

- But the unimodal constraint bounds how big $\pi_0$ can be.

- Use penalized likelihood to make $\pi_0$ ``as big as possible",
subject to the unimodal constraint.

# Approach remains conservative (if unimodal assumption holds)

```{r,echo=FALSE,include=FALSE}
load("../paper/Rcode/sim1.RData") #load simulation results
source("../paper/Rcode/plot_pi0.R")
```

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_pi0_onlyqval(list(simres1a,simres2))
```

# Next steps?

- Incorporate $t$ likelihood as well as normal.

- Incorporate shrinkage of variances and not just means.

- Extend to allow $g(\cdot;\pi)$ to depend on covariates $X$.

- Extend to allow for correlations in the measured $\hat\beta_j$.

# Is this an important problem?

- The original paper introducing FDR (Benjamini and Hochberg, 1995) has
been cited 21,787 times according to Google Scholar.

- That is three times a day for the last 19 years!

# FDR and q values

Although precise definitions vary, roughly

- The FDR at a threshold $P$ is 
$$\text{FDR}(P)=\Pr(\beta_j = 0 |  p_j<P).$$

- The q value for observation $j$ is $q_j=\text{FDR}(p_j)$.

# Estimation and Shrinkage

- Besides allowing one to estimate fdr and fsr, 
this approach also provides a posterior distribution for each $\beta_j$.

- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg compute $\Pr(\beta_j > 2 | \hat\beta_j)$).

- And use it to obtain point estimates and credible intervals for each $\beta_j$, taking account of information from all the other $\beta_j$.

- Because $f(\beta)$ is unimodal, the point estimates will tend to be ``shrunk" towards the overall mean (0).

- Because $f(\beta)$ is estimated from the data, the amount
of shrinkage is adaptive to the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each gene.

- So we call the approach ``Adaptive Shrinkage" (ASH).