### 2014-05-22 Rcode/jash.rmd
add jash.rmd

```{r setup, include=FALSE}
library(limma)
library(qvalue)
library(ashr)
source('jash.R')
set.seed(54321)
```

Simulations
===========

Simulate data: N genes, 2 conditions, nA/nB samples for each condition.
Observations are normal distributed.
```{r}
simfull=function(N,Nnull,nA,nB,mean,betasd,diffmean,a,lambda,c){
  null = c(rep(1, Nnull), rep(0, N - Nnull))
  truetau=rgamma(N,shape=a,rate=a/lambda)
  truebetaA=rnorm(N,mean=mean,sd=betasd)
  truediff=c(rep(0,Nnull),rnorm(N-Nnull,mean=diffmean,sd=1/sqrt(c*truetau[Nnull+1:N]))) # true effect
  truebetaB=truebetaA+truediff
  YA=matrix(rnorm(N*nA,mean=rep(truebetaA,nA),sd=rep(1/sqrt(truetau),nA)),ncol=nA)
  YB=matrix(rnorm(N*nB,mean=rep(truebetaB,nB),sd=rep(1/sqrt(truetau),nB)),ncol=nB)
  return(list(YA=cbind(YA,YB),null=null,truetau=truetau,truebetaA=truebetaA,
              truebetaB=truebetaB,truediff=truediff))
}
```

Shrinkage
=========

An example of effects/variances shrinkage using jash & limma:

```{r}
# Simulate data
d.test=simfull(N=10000,Nnull=7000,nA=2,nB=2,mean=5,betasd=0.5,diffmean=0,a=10,lambda=1,c=0.25)
fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
```

The histogram of true effects/variances of this simulated dataset:

```{r,echo=FALSE}
hist(d.test$truediff,prob=TRUE,breaks=seq(-10.5,10.5,by=1),xlab=expression(paste('true ',beta[g2])),main=expression(paste('Histogram of true ',beta[g2])))
hist(d.test$truetau,breaks=40,xlab=expression(paste('true ',tau[g])),main=expression(paste('Histogram of true ',tau[g])))
```

Use jash and limma to shrink gene-specific variances:

```{r}
# jash
jash.test=jash(d.test$Y,fac,usePointMass=TRUE)

# limma
fit.test <- lmFit(d.test$Y, design=design)
fit.test <- eBayes(fit.test)
```

Compare the original $\hat{\beta}_g$'s and $1/\hat{\tau_g}$'s with their shrinkage estimates (posterior mean) by jash:

```{r,echo=FALSE}
betahat=fit.test$coefficients[,2]
sebetahat=fit.test$sigma
plot(betahat,jash.test$PosteriorMean,xlim=c(-10,10),ylim=c(-10,10),xlab=expression(paste(hat(beta)[g])),ylab=expression(paste('E(',beta[g],'|Y)')),main='(a)')
abline(0,1,lty=2)
plot(log(sebetahat^2),log(1/jash.test$PosteriorPrec),xlim=c(-6,6),ylim=c(-2,2),xlab=expression(paste('log(1/',hat(tau)[g],')')),ylab=expression(paste('log(1/E(',tau[g],'|Y))')),main='(b)')
abline(0,1,lty=2)
```

Compare the true $\tau_g$'s with the estimated $\tau_g$'s by jash or limma:

```{r, echo=FALSE}
plot(fit.test$s2.post,1/jash.test$PosteriorPrec,xlim=c(0,3),ylim=c(0,3),xlab='limma',ylab='jash',
     main='Estimated gene-specific variances by limma and jash')
abline(0,1,lty=2)
plot(log(1/d.test$truetau),log(1/jash.test$PosteriorPrec),xlim=c(-1,1),ylim=c(-1,1),xlab=expression(paste('-log(true ',tau,')')),ylab=expression(paste('-log(estimated ',tau,')')))
points(log(1/d.test$truetau),log(fit.test$s2.post),col=2)
abline(0,1,lty=2)
legend('topleft',pch=c(1,1),col=c('black','red'),legend=c('jash','limma'))
```

```{r}
# MSE of true tau and estimated tau (jash)
mean((log(1/d.test$truetau)-log(1/jash.test$PosteriorPrec))^2)
# MSE of true tau and estimated tau (limma)
mean((log(1/d.test$truetau)-log(fit.test$s2.post))^2)
```

Estimate FDR
============

Compare the true/estimated FDR by jash and limma for 8 simulation scenarios:

```{r,echo=FALSE}
# For each case, simulate 20 times
nsim=20

# Case 1
jash.e.FDR.1=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.1=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.1=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.1=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(c(1,1,2,2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.1=simfull(N=10000,Nnull=5000,nA=2,nB=2,mean=5,betasd=0.5,diffmean=0,a=10,lambda=1,c=1)
  jash.1=jash(d.1$Y,fac,usePointMass=TRUE)
  fit.1 <- lmFit(d.1$Y, design=design)
  fit.1 <- eBayes(fit.1)
  o.jash = order(jash.1$qvalue)
  o.limma = order(fit.1$p.value[,2])
  jash.FDR.1[,i]=cumsum(d.1$null[o.jash])/(1:10000)
  limma.FDR.1[,i]=cumsum(d.1$null[o.limma])/(1:10000)
  jash.e.FDR.1[,i]=sort(jash.1$qvalue)
  limma.e.FDR.1[,i]=sort(qvalue(topTable(fit.1,coef='fac2',n=10000)$P.Value)$qval)
}

ng=150
plot(apply(jash.FDR.1,1,mean)[1:ng],type='l',ylim=c(0,0.2),ylab='FDR' )
lines(apply(limma.FDR.1,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.1,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.1,1,mean)[1:ng],lty=2,col=2)
title('Case 1')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 2
jash.e.FDR.2=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.2=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.2=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.2=matrix(rep(NA,nsim*10000),ncol=nsim)
ash.FDR.2=matrix(rep(NA,nsim*10000),ncol=nsim)
ash.e.FDR.2=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(c(1,1,2,2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.2=simfull(N=10000,Nnull=5000,nA=2,nB=2,mean=5,betasd=0.5,diffmean=0,a=10,lambda=1,c=0.1)
  jash.2=jash(d.2$Y,fac,usePointMass=TRUE)
  fit.2 <- lmFit(d.2$Y, design=design)
  fit.2 <- eBayes(fit.2)
  o.jash = order(jash.2$qvalue)
  o.limma = order(fit.2$p.value[,2])

  jash.FDR.2[,i]=cumsum(d.2$null[o.jash])/(1:10000)
  limma.FDR.2[,i]=cumsum(d.2$null[o.limma])/(1:10000)
  jash.e.FDR.2[,i]=sort(jash.2$qvalue)
  limma.e.FDR.2[,i]=sort(qvalue(topTable(fit.2,coef='fac2',n=10000)$P.Value)$qval)

  ash.2 = ash(fit.2$coef[,2]/(sqrt(fit.2$s2.post)*fit.2$stdev.unscaled[,2]),1,method='fdr')
  o.ash=order(ash.2$qvalue)
  ash.FDR.2[,i]=cumsum(d.2$null[o.ash])/(1:10000)
  ash.e.FDR.2[,i]=sort(ash.2$qvalue)
}

ng=2000
plot(apply(limma.FDR.2,1,mean)[1:ng],col=2,type='l',ylim=c(0,0.05),ylab='FDR',xlab='Number of genes')
lines(apply(jash.FDR.2,1,mean)[1:ng])
lines(apply(jash.e.FDR.2,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.2,1,mean)[1:ng],lty=2,col=2)
title('Case 2')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 3
jash.e.FDR.3=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.3=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.3=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.3=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.ef.FDR.3=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(c(1,1,1,2,2,2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.3=simfull(N=10000,Nnull=5000,nA=3,nB=3,mean=5,betasd=0.5,diffmean=0,a=10,lambda=1,c=0.25)
  jash.3=jash(d.3$Y,fac,usePointMass=TRUE)
  fit.3 <- lmFit(d.3$Y, design=design)
  fit.3 <- eBayes(fit.3)
  o.jash = order(jash.3$qvalue)
  o.limma = order(fit.3$p.value[,2])

  jash.FDR.3[,i]=cumsum(d.3$null[o.jash])/(1:10000)
  limma.FDR.3[,i]=cumsum(d.3$null[o.limma])/(1:10000)
  jash.e.FDR.3[,i]=sort(jash.3$qvalue)
  limma.e.FDR.3[,i]=sort(qvalue(topTable(fit.3,coef='fac2',n=10000)$P.Value)$qval)
}

ng=2500
plot(apply(limma.FDR.3,1,mean)[1:ng],col=2,type='l',ylim=c(0,0.1),ylab='FDR' )
lines(apply(jash.FDR.3,1,mean)[1:ng])
lines(apply(jash.e.FDR.3,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.3,1,mean)[1:ng],lty=2,col=2)
title('Case 3')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 4
jash.e.FDR.4=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.4=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.4=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.4=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=50))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.4=simfull(N=10000,Nnull=5000,nA=50,nB=50,mean=7,betasd=0.5,diffmean=0,a=1,lambda=1,c=0.25)
  jash.4=jash(d.4$Y,fac,usePointMass=TRUE)
  fit.4 <- lmFit(d.4$Y, design=design)
  fit.4 <- eBayes(fit.4)
  o.jash = order(jash.4$qvalue)
  o.limma = order(fit.4$p.value[,2])

  jash.FDR.4[,i]=cumsum(d.4$null[o.jash])/(1:10000)
  limma.FDR.4[,i]=cumsum(d.4$null[o.limma])/(1:10000)
  jash.e.FDR.4[,i]=sort(jash.4$qvalue)
  limma.e.FDR.4[,i]=sort(qvalue(topTable(fit.4,coef='fac2',n=10000)$P.Value)$qval)
}

ng=10000
plot(apply(jash.FDR.4,1,mean)[1:ng],type='l',ylim=c(0,1),ylab='FDR' )
lines(apply(limma.FDR.4,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.4,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.4,1,mean)[1:ng],lty=2,col=2)
title('Case 4')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 5
jash.e.FDR.5=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.5=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.5=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.5=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.5=simfull(N=10000,Nnull=7000,nA=2,nB=2,mean=7,betasd=1,diffmean=0,a=5,lambda=1,c=0.25)
  jash.5=jash(d.5$Y,fac,usePointMass=TRUE)
  fit.5 <- lmFit(d.5$Y, design=design)
  fit.5 <- eBayes(fit.5)
  o.jash = order(jash.5$qvalue)
  o.limma = order(fit.5$p.value[,2])

  jash.FDR.5[,i]=cumsum(d.5$null[o.jash])/(1:10000)
  limma.FDR.5[,i]=cumsum(d.5$null[o.limma])/(1:10000)
  jash.e.FDR.5[,i]=sort(jash.5$qvalue)
  limma.e.FDR.5[,i]=sort(qvalue(topTable(fit.5,coef='fac2',n=10000)$P.Value)$qval)
}

ng=2000
plot(apply(jash.FDR.5,1,mean)[1:ng],type='l',ylim=c(0,0.5),ylab='FDR' )
lines(apply(limma.FDR.5,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.5,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.5,1,mean)[1:ng],lty=2,col=2)
title('Case 5')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 6
jash.e.FDR.6=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.6=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.6=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.6=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.6=simfull(N=10000,Nnull=9000,nA=2,nB=2,mean=7,betasd=4,diffmean=0,a=10,lambda=1,c=0.1)
  jash.6=jash(d.6$Y,fac,usePointMass=TRUE)
  fit.6 <- lmFit(d.6$Y, design=design)
  fit.6 <- eBayes(fit.6)
  o.jash = order(jash.6$qvalue)
  o.limma = order(fit.6$p.value[,2])

  jash.FDR.6[,i]=cumsum(d.6$null[o.jash])/(1:10000)
  limma.FDR.6[,i]=cumsum(d.6$null[o.limma])/(1:10000)
  jash.e.FDR.6[,i]=sort(jash.6$qvalue)
  limma.e.FDR.6[,i]=sort(qvalue(topTable(fit.6,coef='fac2',n=10000)$P.Value)$qval)
}

ng=10000
plot(apply(jash.FDR.6,1,mean)[1:ng],type='l',ylim=c(0,1),ylab='FDR' )
lines(apply(limma.FDR.6,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.6,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.6,1,mean)[1:ng],lty=2,col=2)
title('Case 6')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 7
jash.e.FDR.7=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.7=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.7=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.7=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.7=simfull(N=10000,Nnull=7000,nA=2,nB=2,mean=7,betasd=2,diffmean=0,a=1,lambda=0.25,c=0.1)
  jash.7=jash(d.7$Y,fac,usePointMass=TRUE)
  fit.7 <- lmFit(d.7$Y, design=design)
  fit.7 <- eBayes(fit.7)
  o.jash = order(jash.7$qvalue)
  o.limma = order(fit.7$p.value[,2])

  jash.FDR.7[,i]=cumsum(d.7$null[o.jash])/(1:10000)
  limma.FDR.7[,i]=cumsum(d.7$null[o.limma])/(1:10000)
  jash.e.FDR.7[,i]=sort(jash.7$qvalue)
  limma.e.FDR.7[,i]=sort(qvalue(topTable(fit.7,coef='fac2',n=10000)$P.Value)$qval)
}

ng=1000
plot(apply(jash.FDR.7,1,mean)[1:ng],type='l',ylim=c(0,0.2),ylab='FDR' )
lines(apply(limma.FDR.7,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.7,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.7,1,mean)[1:ng],lty=2,col=2)
title('Case 7')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 8
jash.e.FDR.8=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.8=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.8=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.8=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.8=simfull(N=10000,Nnull=7000,nA=2,nB=2,mean=7,betasd=2,diffmean=0,a=1,lambda=4,c=0.25)
  jash.8=jash(d.8$Y,fac,usePointMass=TRUE)
  fit.8 <- lmFit(d.8$Y, design=design)
  fit.8 <- eBayes(fit.8)
  o.jash = order(jash.8$qvalue)
  o.limma = order(fit.8$p.value[,2])

  jash.FDR.8[,i]=cumsum(d.8$null[o.jash])/(1:10000)
  limma.FDR.8[,i]=cumsum(d.8$null[o.limma])/(1:10000)
  jash.e.FDR.8[,i]=sort(jash.8$qvalue)
  limma.e.FDR.8[,i]=sort(qvalue(topTable(fit.8,coef='fac2',n=10000)$P.Value)$qval)
}

ng=250
plot(apply(jash.FDR.8,1,mean)[1:ng],type='l',ylim=c(0,0.2),ylab='FDR' )
lines(apply(limma.FDR.8,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.8,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.8,1,mean)[1:ng],lty=2,col=2)
title('Case 8')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))
```

Compare with limma, jash, limma+ash (Case 2 for example):

```{r,echo=FALSE}
ng=2000
plot(apply(limma.FDR.2,1,mean)[1:ng],col=2,type='l',ylim=c(0,0.05),ylab='FDR',xlab='Number of genes')
lines(apply(jash.FDR.2,1,mean)[1:ng])
lines(apply(ash.FDR.2,1,mean)[1:ng],col='blue')
lines(apply(jash.e.FDR.2,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.2,1,mean)[1:ng],lty=2,col=2)
lines(apply(ash.e.FDR.2,1,mean)[1:ng], col='blue',lty=2)
legend('topleft',col=c('red','black','blue','red','black','blue'),lty=c(2,2,2,1,1,1),legend=c('limma','jash','ash+limma','true FDR (jash)','true FDR (limma)','true FDR (ash+limma)'))
```

Robustness
==========

Simulate data from the following 3 alternative models:

Case 1: $\beta_g \sim \pi_0\delta_0+\pi_1N(0, \sigma^2)$, $\tau_g \sim Gamma(a,a/\lambda)$, $\hat{\beta}_g\sim N(\beta_g,1/\tau_g)$.
```{r}
simrob=function(N,Nnull,nA,nB,mean,betasd,diffmean,diffsd,a,lambda){
  null = c(rep(1, Nnull), rep(0, N - Nnull))
  truetau=rgamma(N,shape=a,rate=a/lambda)
  truebetaA=rnorm(N,mean=mean,sd=betasd)
  truediff=c(rep(0,Nnull),rnorm(N-Nnull,mean=diffmean,sd=diffsd))
  truebetaB=truebetaA+truediff
  YA=matrix(rnorm(N*nA,mean=rep(truebetaA,nA),sd=rep(1/sqrt(truetau),nA)),ncol=nA)
  YB=matrix(rnorm(N*nB,mean=rep(truebetaB,nB),sd=rep(1/sqrt(truetau),nB)),ncol=nB)
  return(list(YA=cbind(YA,YB),null=null,truetau=truetau,truebetaA=truebetaA,
              truebetaB=truebetaB,truediff=truediff))
}
```
```{r,echo=FALSE}
# Case 1
jash.e.FDR.rob=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.rob=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.rob=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.rob=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.rob=simrob(N=10000,Nnull=5000,nA=2,nB=2,mean=5,betasd=0.5,diffmean=0,diffsd=1,a=5,lambda=2)
  fac=factor(rep(c(1,2),each=2))
  design=model.matrix(~fac)

  jash.rob=jash(d.rob$Y,fac,usePointMass=TRUE)
  fit.rob <- lmFit(d.rob$Y, design=design)
  fit.rob <- eBayes(fit.rob)
  o.jash = order(jash.rob$qvalue)
  o.limma = order(fit.rob$p.value[,2])

  jash.FDR.rob[,i]=cumsum(d.rob$null[o.jash])/(1:10000)
  limma.FDR.rob[,i]=cumsum(d.rob$null[o.limma])/(1:10000)
  jash.e.FDR.rob[,i]=sort(jash.rob$qvalue)
  limma.e.FDR.rob[,i]=sort(qvalue(topTable(fit.rob,coef='fac2',n=10000)$P.Value)$qval)
}

ng=2000
plot(apply(jash.FDR.rob,1,mean)[1:ng],type='l',ylim=c(0,0.3),ylab='FDR' )
lines(apply(limma.FDR.rob,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.rob,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.rob,1,mean)[1:ng],lty=2,col=2)
title('Case 1')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))
```

Case 2: $\beta_g \sim \pi_0\delta_0+\pi_1N(0, 1/(c\tau_g))$, $\tau_g \sim N(\mu,\sigma^2)$, $\hat{\beta}_g\sim N(\beta_g,1/\tau_g)$.
```{r,echo=FALSE}
simrob2=function(N,Nnull,nA,nB,mean,betasd,diffmean,lambda,tausd,c){
  null = c(rep(1, Nnull), rep(0, N - Nnull))
  truetau=rnorm(N,mean=lambda,sd=tausd)
  truebetaA=rnorm(N,mean=mean,sd=betasd)
  truediff=c(rep(0,Nnull),rnorm(N-Nnull,mean=diffmean,sd=1/sqrt(c*truetau[Nnull+1:N])))
  truebetaB=truebetaA+truediff
  YA=matrix(rnorm(N*nA,mean=rep(truebetaA,nA),sd=rep(1/sqrt(truetau),nA)),ncol=nA)
  YB=matrix(rnorm(N*nB,mean=rep(truebetaB,nB),sd=rep(1/sqrt(truetau),nB)),ncol=nB)
  return(list(YA=cbind(YA,YB),null=null,truetau=truetau,truebetaA=truebetaA,
              truebetaB=truebetaB,truediff=truediff))
}
```
```{r,echo=FALSE}
# Case 2
jash.e.FDR.rob2=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.rob2=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.rob2=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.rob2=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.rob2=simrob2(N=10000,Nnull=5000,nA=2,nB=2,mean=5,betasd=0.5,diffmean=0,lambda=3,tausd=0.25,c=1)
  fac=factor(rep(c(1,2),each=2))
  design=model.matrix(~fac)

  jash.rob2=jash(d.rob2$Y,fac,usePointMass=TRUE)
  fit.rob2 <- lmFit(d.rob2$Y, design=design)
  fit.rob2 <- eBayes(fit.rob2)
  o.jash = order(jash.rob2$qvalue)
  o.limma = order(fit.rob2$p.value[,2])

  jash.FDR.rob2[,i]=cumsum(d.rob2$null[o.jash])/(1:10000)
  limma.FDR.rob2[,i]=cumsum(d.rob2$null[o.limma])/(1:10000)
  jash.e.FDR.rob2[,i]=sort(jash.rob2$qvalue)
  limma.e.FDR.rob2[,i]=sort(qvalue(topTable(fit.rob2,coef='fac2',n=10000)$P.Value)$qval)
  }

ng=1000
plot(apply(jash.FDR.rob2,1,mean)[1:ng],type='l',ylim=c(0,0.6),ylab='FDR' )
lines(apply(limma.FDR.rob2,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.rob2,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.rob2,1,mean)[1:ng],lty=2,col=2)
title('Case 2')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))
```

Case 3: $\beta_g \sim \pi_0\delta_0+\pi_1N(0, 1/(c\tau_g))$, $log(\tau_g) \sim N(\mu,\sigma^2)$, $\hat{\beta}_g\sim N(\beta_g,1/\tau_g)$.
```{r,echo=FALSE}
simrob3=function(N,Nnull,nA,nB,mean,betasd,diffmean,lambda,logtausd,c){
  null = c(rep(1, Nnull), rep(0, N - Nnull))
  truetau=rlnorm(N,mean=log(lambda),sd=logtausd)
  truebetaA=rnorm(N,mean=mean,sd=betasd)
  truediff=c(rep(0,Nnull),rnorm(N-Nnull,mean=diffmean,sd=1/sqrt(c*truetau[Nnull+1:N])))
  truebetaB=truebetaA+truediff
  YA=matrix(rnorm(N*nA,mean=rep(truebetaA,nA),sd=rep(1/sqrt(truetau),nA)),ncol=nA)
  YB=matrix(rnorm(N*nB,mean=rep(truebetaB,nB),sd=rep(1/sqrt(truetau),nB)),ncol=nB)
  return(list(YA=cbind(YA,YB),null=null,truetau=truetau,truebetaA=truebetaA,
              truebetaB=truebetaB,truediff=truediff))
}
```
```{r,echo=FALSE}
# Case 3
jash.e.FDR.rob3=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.rob3=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.rob3=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.rob3=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.rob3=simrob3(N=10000,Nnull=5000,nA=2,nB=2,mean=5,betasd=0.5,diffmean=0,lambda=3,logtausd=1,c=1)
  fac=factor(rep(c(1,2),each=2))
  design=model.matrix(~fac)

  jash.rob3=jash(d.rob3$Y,fac,usePointMass=TRUE)
  fit.rob3 <- lmFit(d.rob3$Y, design=design)
  fit.rob3 <- eBayes(fit.rob3)
  o.jash = order(jash.rob3$qvalue)
  o.limma = order(fit.rob3$p.value[,2])

  jash.FDR.rob3[,i]=cumsum(d.rob3$null[o.jash])/(1:10000)
  limma.FDR.rob3[,i]=cumsum(d.rob3$null[o.limma])/(1:10000)
  jash.e.FDR.rob3[,i]=sort(jash.rob3$qvalue)
  limma.e.FDR.rob3[,i]=sort(qvalue(topTable(fit.rob3,coef='fac2',n=10000)$P.Value)$qval)
}

ng=1000
plot(apply(jash.FDR.rob3,1,mean)[1:ng],type='l',ylim=c(0,0.5),ylab='FDR' )
lines(apply(limma.FDR.rob3,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.rob3,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.rob3,1,mean)[1:ng],lty=2,col=2)
title('Case 3')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))
```

Application
===========

BRCA dataset:
```{r,echo=FALSE}
hh = read.table("nejm_brca_release_edit.csv",sep=",",skip=3)
subset = apply(hh, 1, max)<20
hh = hh[subset,]

labs = read.table("nejm_brca_release_edit.csv",sep=",",skip=1,nrows=1)
labs = 1*(labs=="BRCA1") + 2 * (labs=="BRCA2")

hh.betahat = apply(hh[,labs==1],1,mean) - apply(hh[,labs==2],1,mean)
n1 = sum(labs==1)
n2 = sum(labs==2)
hh.sebetahat = sqrt(apply(hh[,labs==1],1,var)/n1 + apply(hh[,labs==2],1,var)/n2)

hh.jash=jasha(hh.betahat,hh.sebetahat,df=n1+n2-2)

library(limma)
hh.new=cbind(hh[,labs==1],hh[,labs==2])
fac=factor(c(rep(1,n1),rep(2,n2)))
design=model.matrix(~fac)
fit <- lmFit(hh.new, design=design)
fit <- eBayes(fit)
```

```{r}
# Number of genes called significant (jash, FDR<0.05)
sum(hh.jash$qvalue<0.05)

# Number of genes called significant (limma, FDR<0.05)
sum(qvalue(topTable(fit,coef='fac2',n=3169)$P.Val)$qval<0.05)
```