### 2014-05-29 scratchwork/makenew.mixEM.Rmd

The aim here is to use the SQUAREM package to improve the speed of convergence of the EM algorithm.

First here is our existing implementation.

```{r}
mixEM = function(matrix_lik, prior, pi.init = NULL,tol=0.0001, maxiter=5000){
  n=nrow(matrix_lik)
  k=ncol(matrix_lik)
  B = rep(0,maxiter)
  pi = pi.init
  if(is.null(pi.init)){
    pi = rep(1/k,k)# Use as starting point for pi
  } 
  pi = ifelse(pi<1e-5,1e-5,pi) #set any estimates that are too small to be just very small
  pi = normalize(pi)

  loglik = rep(0,maxiter)
  priordens= rep(0,maxiter)
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  loglik[1] = sum(log(m.rowsum))
  priordens[1] = sum((prior-1)*log(pi)) 
  classprob = m/m.rowsum #an n by k matrix
  i=1
  if(maxiter >= 2){
    for(i in 2:maxiter){  
      pi = colSums(classprob) + prior-1
      pi = ifelse(pi<1e-5,1e-5,pi) #set any estimates that are less than zero, which can happen with prior<1, to 0
      pi = normalize(pi)

      #Now re-estimate pi
      m  = t(pi * t(matrix_lik)) 
      m.rowsum = rowSums(m)
      loglik[i] = sum(log(m.rowsum))
      priordens[i] = sum((prior-1)*log(pi)) 
      classprob = m/m.rowsum

      if(abs(loglik[i]+priordens[i]-loglik[i-1]-priordens[i-1])<tol) break;
    }
  }
  converged=(abs(loglik[i]+priordens[i]-loglik[i-1]-priordens[i-1])<tol)
  if(!converged){
      warning("EM algorithm in function mixEM failed to converge. Results may be unreliable. Try increasing maxiter and rerunning.")
  }
  return(list(pihat = pi, B=loglik[1:i], 
              niter = i, converged=converged))
}
```

Now a new implementation based on SQUAREM
```{r}
normalize=function(x){return(x/sum(x))}
library(SQUAREM)

fixpoint = function(pi, matrix_lik, prior){  
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  classprob = m/m.rowsum #an n by k matrix
  pinew = normalize(colSums(classprob) + prior-1)
  return(pinew)
}

penloglik = function(pi, matrix_lik, prior){
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  loglik = sum(log(m.rowsum))
  priordens = sum((prior-1)*log(pi))
  return(loglik+priordens)
}

mixEM2 = function(matrix_lik, prior, pi.init = NULL,tol=1e-7, maxiter=5000){
  if(is.null(pi.init)){
    pi = rep(1/k,k)# Use as starting point for pi
  } 
  res = squarem(par=pi.init,fixptfn=fixpoint, objfn=penloglik,matrix_lik=matrix_lik, prior=prior, control=list(maxiter=maxiter,tol=tol))
  return(list(pihat = res$par, B=res$value.objfn, 
              niter = res$iter, converged=res$convergence))
}

```

Now simulate some data
```{r}
set.seed(100)
  sampsize=100000
  sd = c(1,1.1,1.2)
  z = rnorm(sampsize,0,sample(sd,sampsize,replace=TRUE))
  lik = t(vapply(z,dnorm,sd,sd=sd))
  prior=c(1,1,1)
  pi.init=c(0.3,0.2,0.5)
```

```{r}
  system.time(res1<-mixEM(lik,prior,pi.init))
  system.time(res2<-mixEM2(lik, prior, pi.init))
  options(digits=13)
  res1
  res2
```