### 2014-05-29 tests/largedata.Rmd

# Test ash on large data

```{r}
  library(ashr)
  set.seed(111)
  z = rnorm(100000)
  z.ash = ash(z,1)
  z.ash2 = ash(z,1,method="fdr")
```