### 2014-06-02 scratchwork/makenew.mixEM.Rmd
updated variational em

The aim here is to use the SQUAREM package to improve the speed of convergence of the EM algorithm.

First here is our existing implementation.

```{r}
mixEM = function(matrix_lik, prior, pi.init = NULL,tol=0.0001, maxiter=5000){
  n=nrow(matrix_lik)
  k=ncol(matrix_lik)
  B = rep(0,maxiter)
  pi = pi.init
  if(is.null(pi.init)){
    pi = rep(1/k,k)# Use as starting point for pi
  } 
  pi = ifelse(pi<1e-5,1e-5,pi) #set any estimates that are too small to be just very small
  pi = normalize(pi)

  loglik = rep(0,maxiter)
  priordens= rep(0,maxiter)
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  loglik[1] = sum(log(m.rowsum))
  priordens[1] = sum((prior-1)*log(pi)) 
  classprob = m/m.rowsum #an n by k matrix
  i=1
  if(maxiter >= 2){
    for(i in 2:maxiter){  
      pi = colSums(classprob) + prior-1
      pi = ifelse(pi<1e-5,1e-5,pi) #set any estimates that are less than zero, which can happen with prior<1, to 0
      pi = normalize(pi)

      #Now re-estimate pi
      m  = t(pi * t(matrix_lik)) 
      m.rowsum = rowSums(m)
      loglik[i] = sum(log(m.rowsum))
      priordens[i] = sum((prior-1)*log(pi)) 
      classprob = m/m.rowsum

      if(abs(loglik[i]+priordens[i]-loglik[i-1]-priordens[i-1])<tol) break;
    }
  }
  converged=(abs(loglik[i]+priordens[i]-loglik[i-1]-priordens[i-1])<tol)
  if(!converged){
      warning("EM algorithm in function mixEM failed to converge. Results may be unreliable. Try increasing maxiter and rerunning.")
  }
  return(list(pihat = pi, B=loglik[1:i], 
              niter = i, converged=converged))
}
```

Now a new implementation based on SQUAREM
```{r}
normalize=function(x){return(x/sum(x))}
library(SQUAREM)

fixpoint = function(pi, matrix_lik, prior){  
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  classprob = m/m.rowsum #an n by k matrix
  pinew = normalize(colSums(classprob) + prior-1)
  return(pinew)
}

penloglik = function(pi, matrix_lik, prior){
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  loglik = sum(log(m.rowsum))
  subset = (prior-1) != 0
  priordens = sum((prior-1)[subset]*log(pi[subset]))
  return(loglik+priordens)
}

mixEM2 = function(matrix_lik, prior, pi.init = NULL,tol=1e-7, maxiter=5000){
  if(is.null(pi.init)){
    pi = rep(1/k,k)# Use as starting point for pi
  } 
  res = squarem(par=pi.init,fixptfn=fixpoint, objfn=penloglik,matrix_lik=matrix_lik, prior=prior, control=list(maxiter=maxiter,tol=tol))
  return(list(pihat = res$par, B=res$value.objfn, 
              niter = res$iter, converged=res$convergence))
}

#as above but with initial iterations done with smaller sample sizes
#and tolerance set according to sample size
mixEM3 = function(matrix_lik, prior, pi.init = NULL,tol=1e-7, maxiter=5000){
  if(is.null(pi.init)){
    pi = rep(1/k,k)# Use as starting point for pi
  } 
  n = nrow(matrix_lik)
  if(n>1000){       
    res = squarem(par=pi.init,fixptfn=fixpoint, objfn=penloglik,matrix_lik=matrix_lik[1:1000,], prior=prior, control=list(maxiter=maxiter,tol=1e-3,trace=TRUE))
    pi.init = res$par
    print(pi.init)
  }
  res = squarem(par=pi.init,fixptfn=fixpoint, objfn=penloglik,matrix_lik=matrix_lik, prior=prior, control=list(maxiter=maxiter,tol=tol,trace=TRUE))
  return(list(pihat = res$par, B=res$value.objfn, 
              niter = res$iter, converged=res$convergence))
}
```

Now simulate some data
```{r}
  set.seed(100)
  sampsize=100000
  sd = c(1,1.1,1.2)
  z = rnorm(sampsize,0,sample(sd,sampsize,replace=TRUE))
  lik = t(vapply(z,dnorm,sd,sd=sd))
  prior=c(1,1,1)
  pi.init=c(0.3,0.2,0.5)
```

```{r}
  system.time(res1<-mixEM(lik,prior,pi.init))
  system.time(res2<-mixEM2(lik, prior, pi.init))
   system.time(res3<-mixEM3(lik, prior, pi.init))
  options(digits=13)
  res1
  res2
  res3
```

What we found: subsetting by initially running on only 1000 data points
did not work so well - it converged to a value of pi that was pretty different from
the mle for all 100,000 so didn't really save anything.

This section just to run 100 times to see if any warnings results
```{r}
#   options(warn=2)
#    for(i in 1:100){
#      set.seed(i)
#     z0 = rnorm(sampsize,0,1)
#     sd=c(1,2,4,8,16,32)
#     prior=c(1,1,1,1,1,1)
#     pi.init=c(1,1,1,1,1,1)/6
#     lik = t(vapply(z0,dnorm,sd,sd=sd))
#     print(i)
#     res2[[i]]<-mixEM2(lik, prior, pi.init)
#     }
```