### 2014-06-13 tests/largedata.Rmd
updated package with check for negative pis

# Test ash on large data

```{r}
  require(ashr)
  eg5 = matrix(scan("~/Dropbox/eg.j5.txt"), byrow=TRUE, nrow=2)
  temp1=ash(eg5[1,],eg5[2,],cxx=FALSE)
  #temp2=fast.ash(eg5[1,],eg5[2,],prior="nullbiased",pointmass=TRUE,gridmult=2,cxx=FALSE)

```

```{r}
  eg9 = matrix(scan("~/Dropbox/eg.j9.txt"), byrow=TRUE, nrow=2)
  temp1=ash(eg9[1,],eg9[2,],cxx=FALSE)
  #temp2=fast.ash(eg9[1,],eg9[2,],prior="nullbiased",pointmass=TRUE,gridmult=2,cxx=FALSE)

```

```{r}
  library(ashr)
  set.seed(111)
  z = rnorm(100000)
  z.ash = ash(z,1,cxx=FALSE,method="fdr")
  z.ash2 = ash(z,1,method="fdr")
  z.ash$fitted.g
  z.ash2$fitted.g
```