### 2014-06-13 scratchwork/test.turboem.Rmd
corrected error in sign of objective function calculation for SQUAREM

Test turboEM
========================================================

First we have to define a fixpoint function and an objective function
for turboEM to work with
```{r}
normalize=function(x){return(x/sum(x))}

fixpoint = function(pi, matrix_lik, prior){  
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  classprob = m/m.rowsum #an n by k matrix
  pinew = normalize(colSums(classprob) + prior-1)
  return(pinew)
}

negpenloglik = function(pi, matrix_lik, prior){
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  loglik = sum(log(m.rowsum))
  priordens = sum((prior-1)*log(pi))
  return(-(loglik+priordens))
}

#Note this doesn't impose the constraint that they have to sum to 1... might be a problem?
pconstr <- function(par) {
  return(all(0 <= par & par <= 1))
}
```

Now simulate some test data
```{r}
  set.seed(100)
  sampsize=10000
  sd = c(1,1.1,1.2)
  z = rnorm(sampsize,0,sample(sd,sampsize,replace=TRUE))
  lik = t(vapply(z,dnorm,sd,sd=sd))
  prior=c(1,1,1)
  pi.init=c(0.3,0.2,0.5)
```

```{r}
  library(turboEM)
  res = turboem(par=pi.init, control.run=list(convtype = "objfn",tol=1e-5),fixptfn=fixpoint, objfn=negpenloglik, pconstr=pconstr,method=c("em", "squarem", "pem","decme","qn"), matrix_lik=lik,prior=prior)
  options(digits=13)
  res
  library(devtools)
  devtools::load_all("../package/ashr")
  system.time(res2<-mixEM(lik, prior, pi.init))
  length(res2$B)
  res2$B[length(res2$B)]
```

The number of iterations for the ash implementation is `r length(res2$B)` and the objective achieved
is `r res2$B[length(res2$B)]`.

Now try a bigger sample, just comparing EM and squareEM
```{r}
  set.seed(100)
  sampsize=500000
  sd = c(1,1.1,1.2)
  z = rnorm(sampsize,0,sample(sd,sampsize,replace=TRUE))
  lik = t(vapply(z,dnorm,sd,sd=sd))
  prior=c(1,1,1)
  pi.init=c(0.3,0.2,0.5)
  res = turboem(par=pi.init, control.run=list(convtype = "objfn",tol=1e-5),fixptfn=fixpoint, objfn=negpenloglik, pconstr=pconstr,method=c("em", "squarem"), matrix_lik=lik,prior=prior)
  options(digits=13)
  res
```

Now run the squarem package directly
```{r}
  library(SQUAREM)
  system.time(res3<-squarem(par=pi.init,fixptfn=fixpoint, objfn=negpenloglik,matrix_lik=lik, prior=prior))
```

It seems from this limited assessment that i) the squareEM approach is the most effective,
and ii) the SQUAREM package implementation is faster then the turboem implementation.