### 2014-10-14 paper/Rcode/fdr.examples.rmd
updated paper

```{r}
#opts_knit$set(progress = TRUE, verbose = TRUE,root.dir="~/Documents/git/ash/paper/Rcode")
require(ashr)
require(qvalue)
require(fdrtool)
require(mixfdr)
require(locfdr)
require(ggplot2)
source("nullalthist.R")
source("plot_FDReg_hist.R")
```

```{r}
load("sim1.RData")
```

Show illustrative example
```{r}
altcol="cyan" #colors to use
nullcol="blue" 
nc=40 #number of bins in histograms
ncz = 100 # number of bins in z score histograms

source("plot_FDReg_hist.R")
plot_FDReg_hist(simres1a,1)

```

```{r}
plot_lfdr_hist=function(sim,iter=1){
  require(fdrtool)
  hh.pval=sim$pval[[iter]]
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  hh.gren = grenander(ecdf(hh.pval))
  abline(h=min(hh.gren$f.knots),col=nullcol,lwd=2)  
  lines(hh.gren$x.knots,hh.gren$f.knots,lwd=2)
  abline(v=0.1,lwd=2,col=2)
  text(0.1,0.9,labels="a",col=2,cex=1)  
  text(0.1,0.34,labels="b",col=2,cex=1.2)  
  text(0.6,3,labels=paste0("lfdr = b/(a+b) =  ",round(min(hh.gren$f.knots)/approx(hh.gren$x.knots,hh.gren$f.knots,0.1)$y,2)),cex=1.2)
}
plot_lfdr_hist(simres1a,1)
```

```{r}

 # pdf("figures/nullalthist.pdf")
  plotall_hist(simres1a,1,nullalthist)
#  dev.off()

 # pdf("figures/altnullhist.pdf")
  plotall_hist(simres1a,1,altnullhist)
#  dev.off()

```