### 2015-02-06 dsc-shrink/graphs.Rmd
added shiny graphic for showing boxplot of dsc-shrink results

---
title: "Graphs"
runtime: shiny
output: html_document
---

Here's a boxplot of results. 
```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
load("res.RData")
scenario_names = as.character(unique(res$scenario))
method_names = as.character(unique(res$method))
numeric_criteria = names(res)[unlist(lapply(res,is.numeric))]

inputPanel(    
    # Create the checkboxes and select them all by default
    checkboxGroupInput("scen.subset", "Choose Scenarios", 
                        choices  = scenario_names,
                        selected = scenario_names),

    checkboxGroupInput("method.subset", "Choose Methods", 
                        choices  = method_names,
                        selected = method_names),

    selectInput("criteria", "Choose Criteria", 
                        choices  = numeric_criteria,
                        selected = "RMSE")
)

renderPlot({
  res.filter = filter(res,scenario %in% input$scen.subset & method %in% input$method.subset)
  print(input)
  res.filter$value = res.filter[[input$criteria]]
  ggplot(res.filter,aes(method,value,color=method)) + geom_boxplot() + facet_grid(.~scenario)
})
```