### 2015-02-06 scratchwork/checknull.Rmd

Check performance of ash under null
========================================================

Here we simulate all null data sets and look at the 
estimated values of pi0

```{r}
  library(ashr)
  z.ash=list()
  for(i in 1:100){
    set.seed(i)
    z = rnorm(1000) #c(rnorm(950),rnorm(50,0,3))
    z.ash[[i]] = ash(z,1,cxx=FALSE,method="fdr")
    print(i)
    print(get_pi0(z.ash[[i]]))
  }

  nfp = function(a,t=0.05){return(sum(a$lfsr[1:950]<t))} #find number of false positives at threshold t
  mean(unlist(lapply(z.ash,nfp)))
```