### 2015-02-06 scratchwork/makenew.mixVBEM.Rmd

make new mixVBEM
========================================================

First our existing implementation

```{r}
#return the KL-divergence between 2 dirichlet distributions
#p,q are the vectors of dirichlet parameters of same lengths
diriKL = function(p,q){
  p.sum = sum(p)
  q.sum = sum(q)
  k = length(q)
  KL = lgamma(q.sum)-lgamma(p.sum)+sum((q-p)*(digamma(q)-digamma(rep(q.sum,k))))+sum(lgamma(p)-lgamma(q))
  return(KL)
}

#helper function for VBEM
VB.update = function(matrix_lik, pipost){
  avgpipost = matrix(exp(rep(digamma(pipost),n)-rep(digamma(sum(pipost)),k*n)),ncol=k,byrow=TRUE)
  classprob = avgpipost * matrix_lik
  classprob = classprob/rowSums(classprob) # n by k matrix
  B = sum(classprob*log(avgpipost*matrix_lik),na.rm=TRUE) - diriKL(prior,pipost) - sum(classprob*log(classprob)) #negative free energy
  return(list(classprob=classprob,B=B))
}

mixVBEM = function(matrix_lik, prior, post.init=NULL, tol=0.0001, maxiter=5000){
  n=nrow(matrix_lik)
  k=ncol(matrix_lik)
  B = rep(0,maxiter)
  pipost = post.init
  if(is.null(post.init)){
    pipost = prior # Dirichlet posterior on pi
  }
  avgpipost = matrix(exp(rep(digamma(pipost),n)-rep(digamma(sum(pipost)),k*n)),ncol=k,byrow=TRUE)
  classprob = avgpipost * matrix_lik
  classprob = classprob/rowSums(classprob) # n by k matrix  
  B[1] = sum(classprob*log(avgpipost*matrix_lik),na.rm=TRUE) - diriKL(prior,pipost) - sum(classprob*log(classprob))  #negative free energy
  i=1

  if(maxiter>=2){
    for(i in 2:maxiter){  
      pipost = colSums(classprob) + prior

      #Now re-estimate pipost
      avgpipost = matrix(exp(rep(digamma(pipost),n)-rep(digamma(sum(pipost)),k*n)),ncol=k,byrow=TRUE)
      classprob = avgpipost*matrix_lik
      classprob = classprob/rowSums(classprob) # n by k matrix

      B[i] = sum(classprob*log(avgpipost*matrix_lik),na.rm=TRUE) - diriKL(prior,pipost) - sum(classprob*log(classprob))

      if(abs(B[i]-B[i-1])<tol) break;
    } 
    if(i>maxiter){i=maxiter}
  }

  converged=(abs(B[i]-B[i-1])<tol)
  if(!converged){
      warning("EM algorithm in function mixVBEM failed to converge. Results may be unreliable. Try increasing maxiter and rerunning.")
  }

  return(list(pihat = pipost/sum(pipost), B=B[1:i], niter = i, converged=converged,post=pipost))
}

```

Now the squarem version
```{r}
require(SQUAREM)

VBfixpoint = function(pipost, matrix_lik, prior){  
  n=nrow(matrix_lik)
  k=ncol(matrix_lik)
  avgpipost = matrix(exp(rep(digamma(pipost),n)-rep(digamma(sum(pipost)),k*n)),ncol=k,byrow=TRUE)
  classprob = avgpipost*matrix_lik
  classprob = classprob/rowSums(classprob) # n by k matrix
  pipostnew = colSums(classprob) + prior
  return(pipostnew)
}

VBnegpenloglik = function(pipost, matrix_lik, prior){
  n=nrow(matrix_lik)
  k=ncol(matrix_lik)
  avgpipost = matrix(exp(rep(digamma(pipost),n)-rep(digamma(sum(pipost)),k*n)),ncol=k,byrow=TRUE)
  classprob = avgpipost*matrix_lik
  classprob = classprob/rowSums(classprob) # n by k matrix

  B= sum(classprob*log(avgpipost*matrix_lik),na.rm=TRUE) - diriKL(prior,pipost) - sum(classprob*log(classprob)) 
  return(-B)
}

mixVBEM2 = function(matrix_lik, prior, pi.init = NULL,tol=1e-7, maxiter=5000){
  if(is.null(pi.init)){
    pi = rep(1/k,k)# Use as starting point for pi
  } 
  res = squarem(par=pi.init,fixptfn=VBfixpoint, objfn=VBnegpenloglik,matrix_lik=matrix_lik, prior=prior, control=list(maxiter=maxiter,tol=tol))
  return(list(pihat = res$par, B=res$value.objfn, 
              niter = res$iter, converged=res$convergence))
}
```

Now simulate some data
```{r}
  set.seed(100)
  sampsize=100000
  sd = c(1,1.1,1.2)
  z = rnorm(sampsize,0,sample(sd,sampsize,replace=TRUE))
  lik = t(vapply(z,dnorm,sd,sd=sd))
  prior=c(1,1,1)
  pi.init=c(0.3,0.2,0.5)
```

```{r}
  system.time(res1<-mixVBEM(lik,prior,pi.init))
  system.time(res2<-mixVBEM2(lik,prior,pi.init))
  options(digits=13)
  res1
  res2
```