### 2015-02-06 scratchwork/test.gradient.Rmd

Try testing whether gradient of pi0 is negative at pi0=1 (null)
========================================================

The kth element of this vector is the derivative 
of the loglik for $\pi=(\pi_0,...,1-\pi_0,...)$ with respect to $\pi_0$ at $\pi_0=1$.
```{r}
gradient = function(matrix_lik){
  n = nrow(matrix_lik)
  grad = n - colSums(matrix_lik/matrix_lik[,1]) 
  return(grad)
}
```

Simulate some data
```{r}
 set.seed(100)
  sampsize=10000
  sd = c(1,1.1,1.2)
  z0 = rnorm(sampsize,0,1)
  lik0 = t(vapply(z0,dnorm,sd,sd=sd))
  z1 = rnorm(sampsize,0,sample(sd,sampsize,replace=TRUE))
  lik1 = t(vapply(z1,dnorm,sd,sd=sd))

```

```{r}
  gradient(lik0)
  gradient(lik1)
```

```{r}
  for(i in 1:100){
    set.seed(i)
    z0 = rnorm(sampsize,0,1)
  lik0 = t(vapply(z0,dnorm,sd,sd=sd))
    print(i)
    print(gradient(lik0))
  }
```

```{r}
  set.seed(96)
  z0 = rnorm(sampsize,0,1)
  lik0 = t(vapply(z0,dnorm,sd,sd=sd))
  print(gradient(lik0))
  temp=lik0/lik0[,1]

```