### 2015-02-06 Rcode/testnonzeromode.asymmetric.Rmd
added a bunch of files

---
title: "Test non-zero mode in asymmetric case"
output: html_document
---

This example comes from Efron (2008) p16 when examining the false coverage rate (FCR).
It is clearly asymmetric. However, the strongest mode is at 0 (there's a point mass there). I'm hoping the nonzeromode option picks that up.

```{r}
  set.seed(100)
  nsamp=10000
  mu0 = rep(0,nsamp)
  mu1 = rnorm(nsamp,-3,1)
  comp = rbinom(nsamp,1,0.1)
  mu = ifelse(comp==0,mu0,mu1)
  z = rnorm(nsamp,mu,1)
```

```{r}
  library(ashr)
  res.ash = ash(z,1,mixcompdist="halfuniform",method="fdr",nonzeromode=TRUE)
  res.ash
```