### 2015-02-06 talks/shiny/shinytest.Rmd

---
title: Interactive illustration of unimodal assumption
runtime: shiny
output: ioslides_presentation
---

```{r, include=FALSE}
 #plot a histogram of z scores, highlighting the alternative distribution
#of z scores that is implied by pi0
nullalthist_z = function(z,pi0,nullcol="blue",altcol="cyan",...){
  h=hist(z, freq=FALSE,col=nullcol,nclass=40,ylim=c(-0.5,0.5),...)
  nb = length(h$breaks)
  nullexp = pi0 * (pnorm(h$breaks)[-1] - pnorm(h$breaks[-nb]))/(h$breaks[-1]-h$breaks[-nb])
  h$density = h$density - nullexp
  plot(h,add=TRUE,col=altcol,freq=FALSE)
}

 set.seed(100)
z = c(rnorm(5000,0,2),rnorm(5000,0,1))
```

## Illustration of unimodal assumption
Increase the proportion of nulls until the alternative (cyan) looks unimodal.
```{r, echo=FALSE}

inputPanel(  
  sliderInput("pi0", label = "Proportion of nulls:",
              min = 0, max = 1, value = 0, step = 0.1)
)

renderPlot({
  nullalthist_z(z, input$pi0)
})
```