### 2015-02-20 tests/test_nonparam.Rmd
update example

---
title: "test nonparam"
output: html_document
---

```{r}
#Following tests nonparam estimation
library("ashr")
set.seed(100)
# First, try fixing g to a true g (but not unimodal)
beta = c(rnorm(1000,0,1),rnorm(1000,5,1))
betahatsd = 1
betahat = rnorm(length(beta),mean=beta,sd=betahatsd)
g = normalmix(c(0.5,0.5),c(0,5),c(1,1))
res.1=ash(betahat,betahatsd,g=g,control=list(maxiter=0)) # use the mixture components in g, and also the mixture proportions since don't iterate!
res.2=ash(betahat,betahatsd,g=g) # use the mixture components in g, but iterate!
```

Now try non-parametric version by setting guni to be piecewise uniform
```{r}
K=100
bins=seq(-10,10,length=K+1)
binc = (bins[-1]+bins[-(K+1)])/2 # bincenters
guni = unimix(rep(1,K)/K, bins[-(K+1)],bins[-1])

res.nonp=ash(betahat,betahatsd,g=guni) # use the mixture components in g, but iterate!
res.nonp2=ash(betahat,betahatsd,g=guni,randomstart=TRUE)

plot(binc,get_density(res.1,binc)$y,type="l")
lines(binc,get_density(res.2,binc)$y,col=2)
lines(binc,get_density(res.nonp,binc)$y,type="s",col=3)
lines(binc,get_density(res.nonp2,binc)$y,type="s",col=4)
```

Check likelihoods
```{r}
get_loglik(res.1)
get_loglik(res.2)
get_loglik(res.nonp)
get_loglik(res.nonp2)
```

Now try VB to avoid overfitting?
```{r cache=TRUE}
res.VB=ash(betahat,betahatsd,g=guni,randomstart=TRUE,VB = TRUE,prior=rep(1,K)) 
res.VB2=ash(betahat,betahatsd,g=guni,randomstart=TRUE,VB = TRUE,prior=rep(1/2,K)) 
res.VB3=ash(betahat,betahatsd,g=guni,randomstart=TRUE,VB = TRUE,prior=rep(1/K,K))

plot(binc,get_density(res.1,binc)$y,type="l")
lines(binc,get_density(res.VB,binc)$y,col=2)
lines(binc,get_density(res.VB2,binc)$y,col=3)
lines(binc,get_density(res.VB3,binc)$y,col=4)
```

compare evidence
```{r}
  get_loglik(res.VB)
  get_loglik(res.VB2)
  get_loglik(res.VB3)
```

Try same thing with less noise
```{r}
set.seed(100)
# First, try fixing g to a true g (but not unimodal)
beta = c(rnorm(1000,0,1),rnorm(1000,5,1))
betahatsd = 0.05
betahat = rnorm(length(beta),mean=beta,sd=betahatsd)
g = normalmix(c(0.5,0.5),c(0,5),c(1,1))
res.1=ash(betahat,betahatsd,g=g,control=list(maxiter=0)) # use the mixture components in g, and also the mixture proportions since don't iterate!
res.2=ash(betahat,betahatsd,g=g) # use the mixture components in g, but iterate!
```

Now try non-parametric version by setting guni to be piecewise uniform
```{r}
K=100
bins=seq(-10,10,length=K+1)
binc = (bins[-1]+bins[-(K+1)])/2 # bincenters
guni = unimix(rep(1,K)/K, bins[-(K+1)],bins[-1])

res.nonp=ash(betahat,betahatsd,g=guni) # use the mixture components in g, but iterate!
res.nonp2=ash(betahat,betahatsd,g=guni,randomstart=TRUE)

plot(binc,get_density(res.1,binc)$y,type="l")
lines(binc,get_density(res.2,binc)$y,col=2)
lines(binc,get_density(res.nonp,binc)$y,type="s",col=3)
lines(binc,get_density(res.nonp2,binc)$y,type="s",col=4)
```