### 2015-03-16 Rcode/efron.fcr.rmd
some updates to fcr and paper

---
title: "FSR in non-unimodal case"
output: html_document
---

This example comes from Efron (2008) p16 when examining the false coverage rate (FCR).

```{r}
  set.seed(100)
  nsamp=10000
  altmean=-3
  mu0 = rep(0,nsamp)
  mu1 = rnorm(nsamp,altmean,1)
  comp = rbinom(nsamp,1,0.1)
  mu = ifelse(comp==0,mu0,mu1)
  z = rnorm(nsamp,mu,1)
```

```{r}
  library(ashr)
  res.ash = ash(z,1,mixcompdist="halfuniform")
  CImatrix= ashci(res.ash,level=0.95)
```

Compare results with Bayes rule
```{r}
  BayesComparePlot=function(altmean=-3){
    plot(z,mu,xlim=c(-8,0))
    points(z[CImatrix[,1]],CImatrix[,5],col="red")
    points(z[CImatrix[,1]],CImatrix[,4],col="red")

    fdr = 0.9*dnorm(z)/(0.9*dnorm(z)+0.1*dnorm(z,altmean,sqrt(2)))
    o=order(z)
    upper = ifelse(fdr[o]<0.025,(z[o]+altmean)/2+qnorm(0.975+fdr[o])/sqrt(2),0)
    lines(z[o],upper,col="green",lwd=2)
    lines(z[o],(z[o]+altmean)/2-qnorm(0.975)/sqrt(2),col="green",lwd=2)
    abline(v=max(z[fdr<0.05]))
  }
  BayesComparePlot(altmean)
```

It seems the lower 97.5% point is pretty accurate.

Try whole thing again to see if it was a fluke
```{r}
  set.seed(101)
  nsamp=10000
  mu0 = rep(0,nsamp)
  mu1 = rnorm(nsamp,-3,1)
  comp = rbinom(nsamp,1,0.1)
  mu = ifelse(comp==0,mu0,mu1)
  z = rnorm(nsamp,mu,1)
```

```{r}
   res.ash = ash(z,1,mixcompdist="halfuniform")
  CImatrix= ashci(res.ash,level=0.95)
  BayesComparePlot()
```

And try with normal mixture in ash:

```{r}
   res.ash = ash(z,1,mixcompdist="normal")
  CImatrix= ashci(res.ash,level=0.95)
  BayesComparePlot()
```

And uniform mixture:
```{r}
   res.ash = ash(z,1,mixcompdist="uniform")
  CImatrix= ashci(res.ash,level=0.95)
  BayesComparePlot()
```

Things to try:
- simulation study for coverage?
- compare with EB FCR methods in JRSS B?