### 2015-04-15 dsc-shrink/check_mixfdr_lownoise.rmd
automate table generation

---
title: "check_mixfdr_lownoise"
output: html_document
---

Here we check that mixfdr actually does OK at fitting the bimodal distribution when 
noise is small compared with bimodality.

First simulate bimodal data with well-separated peaks. The histogram
shows the multi-modal nature of the underlying g is clear.
```{r}
  source("datamakers/datamaker.R")
  sim.bimodal=rnormmix_datamaker(
              args=list(g=normalmix(c(0.5,0.5),c(-5,5),c(1,1)),
              min_pi0=0.4,
              max_pi0=0.4,
              nsamp=1000,
              betahatsd=1)
            )
  hist(sim.bimodal$input$betahat,nclass=20,xlim=c(-10,10),prob=TRUE)
```

Now run mixfdr on it. Note that the fit captures the multi-modal nature of g, 
although mixfdr continues to overestimate pi0.
```{r}
  source("methods/mixfdr.wrapper.R")
  sim.bimodal.mixfdroutput = mixfdr.wrapper(sim.bimodal$input,args=list(theonull=TRUE))
  g = mixfdr2fitted.g(sim.bimodal.mixfdroutput)$fitted.g
  x=seq(-10,10,length=100)
  plot(ecdf(sim.bimodal$meta$beta),main="Cdf of true beta: empirical (black) and fitted (red)")
  lines(x,mixcdf(g,x),col=2)
```