### 2015-05-24 talks/ilike2015.rmd
add ilike talk on dsc

---
title: "Dynamic Statistical Comparisons, and other adventures in getting organized"
author: Matthew Stephens
date: May 22, 2015
output: ioslides_presentation
---

```{r, include=FALSE}
require("qvalue")
require("ashr")
require("ggplot2")
require("knitr")
```

```{r setup, include=FALSE}
  #set global chunk options
  opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE)
  dep_auto()
```

## Getting Organized!

- Over ~10 years of working with graduate students + postdocs,
I've noticed something.
- Organized researchers get more done (and better!).
- Many of them are more organized than I am.
- Thought: I should get organized; I should help others get organized.

## So what can you do?

- Buy a notebook; bring it to meetings; make notes! 
- Come to meetings with a written agenda. 
- While doing research, record what you did and what the outcome was.
- Use version control ([git](http://git-scm.com)) and internet repositories ([bitbucket](http://www.bitbucket.org), [github](http://www.github.com)) to organize notes, code, etc.
- Use *knitr* to help make your research reproducible.
- Talk about the tools you find useful!

## What are these repository things?

- A repository: a central place in which an aggregation of data is kept and maintained in an organized way (searcharticle.com)
- Great for sharing material across multiple people (eg student and advisor!)
- An amateur example: [http://github.com/stephens999/ash](http://github.com/stephens999/ash)

## What is knitr?

- An R package
- A tool for literate programming
- Text, and R code are interleaved
- When you compile the document, the code is run, and output inserted into the text.
- Great for writing reports, and keeping a track of what you did and what the result was!
- This talk was written with knitr (with RStudio)!

## What is knitr?

 ![Image](screenshot.png)

## What is Reproducible Research?

- Principle: when publishing results of computational procedures, we should
publish the code that produced the results.
- "publishing figures or results without the complete software environment could
be compared to a mathematician publishing an announcement of a mathematical theorem without giving the proof" (Buckheit and Donohoe)
- “an article about a computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result.” [Claerbout]

## Why is reproducibility important?

- Not only because people are forgetful, error-prone, or dishonest!
- Reproducing work is also the first step to extending it.
- Helps communications among researchers (eg student + advisor).
- If you do not publish code implementing your methods, your methods will likely go unused.

## More on git, github, knitr, reproducibility

- Google "The git book", to get started on git.
- Google "Karl Broman github tutorial" for statistics-oriented intro to github.
- Google "donohoe buckheit" for "Wavelab and reproducible research"

## More ridiculously-personal tips

- "Don't try to show them how clever you are" (BD Ripley)
- Minimise email for research communications 
- Chunk time
- Relax! (Exercise, Vacation)
- Admin on Fridays
- Learn some programming principles
- Benchmark early and often

## Minimise email

![Image](slack.screenshot.png)

## Benchmarking:

By which I mean, any quantitative assessment and comparisons of performance (accuracy; speed).

## Benchmarking:

Benefits include:

- Communication
- Tweaking/honing
- Basic quality assurance (cf "testing")

## Benchmarking:

Even basic benchmarks have benefits (...Done is better than perfect)

## Dynamic Statistical Comparisons

- Overall Goal: Improve how we, in the statistics community, make comparisons among methods.

- Reduce work and redundancy; make comparisons more transparent

- How? Centralize reproducible statistical comparisons in easily-extensible public repositories.

## Inspiration: Public Benchmarking

- Protein folding (CASP)

- Image analysis (Pascal)

- The Neflix Prize

- Genetics Analysis Workshop

- RECOMB/DREAM challenge

- Sage Bionetworks

## Dynamic Statistical Comparisons: advantages

- Authors can run their own methods. 
- Ease of adding new data can quickly uncover over-tuning to specific datasets.
- Work is divided, and not duplicated.
- Promotes collaboration through competition. 
- Ensures results are reproducible, and extendable.

- [http://stephens999.github.io/blog/](http://stephens999.github.io/blog/) for more on DSC

## DSC Requirements

- Well-defined input format for methods
- Well-defined output format
- Scoring function (output, input, meta-data)
- Ways to produce input, meta-data
- Methods to turn input -> output

## DSCR - an R package

- Define Scenarios (produce input, meta-data)
- Define Methods (turn input -> output)
- Define Scores

A simple function call then runs all methods on all scenarios, outputs scores

[Simple Vignette](http://htmlpreview.github.com/?https://github.com/stephens999/dscr/blob/master/vignettes/one_sample_location.html)

## DSC examples we're working on

- Covariance Estimation
- Large-scale linear regression
- 1-D Smoothing with Wavelets [after Antoniadis et al, 2001]

## Wavelet Smoothing Example: The "Blocks" Function

```{r, echo=FALSE}
  blocks = function(n){
     t=1:n/n
    pos=c(.1, .13, .15, .23, .25, .40, .44, .65, .76, .78, .81)
    hgt = 2.88/5*c(4, (-5), 3, (-4), 5, (-4.2), 2.1, 4.3, (-3.1), 2.1, (-4.2))
    mu.blk = rep(0,n)
    for(j in 1:length(pos)){
      mu.blk = mu.blk + (1 + sign(t-pos[j]))*(hgt[j]/2)
    }
    mu=0.2+0.6*(mu.blk-min(mu.blk))/max(mu.blk-min(mu.blk))
    return(mu)
  }
len = 1024
x = seq(0,1,length=len)
mu=blocks(len)
plot(x,mu,type="l",col=2,lwd=2)
y=mu+rnorm(len,0,sd=sd(mu))
points(x,y,cex=0.5)
```

## Wavelet Transform of Noisy Data

```{r,echo=FALSE}
  y.wd = wavethresh::wd(y,filter=1,family="DaubExPhase")
  plot(y.wd,first.level=5) #Haar transform
```

## Wavelet Transform of True Mean

```{r, echo=FALSE}
mu.wd = wavethresh::wd(mu,filter=1,family="DaubExPhase")
plot(mu.wd,first.level=5) #Haar transform
```

## Wavelet Shrinkage

- WCs of true mean ($$\beta$$) are "sparse" (or at least, have spiky density about 0)

- WCs for observed data ($$\betahat$$) are noisy observations of true WCs.

- If we shrink the $$\betahat$$ toward 0, they will give a better estimate for $$\beta$$

- And inverting the wavelet transform will give better estimate for $\mu$

## Shrink + Backtransform = Smooth

```{r, echo=FALSE}
y.wd.thresh = wavethresh::threshold(y.wd, policy="BayesThresh")
y.est = wavethresh::wr(y.wd.thresh)
plot(x,mu,type="l",col=2,lwd=2)
lines(x,y.est,type="l")
```

## How to Shrink?

We've been developing generic EB approaches to shrinkage estimation:

- Assume $\beta_j \sim g()$ where $g$ has some concentration about 0 ("sparse").

- Assume we have corresponding estimates $\hat\beta_j$, with standard error $s_j$.

- 
## How to Shrink?

- *Incorporate precision* of each observation $\hat\beta$ by assuming
$$\hat\beta_j | \beta_j \sim N(\beta_j, s_j^2)$$

- Constrain $g$ to be *unimodal* about 0;

- Shrink via two-stage procedure: first estimate $g$ "non-parametrically" from data; then
compute $p(\beta_j | \hat{g}, \hat\beta_j,s_j)$.

## Details

- A convenient way to model $g$: mixture of 0-centered
normal distributions: 
$$g(\beta; \pi) = \sum_{k=1}^K \pi_k N(\beta; 0, \sigma^2_k)$$

- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm).

- By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
we get a flexible unimodal symmetric distribution.

- Can approximate, arbitrarily closely, any scale mixture of normals.
Includes almost all priors used for sparse regression problems (spike-and-slab, double exponential/Laplace/Bayesian Lasso, horseshoe).

## Illustration: $g$ a mixture of 0-centered normals

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, dnorm(x,0,1),type="l",ylim=c(0,2),ylab="density")
lines(x, dnorm(x,0,0.1))
lines(x, dnorm(x,0,0.5))
```

## Illustration: $g$ a mixture of 0-centered normals

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, 0.5*dnorm(x,0,1)+0.5*dnorm(x,0,0.1),type="l",ylim=c(0,2),ylab="density")
```

## Illustration: $g$ a mixture of 0-anchored uniforms

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, dunif(x,0,1),type="s",ylim=c(0,5),ylab="density")
lines(x, dunif(x,0,0.2),type="s")
lines(x, dunif(x,0,0.5),type="s")
lines(x, dunif(x,-0.3,0),type="s",col=2)
lines(x, dunif(x,-0.4,0),type="s",col=2)
lines(x, dunif(x,-2,0),type="s",col=2)
```

## Illustration: $g$ a mixture of 0-anchored uniforms

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, 0.1*dunif(x,0,3)+ 0.3*dunif(x,0,0.2) + 0.2*dunif(x,0,0.5) + 0.1* dunif(x,-0.3,0) + 0.1*dunif(x,-0.3,0)+0.2*dunif(x,-0.4,0),type="s",ylim=c(0,2),ylab="density")
```

## Adaptive Shrinkage

- Approach provides posterior distribution for each $\beta_j$.

- For wavelet shrinkage we just use posterior mean.

- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be "shrunk" towards the overall mean (0).

- Because $g(\beta)$ is estimated from the data, the amount
of shrinkage is adaptive to signal in the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each observation.

- So we call the approach "Adaptive Shrinkage" (ash).

## Smoothing with Adaptive Shrinkage (smash)

```{r, echo=FALSE, results='hide', message=FALSE}
load("../../dscr-smash/res.RData")
library("dplyr")
library("ggplot2")
res.1.v1 = res %>% filter(grepl(".1.v1",scenario)) %>% filter(method %in% c("smash.haar","smash.homo.s8","ebayesthresh","sure.homo.s8"))
PLOTNAMES=c("angles","blip","blocks","bump","corner","doppler","spike")
res.1.v1$scenario=factor(res.1.v1$scenario,levels=paste0(c("ang","blip","blk","bump","cor","dop","sp"),".1.v1"))
levels(res.1.v1$scenario)=PLOTNAMES
ggplot(res.1.v1,aes(x=method,y=mise,fill=method)) + geom_violin() + facet_grid(.~scenario) + theme(axis.text.x=element_blank())
```

## Homoskedastic errors

```{r, echo=FALSE, results='hide', message=FALSE}
res.1.v1 = res %>% filter(grepl(".1.v1",scenario)) %>% filter(method %in% c("smash.haar","smash.homo.s8","ebayesthresh"))

PLOTNAMES=c("angles","blip","blocks","bump","corner","doppler","spike")
res.1.v1$scenario=factor(res.1.v1$scenario,levels=paste0(c("ang","blip","blk","bump","cor","dop","sp"),".1.v1"))
levels(res.1.v1$scenario)=PLOTNAMES

ggplot(res.1.v1,aes(x=method,y=mise,fill=method)) + geom_violin() + facet_grid(.~scenario) +theme(axis.text.x=element_blank())
```

## Heteroskedastic errors

```{r, echo=FALSE, results='hide', message=FALSE}
res.1.v4 = res %>% filter(grepl(".1.v4",scenario)) %>% filter(method %in% c("smash.haar","smash.homo.s8","ebayesthresh"))
res.1.v4$scenario=factor(res.1.v4$scenario,levels=paste0(c("ang","blip","blk","bump","cor","dop","sp"),".1.v4"))
levels(res.1.v4$scenario)=PLOTNAMES
ggplot(res.1.v4,aes(x=method,y=mise,fill=method)) + geom_violin() + facet_grid(.~scenario) + theme(axis.text.x=element_blank())
```

## Question

- Does anyone have applications where allowing for heteroskedastic errors might be useful?

## Summary

- We're looking for conspirators interested in organizing benchmarking!

## Thanks

- Zhengrong Xing

- to the developers of **R**, **knitr**, **Rstudio** and **Pandoc**.
Also the very cool **SQUAREM** and **wavethresh** packages.

- to the several postdoctoral researchers and students
who have worked with me on related topics.

- Including Chaoxing Dai, Mengyin Lu, Ester Pantaleo, Scott Powers, Sen Tian, Wei Wang, .

- NHGRI, Moore Foundation for funding.

- `ashr` package: `http://www.github.com/stephens999/ashr`

- `dscr` package: `http://www.github.com/stephens999/dscr`