### 2015-06-23 examples/adaptive_shrinkage.rmd
update

---
title: "Adaptive Shrinkage Illustration"
output: html_document
---

The goal here is to illustrate the "adaptive" nature of the adaptive shrinkage. The shrinkage is adaptive in two senses. First,
the amount of shrinkage depends on the distribution $g$ of the true effects, which is learned from the data: when $g$ is very peaked about zero
then ash learns this and deduces that signals should be more strongly shrunk towards zero than when g is less peaked about zero.
Second, the amount of shrinkage of each observation depends on its standard error: the smaller the standard error, the more informative
the data, and so the less shrinkage that occurs. From an Empirical Bayesian point of view both of these 
points are entirely natural: the posterior depends on both the prior
and the likelihood; the prior, $g$, is learned from the data, and the likelihood incorporates the standard error of each observation.

We simulate from two scenarios, the first with effects more peaked about zero, the second with effects less peaked.
```{r}
set.seed(100)
library(ashr)
beta1 = c(rnorm(500,0,0.01),rnorm(500,0,0.5))
beta2 = rnorm(1000,0,1)
s = 1/rgamma(1000,10,10)
e = rnorm(1000,0,s)
beta1hat = beta1 + e
beta2hat = beta2 + e
```

Now run ash on both:
```{r}
beta1.ash = ash(beta1hat, s)
beta2.ash = ash(beta2hat, s)
```

```{r}
df = data.frame(betahat = c(beta1hat,beta2hat), beta_est = c(beta1.ash$PosteriorMean, beta2.ash$PosteriorMean),s=c(s,s),scenario=c(rep("peaked",1000),rep("less peaked",1000)))
library(ggplot2)
ggplot(df,aes(x=betahat,y=beta_est)) + geom_point(col=s) + facet_grid(.~scenario) 
#need to add equal scales, and color gradient

```