### 2015-06-23 talks/wavelet_vignette.rmd

---
title: "Wavelet Vignette"
author: Matthew Stephens
date: May 22, 2015
output: ioslides_presentation
---

## Example: The "Blocks"" Function

```{r, echo=FALSE}
  blocks = function(n){
     t=1:n/n
    pos=c(.1, .13, .15, .23, .25, .40, .44, .65, .76, .78, .81)
    hgt = 2.88/5*c(4, (-5), 3, (-4), 5, (-4.2), 2.1, 4.3, (-3.1), 2.1, (-4.2))
    mu.blk = rep(0,n)
    for(j in 1:length(pos)){
      mu.blk = mu.blk + (1 + sign(t-pos[j]))*(hgt[j]/2)
    }
    mu=0.2+0.6*(mu.blk-min(mu.blk))/max(mu.blk-min(mu.blk))
    return(mu)
  }
len = 1024
x = seq(0,1,length=len)
mu=blocks(len)
plot(x,mu,type="l",col=2,lwd=2)
y=mu+rnorm(len,0,sd=sd(mu))
points(x,y,cex=0.5)
```

## Wavelet Transform of Noisy Data

```{r,echo=FALSE}
  y.wd = wavethresh::wd(y,filter=1,family="DaubExPhase")
  plot(y.wd,first.level=5) #Haar transform
```

## Wavelet Transform of True Mean

```{r, echo=FALSE}
mu.wd = wavethresh::wd(mu,filter=1,family="DaubExPhase")
plot(mu.wd,first.level=5) #Haar transform
```

## Shrink + Backtransform = Smooth
```{r, echo=FALSE}
y.wd.thresh = wavethresh::threshold(y.wd, policy="BayesThresh")
y.est = wavethresh::wr(y.wd.thresh)
plot(x,mu,type="l",col=2,lwd=2)
lines(x,y.est,type="l")
```