### 2015-09-22 talks/shiny/shinytest.Rmd
update paper

---
title: Illustration of unimodal assumption and role of penalty in ashr
runtime: shiny
output: 
  ioslides_presentation:
    smaller: true
---

```{r, include=FALSE}
 #plot a histogram of z scores, highlighting the alternative distribution
#of z scores that is implied by pi0
nullalthist_z = function(z,pi0,nullcol="blue",altcol="cyan",...){
  h=hist(z, freq=FALSE,col=nullcol,nclass=50,ylim=c(-0.5,0.5),...)
  nb = length(h$breaks)
  nullexp = pi0 * (pnorm(h$breaks)[-1] - pnorm(h$breaks[-nb]))/(h$breaks[-1]-h$breaks[-nb])
  h$density = h$density - nullexp
  plot(h,add=TRUE,col=altcol,freq=FALSE)
}

 set.seed(100)
z = c(rnorm(5000,0,2),rnorm(5000,0,1))
```

##

Slide the slider to see how the proportion of nulls (pi0) changes the decomposition of $z$ scores into null (dark blue) and alternative (cyan).

If you increase pi0 until the alternative (cyan) just stops looking unimodal, this mimics the UA of ashr. The ashr penalty pushes pi0 as far as possible in this way.

If you increase pi0 until cyan hits 0 then you are making the ZA of locfdr and qvalue.
```{r, echo=FALSE}

inputPanel(  
  sliderInput("pi0", label = "Proportion of nulls:",
              min = 0, max = 1, value = 0, step = 0.1)
)

renderPlot({
  nullalthist_z(z, input$pi0)
})
```