### 2015-10-06 subsubsectionReorder.tex

\subsubsection*{Reordering of significance, and the ``$p$ value prior"}

Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
 with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
 We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
 The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
 and so the $\lfsr$ for these poor-precision measurements is always high.

The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
was previously discussed in \cite{guan.stephens.08}. However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores). He named this prior the ``$p$-value prior" because it can be thought of as the 
implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)

This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))]. Setting $\alpha=0$ implies that $\beta_j$ is
independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
effects (in absolute value). This latter assumption may often be qualitatively plausible: 
for example, in gene expression studies the standard error for gene $j$ depends partly on
the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
which will be dataset-specific. 
Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
which has been implemented in our software by C. Dai.