### 2015-10-13 examples/rosenblatt.rmd
added rosenblatt example

---
title: "mixture of zero and non-zero mean components"
output: html_document
---

Based on email exchange with Jonathan Rosenblatt (October 13, 2015) 
let's try to fit data to a model that is a 50-50 mixture of a distribution g0 centered at 0
and a distribution g1 centered at mu.

First I'll simulate from such a model:
```{r}
library(ashr)
set.seed(100)
ns = 1000 # number of samples
beta = c(rnorm(ns/2,0,1),rnorm(ns/2,3,1))
shat = 1 #just set noise to 1
betahat = rnorm(ns,beta,shat)
hist(betahat)
```

Now fit the model for various values of mu (I just used -1 to 5)
```{r}
mixsd = ashr:::autoselect.mixsd(betahat,shat,mult=sqrt(2)) #probably can improve choice of grid, but just to get something rolling
k= length(mixsd)

mu = seq(-1,5,length=7)
fit.ash=list()

for(i in 1:length(mu)){ #fit ash model for each mu
  g = normalmix(pi=rep(1/(2*k),2*k), mean = c(rep(0,k),rep(mu[i],k)), sd = c(mixsd,mixsd) )
  fit.ash[[i]] = ash(betahat,shat,g=g,prior="uniform")
}
```

Plot log-likelihood for different mu: it seems to be maximized near mu=3 which is promising.
```{r}
l=lapply(fit.ash,get_loglik)
plot(mu,l,xlab="mu",ylab="log-likelihood (maximized over g0,g1)")
```

Look at the estimated proportions of the zero vs 3 component
```{r}
x=seq(-1,5,length=100)
a = fit.ash[[which.max(l)]]
plot(get_density(a,x),type="l")
fit.ash[[which.max(l)]]
```